### 2012-09-16 ProjectProposal/Project.tex
Proposal

\documentclass[a4paper,10pt]{article}
%\documentclass[a4paper,10pt]{scrartcl}

\usepackage[utf8]{inputenc}

\title{Project Proposal Machine Learning}
\author{}
\date{}

\begin{document}
\maketitle

% 
% 
% > 1) Title project 
% > 2) Student member names with student numbers 
% > 3) Application, explain how you get data or simulator, how many examples, how 
% > many inputs, outputs, whether it is an RL, classification, or regression problem. 
% > 4) Methods. Explain the methods that you will compare. Identify the learning 
% > algorithm and the optimization method.

\paragraph{Title project:}
Structured Prediction for Named Entity Recognition

\paragraph{Student member names \& number:}

Joachim Daiber: 2397331\\
Carmen Klaussner: 2401541\\

\paragraph{Application}
We will obtain our training data from CoNLL-2003\cite{TjongKimSang:2003:ICS:1119176.1119195}, which contains
data in English and German.Additionally, we may also use the CoNLL-2002 data for Dutch and Spanish. 
Named Entity Recognition is a classification problem in which the goal is to correctly predict the named entity types for
the tokens in a text. Table blala shows the input/output distribution.

\begin{table}[h!]
\centering
 \begin{tabular}{l|l|l|l}
 \textbf{English Data}& Articles & Sentences & Tokens \\ \hline
 Training set &   946       &   14,987        &   203,621     \\
 Development set& 216        &  3,466         &    51,362     \\
 Test set       & 231        &   3,684        &     46,435    \\

 \end{tabular}
\end{table}

\begin{table}[h!]
\centering
 \begin{tabular}{l|l|l|l|l}
 \textbf{English Data} & LOC      & MISC     & ORG &PER       \\ \hline
 Training set &   7140       &  3438  & 6321     & 6600    \\
 Development set&  1837       &  922         &  1341    &  1842    \\
 Test set       &  1668       &   702        &   1661    &  1617   \\

 \end{tabular}
\end{table}

\begin{table}[h!]
\centering
 \begin{tabular}{l|l|l|l}
 \textbf{German Data} & Articles & Sentences & Tokens \\ \hline
 Training set &     553     &  12,705         &  206,931      \\ 
 Development set&    201     &  3,068         &  51,444       \\
 Test set       &    155     &   3,160        &   51,943      \\

 \end{tabular}
\end{table}

\begin{table}[h!]
\centering
 \begin{tabular}{l|l|l|l|l}
 \textbf{German Data} & LOC      & MISC     & ORG &PER       \\        \hline
 Training set &   4363       &   2288        & 2427     & 2773    \\
 Development set&  1181       &   1010        &  1241    &  1401    \\
 Test set       &   1035      &    670       &  773     &  1195   \\

 \end{tabular}
\end{table}

\paragraph{Methods:}
We believe that Structured Prediction provides a flexible and efficient model for Named Entity Recognition. 
The learning algorithm Structured Perceptron with Averaging \cite{Collins:2002:DTM:1118693.1118694}.

% > 6) Chosen programming language 
% > 7) Planning. Describe the different phases and the planning deadlines (until November).
% 
\paragraph{Setup of Experiments:}
We will compare multiple sets of features on the given training and test set, 
while trying to find a good balance between complexity and performance.

\paragraph{Chosen programming language:}
Python with the NumPy package

\paragraph{Planning:}
17-23 Sep.: Data Preparation/Literature Review

24-30 Sep.: Implementation of Learning Algorithm

1-7 Okt.: Implementation of Decoding

8-14 Okt.: Improvement of Features

15-21 Okt.: Evaluation

22-28 Okt.: Paper

29-4 Nov.: Paper

{\small
\bibliographystyle{plain}
\bibliography{ML.bib}
}

\end{document}

### 2012-09-16 ProjectProposal/Project.tex
Changed planing table.

@@
 
+\section*{Project Proposal}
+
 \paragraph{Title project:}
@@
 
-\paragraph{Student member names \& number:}
-
+\paragraph{Student member names \& number:}\\
 Joachim Daiber: 2397331\\
@@
 We believe that Structured Prediction provides a flexible and efficient model for Named Entity Recognition. 
-The learning algorithm Structured Perceptron with Averaging \cite{Collins:2002:DTM:1118693.1118694}.
+The learning algorithm is the Structured Perceptron with Averaging \cite{Collins:2002:DTM:1118693.1118694}.
 
@@
 
-\paragraph{Planning:}
-17-23 Sep.: Data Preparation/Literature Review
+\section*{Planning}
 
-24-30 Sep.: Implementation of Learning Algorithm
-
-1-7 Okt.: Implementation of Decoding
-
-8-14 Okt.: Improvement of Features
-
-15-21 Okt.: Evaluation
-
-22-28 Okt.: Paper
-
-29-4 Nov.: Paper
+\begin{table}[h!]
+\begin{tabular}{l l}
+17-23 Sep. & Data Preparation/Literature Review \\
+24-30 Sep. & Implementation of Learning Algorithm\\
+1-7 Oct.   & Implementation of Decoding\\
+8-14 Oct.  & Improvement of Features\\
+15-21 Oct. & Evaluation \\
+22-28 Oct. & Paper \\
+29-4 Nov.  & Paper\\
+\end{tabular}
+\end{table}
 

### 2012-09-17 ProjectProposal/Project.tex
Minor layout changes, wording, sources.

@@
 \date{}
+
+\usepackage{hyperref}
+\usepackage{xcolor}
+\definecolor{dark-red}{rgb}{0.4,0.15,0.15}
+\definecolor{dark-blue}{rgb}{0.15,0.15,0.4}
+\definecolor{medium-blue}{rgb}{0,0,0.5}
+\hypersetup{
+    colorlinks, linkcolor={dark-red},
+    citecolor={dark-blue}, urlcolor={medium-blue}
+}
 
@@
 
-\paragraph{Title project:}
+\paragraph{Project title:}
 Structured Prediction for Named Entity Recognition
 
-\paragraph{Student member names \& number:}\\
-Joachim Daiber: 2397331\\
-Carmen Klaussner: 2401541\\
+\paragraph{Student member names \& number:\\\\}
+
+\begin{tabular}{l l}
+Joachim Daiber   & 2397331 \\
+Carmen Klaussner & 2401541\\
+\end{tabular}
 
 \paragraph{Application}
-We will obtain our training data from CoNLL-2003\cite{TjongKimSang:2003:ICS:1119176.1119195}, which contains
-data in English and German.Additionally, we may also use the CoNLL-2002 data for Dutch and Spanish. 
+We will obtain our training data from the CoNLL-2003\cite{TjongKimSang:2003:ICS:1119176.1119195} dataset, which contains
+data in English and German. Additionally, we may also use the CoNLL-2002 data for Dutch and Spanish.
+
 Named Entity Recognition is a classification problem in which the goal is to correctly predict the named entity types for
-the tokens in a text. Table blala shows the input/output distribution.
+the tokens in a text.
+Table~\ref{table:input} shows the size of the input data and Table~\ref{table:output} shows the types and amount of Named Entity annotations on the input data for German and English.
 
 \begin{table}[h!]
+
 \centering
@@
  \end{tabular}
+
+\vspace*{0.2cm}
+
+ \begin{tabular}{l|l|l|l}
+ \textbf{German Data} & Articles & Sentences & Tokens \\ \hline
+ Training set &     553     &  12,705         &  206,931      \\ 
+ Development set&    201     &  3,068         &  51,444       \\
+ Test set       &    155     &   3,160        &   51,943      \\
+
+ \end{tabular}
+\caption{Sizes of the training, development and test sets.}
+\label{table:input}
+
 \end{table}
@@
 \begin{table}[h!]
+
 \centering
@@
  \end{tabular}
-\end{table}
 
-\begin{table}[h!]
-\centering
- \begin{tabular}{l|l|l|l}
- \textbf{German Data} & Articles & Sentences & Tokens \\ \hline
- Training set &     553     &  12,705         &  206,931      \\ 
- Development set&    201     &  3,068         &  51,444       \\
- Test set       &    155     &   3,160        &   51,943      \\
+\vspace*{0.2cm}
 
- \end{tabular}
-\end{table}
-
-\begin{table}[h!]
 \centering
@@
  \end{tabular}
+\caption{Named Entity annotations in the training, development and test sets.}
+\label{table:output}
+
 \end{table}
@@
 \paragraph{Methods:}
-We believe that Structured Prediction provides a flexible and efficient model for Named Entity Recognition. 
+We believe that Structured Prediction provides a flexible and efficient model for Named Entity Recognition (see \cite{strlearn}).
 The learning algorithm is the Structured Perceptron with Averaging \cite{Collins:2002:DTM:1118693.1118694}.
@@
 \begin{tabular}{l l}
-17-23 Sep. & Data Preparation/Literature Review \\
+17-23 Sep. & Data Preparation and Literature Review \\
 24-30 Sep. & Implementation of Learning Algorithm\\

### 2012-09-17 ProjectProposal/Project.tex
Wording.

@@
 \paragraph{Application}
-We will obtain our training data from the CoNLL-2003\cite{TjongKimSang:2003:ICS:1119176.1119195} dataset, which contains
-data in English and German. Additionally, we may also use the CoNLL-2002 data for Dutch and Spanish.
 
@@
 the tokens in a text.
+
+We will obtain our training data from the CoNLL-2003\cite{TjongKimSang:2003:ICS:1119176.1119195} dataset, which contains
+data in English and German.
+Additionally, we may also use the CoNLL-2002 data for Dutch and Spanish.
 Table~\ref{table:input} shows the size of the input data and Table~\ref{table:output} shows the types and amount of Named Entity annotations on the input data for German and English.

### 2012-09-17 ProjectProposal/Project.tex
Last changes.

@@
 
-\paragraph{Application}
+\paragraph{Application:}
 
@@
 
-We will obtain our training data from the CoNLL-2003\cite{TjongKimSang:2003:ICS:1119176.1119195} dataset, which contains
+We will obtain our training data from the CoNLL-2003 dataset \cite{TjongKimSang:2003:ICS:1119176.1119195}, which contains
 data in English and German.
 Additionally, we may also use the CoNLL-2002 data for Dutch and Spanish.
-Table~\ref{table:input} shows the size of the input data and Table~\ref{table:output} shows the types and amount of Named Entity annotations on the input data for German and English.
+Table~\ref{table:input} shows the size of the input data and Table~\ref{table:output} shows the types and counts of Named Entity annotations on the input data for German and English.
 

### 2012-10-19 doc/report/paper.tex
Paper.

--- 2012-09-17 ProjectProposal/Project.tex
@@
 
-\title{Project Proposal Machine Learning}
+\title{Named Entity Recognition using Structured Prediction}
 \author{}
@@
 
-% 
-% 
-% > 1) Title project 
-% > 2) Student member names with student numbers 
-% > 3) Application, explain how you get data or simulator, how many examples, how 
-% > many inputs, outputs, whether it is an RL, classification, or regression problem. 
-% > 4) Methods. Explain the methods that you will compare. Identify the learning 
-% > algorithm and the optimization method.
-
-\section*{Project Proposal}
-
-\paragraph{Project title:}
-Structured Prediction for Named Entity Recognition
-
-\paragraph{Student member names \& number:\\\\}
-
-\begin{tabular}{l l}
-Joachim Daiber   & 2397331 \\
-Carmen Klaussner & 2401541\\
-\end{tabular}
-
-\paragraph{Application}
-
-Named Entity Recognition is a classification problem in which the goal is to correctly predict the named entity types for
-the tokens in a text.
-
-We will obtain our training data from the CoNLL-2003\cite{TjongKimSang:2003:ICS:1119176.1119195} dataset, which contains
-data in English and German.
-Additionally, we may also use the CoNLL-2002 data for Dutch and Spanish.
-Table~\ref{table:input} shows the size of the input data and Table~\ref{table:output} shows the types and amount of Named Entity annotations on the input data for German and English.
-
-\begin{table}[h!]
-
-\centering
- \begin{tabular}{l|l|l|l}
- \textbf{English Data}& Articles & Sentences & Tokens \\ \hline
- Training set &   946       &   14,987        &   203,621     \\
- Development set& 216        &  3,466         &    51,362     \\
- Test set       & 231        &   3,684        &     46,435    \\
-
- \end{tabular}
-
-\vspace*{0.2cm}
-
- \begin{tabular}{l|l|l|l}
- \textbf{German Data} & Articles & Sentences & Tokens \\ \hline
- Training set &     553     &  12,705         &  206,931      \\ 
- Development set&    201     &  3,068         &  51,444       \\
- Test set       &    155     &   3,160        &   51,943      \\
-
- \end{tabular}
-\caption{Sizes of the training, development and test sets.}
-\label{table:input}
-
-\end{table}
-
-\begin{table}[h!]
-
-\centering
- \begin{tabular}{l|l|l|l|l}
- \textbf{English Data} & LOC      & MISC     & ORG &PER       \\ \hline
- Training set &   7140       &  3438  & 6321     & 6600    \\
- Development set&  1837       &  922         &  1341    &  1842    \\
- Test set       &  1668       &   702        &   1661    &  1617   \\
-
- \end{tabular}
-
-\vspace*{0.2cm}
-
-\centering
- \begin{tabular}{l|l|l|l|l}
- \textbf{German Data} & LOC      & MISC     & ORG &PER       \\        \hline
- Training set &   4363       &   2288        & 2427     & 2773    \\
- Development set&  1181       &   1010        &  1241    &  1401    \\
- Test set       &   1035      &    670       &  773     &  1195   \\
-
- \end{tabular}
-\caption{Named Entity annotations in the training, development and test sets.}
-\label{table:output}
-
-\end{table}
-
-\paragraph{Methods:}
-We believe that Structured Prediction provides a flexible and efficient model for Named Entity Recognition (see \cite{strlearn}).
-The learning algorithm is the Structured Perceptron with Averaging \cite{Collins:2002:DTM:1118693.1118694}.
-
-% > 6) Chosen programming language 
-% > 7) Planning. Describe the different phases and the planning deadlines (until November).
-% 
-\paragraph{Setup of Experiments:}
-We will compare multiple sets of features on the given training and test set, 
-while trying to find a good balance between complexity and performance.
-
-\paragraph{Chosen programming language:}
-Python with the NumPy package
-
-\section*{Planning}
-
-\begin{table}[h!]
-\begin{tabular}{l l}
-17-23 Sep. & Data Preparation and Literature Review \\
-24-30 Sep. & Implementation of Learning Algorithm\\
-1-7 Oct.   & Implementation of Decoding\\
-8-14 Oct.  & Improvement of Features\\
-15-21 Oct. & Evaluation \\
-22-28 Oct. & Paper \\
-29-4 Nov.  & Paper\\
-\end{tabular}
-\end{table}
-
 {\small
 \bibliographystyle{plain}
-\bibliography{ML.bib}
+\bibliography{paper.bib}
 }

### 2012-10-21 src/structured_ner/case/MosesTrueCaser.py
Adding TrueCasing, extended features, added error analysis.

@@ -1,0 +1,51 @@
+from copy import copy
+import nltk
+from TrueCaser import TrueCaser
+
+class MosesTrueCaser(TrueCaser):
+
+    """
+    This is a Python port of the truecasing Perl script in the Moses SMT system:
+
+    http://code.google.com/p/smt/source/browse/trunk/moses64/tools/moses-scripts/recaser/truecase.perl?r=32
+
+    """
+
+    def __init__(self, model_file):
+
+        self.best  = {}
+        self.known = {}
+
+        for line in model_file:
+            entry = line.split(" ")
+
+            self.best[ entry[0].lower() ] = entry[0]
+            self.known[ entry[0] ] = 1
+
+            for i in range(2, len(entry), 2):
+                self.known[ entry[i] ] = 1
+
+        model_file.close()
+
+    def case(self, tokens, tags):
+
+        l_tokens = map(lambda s: s.lower(), tokens)
+        cased_tokens = copy(l_tokens)
+
+        for i in range(len(tokens)):
+            word = l_tokens[i]
+
+            if i == 0 and word.lower() in self.best:
+                #Truecase sentence start
+                cased_tokens[i] = self.best[ word.lower() ]
+            elif word in self.known:
+                #don't change known words
+                cased_tokens[i] = word
+            elif word.lower() in self.best:
+                # truecase otherwise unknown words
+                cased_tokens[i] = self.best[ word.lower() ]
+            else:
+                # unknown, nothing to do
+                cased_tokens[i] = word
+
+        return cased_tokens

### 2012-10-21 src/structured_ner/demo.py

@@ -11,58 +11,97 @@ import nltk
 def index():
     return """
 
+
+
+
 <!DOCTYPE html PUBLIC "-//W3C//DTD XHTML 1.0 Strict//EN"
-	"http://www.w3.org/TR/xhtml1/DTD/xhtml1-strict.dtd">
+    "http://www.w3.org/TR/xhtml1/DTD/xhtml1-strict.dtd">
 
 <html xmlns="http://www.w3.org/1999/xhtml" xml:lang="en" lang="en">
 <head>
-	<meta http-equiv="Content-Type" content="text/html; charset=utf-8"/>
+    <meta http-equiv="Content-Type" content="text/html; charset=utf-8"/>
 
-	<title>NER Demo</title>
+    <title>NER Demo</title>
 
-	<script src="http://ajax.googleapis.com/ajax/libs/jquery/1.8.2/jquery.min.js"></script>
-	<script type="text/javascript" charset="utf-8">
-	function update() {
-	  var t = $("textarea#input").val();
-	  $("#out").load("http://localhost:8086/ner/"+ encodeURI(t));
-	}
+    <script src="http://ajax.googleapis.com/ajax/libs/jquery/1.8.2/jquery.min.js"></script>
+    <script type="text/javascript" charset="utf-8">
+    function update() {
+      var t = $("textarea#input").val();
+      $("#out").load("http://localhost:8086/ner/"+ encodeURI(t));
+    }
 
     $(document).ready(function() {
-	//setup before functions
-	var typingTimer;                //timer identifier
-	var doneTypingInterval = 700;  //time in ms, 5 second for example
+    //setup before functions
+    var typingTimer;                //timer identifier
+    var doneTypingInterval = 700;  //time in ms, 5 second for example
 
-	//on keyup, start the countdown
-	$('#input').keyup(function(){
-	    typingTimer = setTimeout(update, doneTypingInterval);
-	});
+    //on keyup, start the countdown
+    $('#input').keyup(function(){
+        typingTimer = setTimeout(update, doneTypingInterval);
+    });
 
-	//on keydown, clear the countdown
-	$('#input').keydown(function(){
-	    clearTimeout(typingTimer);
-	});
-	});
+    //on keydown, clear the countdown
+    $('#input').keydown(function(){
+        clearTimeout(typingTimer);
+    });
+    });
 
-	</script>
+    </script>
+
[…]

### 2012-10-21 src/structured_ner/evaluate.py

@@ -3,9 +3,11 @@ from nltk.corpus import conll2002
 from nltk.corpus.reader.conll import ConllChunkCorpusReader
 from nltk.corpus.util import LazyCorpusLoader
 from sentence import load_conll
+from nltk.stem.wordnet import WordNetLemmatizer
 from features.features_labels import LabelInteractionFeatures
 from features.features_node import SimpleNodeFeatures
 from features.features_gazetteer import GazetteerFeatures
+from case.MosesTrueCaser import MosesTrueCaser
 from train_supervised import train_ner
 
 conll2002_labels = ['LOC', 'PER', 'ORG', 'MISC', 'O']
@@ -15,31 +17,48 @@ conll2003 = LazyCorpusLoader('conll2003', ConllChunkCorpusReader, '.*\.(test|tra
 # Feature sets and Gazetteers for Dutch
 ########################################
 
-nl_gazetteer = GazetteerFeatures(['data/loc_nl.txt', 'data/person_nl.txt', 'data/org_nl.txt'])
+#nl_gazetteer = GazetteerFeatures(['data/loc_nl.txt', 'data/person_nl.txt', 'data/org_nl.txt'])
 
-dutch_train     = load_conll(conll2002.chunked_sents('ned.train'))
-dutch_heldout   = load_conll(conll2002.chunked_sents('ned.testa'))
-dutch_test      = load_conll(conll2002.chunked_sents('ned.testb'))
+#dutch_train     = load_conll(conll2002.chunked_sents('ned.train'))[:6000]
+#dutch_heldout   = load_conll(conll2002.chunked_sents('ned.testa'))[:1000]
+#dutch_test      = load_conll(conll2002.chunked_sents('ned.testb'))
 
-dutch_ner        = train_ner('nl', conll2002_labels, dutch_train, dutch_heldout, dutch_test, [SimpleNodeFeatures(), LabelInteractionFeatures(), nl_gazetteer], verbose=True)
-dutch_ner2       = train_ner('nl', conll2002_labels, dutch_train, dutch_heldout, dutch_test, [SimpleNodeFeatures(), LabelInteractionFeatures()], verbose=True)
-dutch_ner1       = train_ner('nl', conll2002_labels, dutch_train, dutch_heldout, dutch_test, [SimpleNodeFeatures()], verbose=True)
[…]

### 2012-10-21 src/structured_ner/features/features_gazetteer.py

@@ -1,4 +1,7 @@
+from __future__ import division
 import codecs
+from collections import defaultdict
+import sys
 from feature_generator import FeatureSet
 
 class GazetteerFeatures(FeatureSet):
@@ -21,9 +24,38 @@ class GazetteerFeatures(FeatureSet):
                         self.gazetteer_tokens[token].append(g_file)
 
 
+    def filter(self, corpus):
+
+        print >>sys.stderr, "Filtering %d Gaz. tokens" % len(self.gazetteer_tokens)
+
+        token_NE_count = defaultdict(int)
+        token_O_count  = defaultdict(int)
+
+        for sent in corpus:
+            for (token_tag, label) in zip(sent.x, sent.y):
+                token, tag = token_tag
+                if label == 'O':
+                    token_O_count[token]  += 1
+                else:
+                    token_NE_count[token] += 1
+
+        N = len(set(token_O_count.keys() + token_NE_count.keys()))
+
+        for token in self.gazetteer_tokens.keys():
+
+            pNE = (token_NE_count[token] + 1) / (token_NE_count[token] + token_O_count[token] + N)
+            pO  = (token_O_count[token] + 1)  / (token_NE_count[token] + token_O_count[token] + N)
+
+            #if P(NE|token) < P(O|token): remove it
+            if pNE < pO:
+                del self.gazetteer_tokens[token]
+
+        print >>sys.stderr, "Tokens after filtering: %d" % len(self.gazetteer_tokens)
+
 
     def generate_features(self, feature_storage, sequence, i, last_label, label):
-        token, pos_tag = sequence.x[i]
+        _, pos_tag = sequence.x[i]
+        token = sequence.true_case[i]
         features_fired = []
 
         if token in self.gazetteer_tokens:

### 2012-10-21 src/structured_ner/features/features_node.py

@@ -4,9 +4,18 @@ from feature_generator import FeatureSet
 
 class SimpleNodeFeatures(FeatureSet):
 
+    initial_pattern = re.compile(r"[A-Z]\.")
+    digit_pattern_2 = re.compile(r"[0-9]{2}")
+    digit_pattern_4 = re.compile(r"[0-9]{4}")
+
+
+    def __init__(self):
+        self.lemmatizer = None
+
     @lru_cache(maxsize=100)
     def generate_features(self, feature_storage, sequence, i, last_label, label):
-        token, pos_tag = sequence.x[i]
+        _, pos_tag = sequence.x[i]
+        token = sequence.true_case[i]
 
         features_fired = []
 
@@ -14,9 +23,39 @@ class SimpleNodeFeatures(FeatureSet):
         feature_id = feature_storage.add_feature("label:%s" % label)
         features_fired.append(feature_id)
 
-        #Uppercase?:
-        if token[0].isupper():
-            feature_id = feature_storage.add_feature("capital_first-%s" % label)
+        #First word.
+        if i == 0:
+            feature_id = feature_storage.add_feature("first_word-%s" % label)
+            features_fired.append(feature_id)
+
+        #In quotes?
+        if i < len(sequence)-1 and sequence.x[i-1] == '"' and sequence.x[i+1] == '"':
+            feature_id = feature_storage.add_feature("in_quotes-%s" % label)
+            features_fired.append(feature_id)
+
+        #Numbers?
+        if self.digit_pattern_2.match(token):
+            feature_id = feature_storage.add_feature("2_digit_number-%s" % label)
+            features_fired.append(feature_id)
+
+        if self.digit_pattern_4.match(token):
+            feature_id = feature_storage.add_feature("2_digit_number-%s" % label)
+            features_fired.append(feature_id)
+
+
+        #Initial of name?
+        if self.initial_pattern.match(token):
+            feature_id = feature_storage.add_feature("name_initial-%s" % label)
+            features_fired.append(feature_id)
+
+        #Uppercase at start of sentence?:
+        if token[0].isupper() and i == 0:
+            feature_id = feature_storage.add_feature("capital_first_sentence_initial-%s" % label)
+            features_fired.append(feature_id)
+
+        #Uppercase within sentence?:
+        if token[0].isupper() and i > 0:
+            feature_id = feature_storage.add_feature("capital_first_non_sentence_initial-%s" % label)
             features_fired.append(feature_id)
 
[…]

### 2012-10-21 src/structured_ner/linear_classifier.py

@@ -1,6 +1,8 @@
 from __future__ import division
 import numpy as np
 import sys
+from setuptools.command.sdist import entities
+from util import entities_from_list
 
 class LinearClassifier(object):
 
@@ -36,27 +38,30 @@ class LinearClassifier(object):
         total = 0
         incorrect = 0
 
-        tp_for_label = dict([ (l, 0) for l in self.labels ])
-        fp_for_label = dict([ (l, 0) for l in self.labels ])
-        fn_for_label = dict([ (l, 0) for l in self.labels ])
+        errors = []
+
+        gold_count_for_label      = dict([ (l, 0) for l in self.labels ])
+        predicted_count_for_label = dict([ (l, 0) for l in self.labels ])
+        correct_count_for_label   = dict([ (l, 0) for l in self.labels ])
 
         for sentence in sentences:
             z = self.viterbi_decode(sentence)
 
-            for i in xrange(len(sentence.x)):
-                total += 1.
+            for i in range(len(z)):
+                if z[i] != sentence.y[i]:
+                    errors.append( sentence.x[i][0] + ' ' + sentence.x[i][1] + ' ' + sentence.true_case[i] + ' ' + sentence.y[i] + ' ' + z[i] )
 
-                if sentence.y[i] != z[i]:
-                    incorrect += 1
+            total += len(sentence.y)
+            incorrect += len( [i for i in range(len(sentence.y)) if sentence.y[i] != z[i]])
 
-                    if sentence.y[i] == 'O':
-                        fp_for_label[ z[i] ] += 1
+            gold_entities      = entities_from_list(sentence.y)
+            predicted_entities = entities_from_list(z)
 
-                    if z[i] == 'O':
-                        fn_for_label[ sentence.y[i] ] += 1
-
-                elif z[i] != 'O':
-                    tp_for_label[ z[i] ] += 1
+            for l in self.labels:
+                if l != 'O':
+                    gold_count_for_label[l]      += len( [ e for e in gold_entities      if e[2] == l] )
+                    predicted_count_for_label[l] += len( [ e for e in predicted_entities if e[2] == l] )
+                    correct_count_for_label[l]   += len( [ e for e in gold_entities      if e[2] == l and e in predicted_entities] )
 
         out = "Test Accuracy: %f\n" % (1.0 - (incorrect/total))
 
@@ -67,17 +72,26 @@ class LinearClassifier(object):
             if l != 'O':
 
[…]

### 2012-10-21 src/structured_ner/ner.py

@@ -14,7 +14,7 @@ class StructuredNER(NamedEntityRecognizer):
         self.tokenizer = tokenizer
         self.tagger = tagger
 
-    def recognize(self, text):
-        tokens = self.tokenizer.tokenize(text)
+    def recognize(self, tokens_and_tags):
+        tokens = self.tokenizer.tokenize(tokens_and_tags)
         tags = self.tagger.tag(tokens)
         return zip(tokens, self.perceptron.viterbi_decode(Sentence(tags, [])))

### 2012-10-21 src/structured_ner/sentence.py

@@ -1,26 +1,41 @@
-def sentence_from_conll(chunked_sent):
+from __future__ import division
+
+def true_case_sentence(tokens, tags):
+    return sum([ (tokens[i].isupper() or (tags[i] in ['CD', '.', ',', ':'])) for i in range(len(tokens))]) / float(len(tokens)) >= 0.95
+
+def sentence_from_conll(chunked_sent, lemmatizer, truecaser):
     x = []
     y = []
 
     for element in chunked_sent:
         if type(element) == tuple:
-            x.append(element)
+            x.append( (element[0], element[1]) )
             y.append('O')
         else:
             for ne_token in element.leaves():
-                x.append(ne_token)
+                x.append( (ne_token[0], ne_token[1]) )
                 y.append(element.node)
 
-    return Sentence(x, y)
+    tokens, tags = zip(*x)
+    if truecaser is not None and true_case_sentence(tokens, tags):
+        true_case = truecaser.case(tokens, map(lambda e: e[1], x))
+    else:
+        true_case = tokens
 
-def load_conll(corpus):
-    return [ sentence_from_conll(sent) for sent in corpus if len(sent) > 0 ]
+    lemmas = map(lambda tc: lemmatizer.lemmatize(tc), true_case)
+
+    return Sentence(x, y, lemmas, true_case)
+
+def load_conll(corpus, lemmatizer, truecaser):
+    return [ sentence_from_conll(sent, lemmatizer, truecaser) for sent in corpus if len(sent) > 0 ]
 
 class Sentence:
 
-    def __init__(self, x, y):
-        self.x = x
-        self.y = y
+    def __init__(self, x, y, lemmas, true_case):
+        self.x         = x
+        self.y         = y
+        self.lemmas    = lemmas
+        self.true_case = true_case
 
     def __len__(self):
         return len(self.x)

### 2012-10-21 src/structured_ner/structured_perceptron.py

@@ -38,6 +38,10 @@ class StructuredPerceptron(LinearClassifier):
             accuracy = 1.0 - (incorrect/total)
             training_accuracy.append(accuracy)
 
+            #Stop if the error on the training data does not decrease
+            if training_accuracy[-1] <= training_accuracy[-2]:
+                break
+
             if verbose:
                 _, acc = self.test(heldout)
                 heldout_accuracy.append(acc)

### 2012-10-21 src/structured_ner/train_supervised.py

@@ -7,10 +7,15 @@ from structured_perceptron import StructuredPerceptron
 
 def train_ner(lang, labels, data, heldout, test, feature_sets, verbose=False, additional_run_label=None):
 
+    run_label = "%s_%s" % (lang, '+'.join(map(lambda fs: fs.__class__.__name__, feature_sets)))
+
+    if additional_run_label:
+        run_label += "_" + additional_run_label
+
     print >>sys.stderr, '\n' * 2
 
     print >>sys.stderr, '-' * 60
-    print >>sys.stderr, "Training: %s (%s)" % (lang, '+'.join(map(lambda fs: fs.__class__.__name__, feature_sets)))
+    print >>sys.stderr, "Training: %s (%s)" % (lang, run_label)
     print >>sys.stderr, '-' * 60
 
     print >>sys.stderr, "Extracting features from corpus..."
@@ -18,16 +23,11 @@ def train_ner(lang, labels, data, heldout, test, feature_sets, verbose=False, ad
 
     #Create the Perceptron
     print >>sys.stderr, "Creating the Perceptron..."
-    perceptron = StructuredPerceptron(labels, feature_generator, epochs=10)
+    perceptron = StructuredPerceptron(labels, feature_generator, epochs=15)
 
     #Train it!
     print >>sys.stderr, "Training the Perceptron..."
     random.shuffle(data)
-
-    run_label = "%s_%s" % (lang, '+'.join(map(lambda fs: fs.__class__.__name__, feature_sets)))
-
-    if additional_run_label:
-        run_label += "_" + additional_run_label
 
     perceptron.train(data, heldout, verbose=verbose, run_label=run_label)
 

### 2012-10-21 src/structured_ner/util.py

@@ -1,0 +1,24 @@
+def entities_from_list(list):
+    b = False
+    current_start = 0
+    current_tag = None
+
+    entities = []
+
+    for i in range(len(list)):
+      if list[i] == 'O' or i == len(list)-1 or (b and list[i] != current_tag):
+          if b:
+              entities.append( (current_start, i-1, current_tag) )
+              b = False
+              current_start = 0
+              current_tag = None
+
+      if list[i] != 'O' and not b:
+          b = True
+          current_start = i
+          current_tag = list[i]
+
+      if i == len(list)-1 and list[i] != 'O':
+          entities.append( (i, i, list[i]) )
+
+    return entities

### 2012-10-21 src/structured_ner/confMatrix.py
confusionmatrix

@@ -1,0 +1,35 @@
+import numpy as N
+import itertools
+
+class confMatrix():
+  def __init__(self, vals):
+    self.x = vals
+  
+ 
+
+  def confusion_matrix(self):
+    Ytrue = []
+    Ypred = []
+
+    
+    for i in self.x:
+      if len(i) != 0:
+	Ytrue.append(i[1])
+	Ypred.append(i[2])
+	  
+
+
+
+    classes = list(set(Ytrue))
+    n = len(classes)
+
+    #error = N.array([zip(Ytrue,Ypred).count(x) for x in itertools.product(classes,repeat=2)]).reshape(n,n)
+    #print error
+
+    error = N.array([z.count(x) for z in [zip(Ytrue,Ypred)] for x in itertools.product(classes,repeat=2)]).reshape(n,n)
+    print error
+    return error
+
+
+
+ 

### 2012-10-22 src/structured_ner/case/MosesTrueCaser.py
Changed Gazetteer to DAWG.

@@ -33,7 +33,7 @@ class MosesTrueCaser(TrueCaser):
         cased_tokens = copy(l_tokens)
 
         for i in range(len(tokens)):
-            word = l_tokens[i]
+            word = tokens[i]
 
             if i == 0 and word.lower() in self.best:
                 #Truecase sentence start
@@ -49,3 +49,6 @@ class MosesTrueCaser(TrueCaser):
                 cased_tokens[i] = word
 
         return cased_tokens
+
+t = MosesTrueCaser(open('/Users/jodaiber/Desktop/Groningen/structured-ner/src/structured_ner/models/truecase/truecase-model.en'))
+print t.case_pairs( nltk.pos_tag( "New York".split(' ')) )

### 2012-10-22 src/structured_ner/evaluate.py

@@ -46,12 +46,14 @@ for (lang, corpus) in [
     lemmatizer = WordNetLemmatizer()
     truecaser  = MosesTrueCaser(open('/Users/jodaiber/Desktop/Groningen/structured-ner/src/structured_ner/models/truecase/truecase-model.en'))
 
-    train   = load_conll(corpus.chunked_sents('%s.train' % lang), lemmatizer, None)
-    heldout = load_conll(corpus.chunked_sents('%s.testa' % lang), lemmatizer, truecaser)
-    test    = load_conll(corpus.chunked_sents('%s.testb' % lang), lemmatizer, truecaser)
+    gazetteer = GazetteerFeatures(['data/loc_%s.txt' % lang, 'data/person_%s.txt' % lang, 'data/org_%s.txt' % lang, 'data/misc_%s.txt' % lang], truecaser)
 
-    gazetteer = GazetteerFeatures(['data/loc_%s.txt' % lang, 'data/person_%s.txt' % lang, 'data/org_%s.txt' % lang])
-    gazetteer.filter(train)
+    train   = load_conll(corpus.chunked_sents('%s.train' % lang), lemmatizer, None,      gazetteer)
+    heldout = load_conll(corpus.chunked_sents('%s.testa' % lang), lemmatizer, truecaser, gazetteer)
+    test    = load_conll(corpus.chunked_sents('%s.testb' % lang), lemmatizer, truecaser, gazetteer)
+
+
+    #gazetteer.filter(train)
 
     node_features = SimpleNodeFeatures()
 
@@ -59,6 +61,5 @@ for (lang, corpus) in [
     pickle.dump(model_gaz, open("models/%s_gaz.pickle" % lang, 'w'))
 
     model     = train_ner(lang, conll2002_labels, train, heldout, test, [node_features, LabelInteractionFeatures()])
-
     pickle.dump(model, open("models/%s.pickle" % lang, 'w'))
 

### 2012-10-22 src/structured_ner/features/features_gazetteer.py

@@ -1,66 +1,85 @@
+"""
+>>> GazetteerFeatures([[u"The New York Times", u"New York"]]).find(u"The New York Times is a newspaper in New York .".split(' '))
+['0', '0', '0', '0', 'O', 'O', 'O', 'O', '0', '0', 'O']
+
+"""
+
 from __future__ import division
 import codecs
 from collections import defaultdict
 import sys
 from feature_generator import FeatureSet
+import dawg
+import struct
 
 class GazetteerFeatures(FeatureSet):
 
-    def __init__(self, gazetteer_files):
+    def __init__(self, gazetteer_files, truecaser):
 
-        self.gazetteer_tokens = {}
+        data = []
 
-        for g_file in gazetteer_files:
-            for entry in codecs.open(g_file, encoding='utf-8', mode='r'):
-                for token in entry.strip().split(' '):
+        if type(gazetteer_files[0]) != list:
+            for i in range(len(gazetteer_files)):
+                gazetteer_files[i] = codecs.open(gazetteer_files[i], encoding='utf-8', mode='r')
 
-                    if token.startswith('('):
-                        continue
+        for i in range(len(gazetteer_files)):
+            for entry in gazetteer_files[i]:
+                e = entry.rstrip()
+                data.append( (e, bytes(i)) )
 
-                    if token not in self.gazetteer_tokens:
-                        self.gazetteer_tokens[token] = []
+                #Only store true-case versions of 2+ token entries if they are different from the original e
+                if ' ' in entry:
+                    e_true_case = u' '.join(truecaser.case(entry.rstrip().split(' '), []))
 
-                    if g_file not in self.gazetteer_tokens[token]:
-                        self.gazetteer_tokens[token].append(g_file)
+                    if e_true_case != e:
+                        data.append( (e_true_case, bytes(i)) )
+
+        self.gazetteer_dawg = dawg.BytesDAWG(data)
+
+    def find(self, tokens):
+
+        matches = []
+
+        i = 0
+        while i < len(tokens):
+
+            j = i+1
+            while j <= len(tokens) and len( self.gazetteer_dawg.keys( u' '.join(tokens[i:j]) ) ) > 0:
+                j += 1
+
+            match_found = False
+            for z in reversed(range(i, j)):
+                try:
+                    matches.append( (i, z, self.gazetteer_dawg[ u' '.join(tokens[i:z]) ] ) )
+                    match_found = True
+                    break
+                except KeyError:
[…]

### 2012-10-22 src/structured_ner/features/features_labels.py

@@ -10,5 +10,10 @@ class LabelInteractionFeatures(FeatureSet):
         feature_id = feature_storage.add_feature("label_bigram:%s-%s" % (last_label, label))
         features_fired.append(feature_id)
 
+        #Last label and current POS tag:
+        if sequence.x[i][1] in ['IN', ')', '(', 'PRP', 'POS']:
+            feature_id = feature_storage.add_feature("labels_and_pos_tag:%s-%s-%s" % (last_label, sequence.x[i][0], label))
+            features_fired.append(feature_id)
+
         return features_fired
 

### 2012-10-22 src/structured_ner/gazetteer/words_in_category.sh

@@ -1,0 +1,3 @@
+#! /bin/sh
+
+curl -H "Accept: text/xml" "http://en.wiktionary.org/w/api.php?action=query&list=categorymembers&cmtitle=Category:$1&cmlimit=500&format=xml" | egrep -o '<cm pageid="[0-9]+" ns="[0-9]+" title="[^"]+"' | sed 's/.* title=\"\([^"]*\)\"$/\1/g' | grep -v ^Category: > $1

### 2012-10-22 src/structured_ner/sentence.py

@@ -1,9 +1,12 @@
 from __future__ import division
 
-def true_case_sentence(tokens, tags):
+def all_uppercase(tokens, tags):
     return sum([ (tokens[i].isupper() or (tags[i] in ['CD', '.', ',', ':'])) for i in range(len(tokens))]) / float(len(tokens)) >= 0.95
 
-def sentence_from_conll(chunked_sent, lemmatizer, truecaser):
+def headline(tokens, tags):
+    return sum([ not tokens[i].islower() for i in range(len(tokens))]) / float(len(tokens)) == 1.0 and tokens[-1] not in ('.', '?', '!') and len([ 1 for i in range(len(tokens)) if tags[i] not in ['CD', '.', ',', ':'] ]) >= 3
+
+def sentence_from_conll(chunked_sent, lemmatizer, truecaser, gazetteer):
     x = []
     y = []
 
@@ -17,25 +20,35 @@ def sentence_from_conll(chunked_sent, lemmatizer, truecaser):
                 y.append(element.node)
 
     tokens, tags = zip(*x)
-    if truecaser is not None and true_case_sentence(tokens, tags):
+    tokens, tags = list(tokens), list(tags)
+
+    if truecaser is not None and all_uppercase(tokens, tags):
+        true_case = truecaser.case(map(lambda x: x.lower(), tokens), map(lambda e: e[1], x))
+    elif truecaser is not None and headline(tokens, tags):
         true_case = truecaser.case(tokens, map(lambda e: e[1], x))
     else:
         true_case = tokens
 
+    if gazetteer:
+        gazetteer_entries = gazetteer.find( true_case )
+    else:
+        gazetteer_entries = map(lambda _: 'O', true_case)
+
     lemmas = map(lambda tc: lemmatizer.lemmatize(tc), true_case)
 
-    return Sentence(x, y, lemmas, true_case)
+    return Sentence(x, y, lemmas, true_case, gazetteer_entries)
 
-def load_conll(corpus, lemmatizer, truecaser):
-    return [ sentence_from_conll(sent, lemmatizer, truecaser) for sent in corpus if len(sent) > 0 ]
+def load_conll(corpus, lemmatizer, truecaser, gazetteer):
+    return [ sentence_from_conll(sent, lemmatizer, truecaser, gazetteer) for sent in corpus if len(sent) > 0 ]
 
 class Sentence:
 
-    def __init__(self, x, y, lemmas, true_case):
+    def __init__(self, x, y, lemmas, true_case, gazetteer_entries):
[…]

### 2012-10-22 src/structured_ner/setup.py

@@ -8,5 +8,5 @@ setup(name='Structured NER',
       #author='Joachim Daiber',
       #author_email='gward@python.net',
       url='https://github.com/jodaiber/structured-ner',
-      packages=['nltk', 'repoze.lru', 'bottle'],
+      packages=['nltk', 'repoze.lru', 'bottle', 'dawg'],
      )

### 2012-10-22 src/structured_ner/test.py

@@ -9,10 +9,17 @@ from case.MosesTrueCaser import MosesTrueCaser
 lemmatizer = WordNetLemmatizer()
 truecaser  = MosesTrueCaser(open('models/truecase/truecase-model.en'))
 
-m = pickle.load(open("models/_eng_gaz.pickle"))
+m = pickle.load(open("models/eng_gaz.pickle"))
+
+gazetteer = m.feature_generator.feature_sets[-1]
 
 corpus = LazyCorpusLoader('conll2003', ConllChunkCorpusReader, '.*\.(test|train).*', ('LOC', 'PER', 'ORG', 'MISC'), encoding='utf-8')
-test    = load_conll(corpus.chunked_sents('eng.testa'), lemmatizer, truecaser)
+test    = load_conll(corpus.chunked_sents('eng.testa'), lemmatizer, truecaser, gazetteer=gazetteer)
+
+for s in test:
+    print s.true_case
+    print s.gazetteer_entries
+    print
 
 out, acc = m.test(test)
 

### 2012-10-22 src/structured_ner/train_supervised.py

@@ -23,7 +23,7 @@ def train_ner(lang, labels, data, heldout, test, feature_sets, verbose=False, ad
 
     #Create the Perceptron
     print >>sys.stderr, "Creating the Perceptron..."
-    perceptron = StructuredPerceptron(labels, feature_generator, epochs=15)
+    perceptron = StructuredPerceptron(labels, feature_generator, epochs=16)
 
     #Train it!
     print >>sys.stderr, "Training the Perceptron..."

### 2012-10-23 src/structured_ner/demo.py
Finished demo, minor changes.

@@ -3,9 +3,11 @@ import pickle
 
 from bottle import route, run, template
 import webbrowser
+from nltk.stem.wordnet import WordNetLemmatizer
 
 from ner import StructuredNER
 import nltk
+from case.MosesTrueCaser import MosesTrueCaser
 
 @route('/')
 def index():
@@ -24,10 +26,13 @@ def index():
     <title>NER Demo</title>
 
     <script src="http://ajax.googleapis.com/ajax/libs/jquery/1.8.2/jquery.min.js"></script>
+
     <script type="text/javascript" charset="utf-8">
     function update() {
       var t = $("textarea#input").val();
-      $("#out").load("http://localhost:8086/ner/"+ encodeURI(t));
+      $("#out").load("http://localhost:8086/ner/"+ encodeURI(t), function() {
+        $("[rel=tooltip]").tooltip({html: true, placement: 'bottom'});
+      });
     }
 
     $(document).ready(function() {
@@ -48,7 +53,11 @@ def index():
 
     </script>
 
-    <link rel="stylesheet" href="http://twitter.github.com/bootstrap/1.4.0/bootstrap.min.css" type="text/css" charset="utf-8">
+    <script src="http://netdna.bootstrapcdn.com/twitter-bootstrap/2.1.1/js/bootstrap.min.js"></script>
+    <link href="http://netdna.bootstrapcdn.com/twitter-bootstrap/2.1.1/css/bootstrap-combined.min.css" rel="stylesheet">
+
+
+    <link rel="stylesheet" href="https://raw.github.com/drewwilson/TipTip/master/tipTip.css" type="text/css" charset="utf-8">
 
     <style type="text/css" media="screen">
     body {
@@ -59,14 +68,19 @@ def index():
         text-align: center;
     }
     #out {
-
+        display: block;
     }
     #input, #out {
         outline: none;
         padding: 10px;
-        display: inline-block;
+        display: block;
+        margin-left: -200px;
+
+        position: relative;
+        left: 50%;
         width: 400px;
         height: 100px;
+
         background: white;
         border: 0;
         border-radius: 7px;
@@ -76,8 +90,69 @@ def index():
     }
     #out {
         margin-top: 1em;
-        background: #fff;
-        border: 3px solid #f0f0f0;
+
+        line-height: 2em;
+
+        position: relative;
+        left: 50%;
+        width: 700px;
[…]

### 2012-10-23 src/structured_ner/evaluate.py

@@ -57,7 +57,7 @@ for (lang, corpus) in [
 
     node_features = SimpleNodeFeatures()
 
-    model_gaz = train_ner(lang, conll2002_labels, train, heldout, test, [node_features, LabelInteractionFeatures(), gazetteer])
+    model_gaz = train_ner(lang, conll2002_labels, train, heldout, test, [node_features, LabelInteractionFeatures(), gazetteer], verbose=True)
     pickle.dump(model_gaz, open("models/%s_gaz.pickle" % lang, 'w'))
 
     model     = train_ner(lang, conll2002_labels, train, heldout, test, [node_features, LabelInteractionFeatures()])

### 2012-10-23 src/structured_ner/ner.py

@@ -1,5 +1,9 @@
+# -*- coding: utf-8 -*-
+
 import pickle
+from traits.trait_types import self
 from sentence import Sentence
+from sentence import all_uppercase, headline
 
 class NamedEntityRecognizer:
 
@@ -9,12 +13,50 @@ class NamedEntityRecognizer:
 
 class StructuredNER(NamedEntityRecognizer):
 
-    def __init__(self, perceptron_model, tokenizer, tagger):
+    def __init__(self, perceptron_model, tokenizer, tagger, lemmatizer, truecaser):
         self.perceptron = pickle.load(perceptron_model)
         self.tokenizer = tokenizer
         self.tagger = tagger
+        self.truecaser = truecaser
+        self.lemmatizer = lemmatizer
+        self.gazetteer = self.perceptron.feature_generator.feature_sets[-1]
 
-    def recognize(self, tokens_and_tags):
-        tokens = self.tokenizer.tokenize(tokens_and_tags)
-        tags = self.tagger.tag(tokens)
-        return zip(tokens, self.perceptron.viterbi_decode(Sentence(tags, [])))
+        self.feature_by_id = {}
+
+        for (feature, id) in self.perceptron.feature_generator.feature_ids.items():
+            self.feature_by_id[id] = feature
+
+
+    def recognize(self, sentence):
+        tokens = self.tokenizer.tokenize( sentence )
+        x = self.tagger.tag(tokens)
+        _, tags = map(list, zip(*x))
+
+        if self.truecaser is not None and all_uppercase(tokens, tags):
+            true_case = self.truecaser.case(map(lambda x: x.lower(), tokens), map(lambda e: e[1], x))
+        elif self.truecaser is not None and headline(tokens, tags):
+            true_case = self.truecaser.case(tokens, map(lambda e: e[1], x))
+        else:
+            true_case = tokens
+
+        if self.gazetteer:
+            gazetteer_entries = self.gazetteer.find( true_case )
+        else:
+            gazetteer_entries = map(lambda _: 'O', true_case)
+
+        lemmas = map(lambda tc: self.lemmatizer.lemmatize(tc), true_case)
+
+        sentence = Sentence(x, [], lemmas, true_case, gazetteer_entries)
+
+        labels = self.perceptron.viterbi_decode(sentence)
+
+        features = []
+        for i in range(len(tokens)):
+            if i == 0:
+                last_label = "<S>"
+            else:
+                last_label = labels[i-1]
+
[…]

### 2012-10-23 src/structured_ner/test.py

@@ -16,11 +16,6 @@ gazetteer = m.feature_generator.feature_sets[-1]
 corpus = LazyCorpusLoader('conll2003', ConllChunkCorpusReader, '.*\.(test|train).*', ('LOC', 'PER', 'ORG', 'MISC'), encoding='utf-8')
 test    = load_conll(corpus.chunked_sents('eng.testa'), lemmatizer, truecaser, gazetteer=gazetteer)
 
-for s in test:
-    print s.true_case
-    print s.gazetteer_entries
-    print
-
 out, acc = m.test(test)
 
 codecs.open('test.txt', 'w', encoding='utf-8').write(out)

### 2012-10-23 src/structured_ner/case/MosesTrueCaser.py
Bugfix.

@@ -49,6 +49,3 @@ class MosesTrueCaser(TrueCaser):
                 cased_tokens[i] = word
 
         return cased_tokens
-
-t = MosesTrueCaser(open('/Users/jodaiber/Desktop/Groningen/structured-ner/src/structured_ner/models/truecase/truecase-model.en'))
-print t.case_pairs( nltk.pos_tag( "New York".split(' ')) )

### 2012-10-28 src/structured_ner/demo.py
Latest version for demo, added Makefile for truecase models.

@@ -174,9 +174,9 @@ def index():
     """
 
 
-tokenizer = nltk.tokenize.RegexpTokenizer(r'\w+|[^\w\s]+')
-tagger = nltk.data.load('taggers/maxent_treebank_pos_tagger/english.pickle')
-truecaser = MosesTrueCaser(open('models/truecase/truecase-model.en'))
+tokenizer    = nltk.tokenize.RegexpTokenizer(r'\w+|[^\w\s]+')
+tagger       = nltk.data.load('taggers/maxent_treebank_pos_tagger/english.pickle')
+truecaser    = MosesTrueCaser(open('models/truecase/truecase-model.en'))
 nerecognizer = StructuredNER(open("models/eng_gaz.pickle"), tokenizer, tagger, WordNetLemmatizer(), truecaser)
 
 @route('/ner/:sent')
@@ -192,10 +192,6 @@ def ner(sent):
             f += ft.replace("<", "-").replace(">", "-") + "<br />"
 
         s += "<span class='ne_label' rel='tooltip' title='%s'>%s</span>" % (f, label)
-        #s += "<div class='features'><strong>Features:</strong><br />"
-
-
-        #s += "</div>"
 
         s += "</span>"
 

### 2012-10-28 src/structured_ner/evaluate.py

@@ -2,6 +2,7 @@ import pickle
 from nltk.corpus import conll2002
 from nltk.corpus.reader.conll import ConllChunkCorpusReader
 from nltk.corpus.util import LazyCorpusLoader
+from nltk.stem.snowball import SnowballStemmer
 from sentence import load_conll
 from nltk.stem.wordnet import WordNetLemmatizer
 from features.features_labels import LabelInteractionFeatures
@@ -17,34 +18,35 @@ conll2003 = LazyCorpusLoader('conll2003', ConllChunkCorpusReader, '.*\.(test|tra
 # Feature sets and Gazetteers for Dutch
 ########################################
 
-#nl_gazetteer = GazetteerFeatures(['data/loc_nl.txt', 'data/person_nl.txt', 'data/org_nl.txt'])
+lemmatizer = SnowballStemmer("dutch")
+truecaser  = MosesTrueCaser(open("models/truecase/nl"))
+nl_gazetteer = GazetteerFeatures(['data/loc_nl.txt', 'data/person_nl.txt', 'data/org_nl.txt'], truecaser)
 
-#dutch_train     = load_conll(conll2002.chunked_sents('ned.train'))[:6000]
-#dutch_heldout   = load_conll(conll2002.chunked_sents('ned.testa'))[:1000]
-#dutch_test      = load_conll(conll2002.chunked_sents('ned.testb'))
+dutch_train     = load_conll(conll2002.chunked_sents('ned.train'), lemmatizer, None, nl_gazetteer )
+dutch_heldout   = load_conll(conll2002.chunked_sents('ned.testa'), lemmatizer, truecaser, nl_gazetteer)
+dutch_test      = load_conll(conll2002.chunked_sents('ned.testb'), lemmatizer, truecaser, nl_gazetteer)
 
-#dutch_ner        = train_ner('nl', conll2002_labels, dutch_train, dutch_heldout, dutch_test, [SimpleNodeFeatures(), LabelInteractionFeatures(), nl_gazetteer], verbose=True)
-#pickle.dump(dutch_ner, open("models/%s.pickle" % 'nl', 'w'))
+dutch_ner        = train_ner('nl', conll2002_labels, dutch_train, dutch_heldout, dutch_test, [SimpleNodeFeatures(), LabelInteractionFeatures(), nl_gazetteer], verbose=True)
+pickle.dump(dutch_ner, open("models/%s.pickle" % 'nl', 'w'))
 
[…]

### 2012-10-28 src/structured_ner/linear_classifier2.py

@@ -1,111 +1,0 @@
-from __future__ import division
-import numpy as np
-import sys
-from setuptools.command.sdist import entities
-from util import entities_from_list
-
-class LinearClassifier(object):
-
-    def __init__(self, labels, feature_generator):
-        self.w = None
-        self.feature_generator = feature_generator
-        self.labels = labels
-
-    def fs(self, sentence, i, last_label, current_label):
-        return filter(lambda f: f != -1, self.feature_generator.generate(sentence, i, last_label, current_label))
-
-    def viterbi_decode(self, sentence, k=1):
-        alphas = np.zeros((len(sentence), len(self.labels)), dtype=float)
-        path = {}
-
-        for y in xrange(len(self.labels)):
-            path[y] = [y]
-            alphas[0][y] = np.sum(self.w[self.fs(sentence, 0, '<S>', self.labels[y])])
-
-        for i in xrange(1, len(sentence)):
-            new_path = {}
-
-            for y in xrange(len(self.labels)):
-                (alphas[i][y], b) = max([(alphas[i-1][y_last] + np.sum(self.w[self.fs(sentence, i, self.labels[y_last], self.labels[y])] ), y_last) for y_last in xrange(len(self.labels))])
-                new_path[y] = path[b] + [y]
-
-            path = new_path
-
-        return map(lambda i: self.labels[i], path[ alphas[-1:,].argmax() ])
-
-
-    def test(self, sentences):
-        total = 0
-        incorrect = 0
-
-        errors = []
-        comb = []
-
-        gold_count_for_label      = dict([ (l, 0) for l in self.labels ])
-        predicted_count_for_label = dict([ (l, 0) for l in self.labels ])
-        correct_count_for_label   = dict([ (l, 0) for l in self.labels ])
-
-        for sentence in sentences:
-            z = self.viterbi_decode(sentence)
-
-            for i in range(len(z)):
-                
-                if z[i] != sentence.y[i]:
-                    errors.append( sentence.x[i][0] + ' ' + sentence.x[i][1] + ' ' + sentence.true_case[i] + ' ' + sentence.y[i] + ' ' + z[i] )
-            
-                comb.append([sentence.x[i][0],sentence.y[i],z[i]]);
-            
-            
-            total += len(sentence.y)
-            incorrect += len( [i for i in range(len(sentence.y)) if sentence.y[i] != z[i]])
-
-            gold_entities      = entities_from_list(sentence.y)
[…]

### 2012-10-28 src/structured_ner/ner.py

@@ -1,7 +1,6 @@
 # -*- coding: utf-8 -*-
 
 import pickle
-from traits.trait_types import self
 from sentence import Sentence
 from sentence import all_uppercase, headline
 

### 2012-10-28 src/structured_ner/structured_perceptron.py

@@ -38,13 +38,13 @@ class StructuredPerceptron(LinearClassifier):
             accuracy = 1.0 - (incorrect/total)
             training_accuracy.append(accuracy)
 
-            #Stop if the error on the training data does not decrease
-            if training_accuracy[-1] <= training_accuracy[-2]:
-                break
-
             if verbose:
                 _, acc = self.test(heldout)
                 heldout_accuracy.append(acc)
+
+            #Stop if the error on the training data does not decrease
+            if training_accuracy[-1] <= training_accuracy[-2]:
+                break
 
             print >>sys.stderr, "Epoch %i, Accuracy: %f" % (i_epoch, accuracy)
 
@@ -61,7 +61,7 @@ class StructuredPerceptron(LinearClassifier):
 
         #Export training info in verbose mode:
         if verbose:
-            x = np.arange(0, self.n_epochs+1, 1.0)
+            x = np.arange(0, len(training_accuracy), 1.0)
             plt.plot(x, training_accuracy, marker='o', linestyle='--', color='r', label='Training')
             plt.plot(x, heldout_accuracy,  marker='o', linestyle='--', color='b', label='Heldout')
 

### 2012-10-28 src/structured_ner/test.py

@@ -9,9 +9,9 @@ from case.MosesTrueCaser import MosesTrueCaser
 lemmatizer = WordNetLemmatizer()
 truecaser  = MosesTrueCaser(open('models/truecase/truecase-model.en'))
 
-m = pickle.load(open("models/eng_gaz.pickle"))
+m = pickle.load(open("models/eng.pickle"))
 
-gazetteer = m.feature_generator.feature_sets[-1]
+gazetteer = None #m.feature_generator.feature_sets[-1]
 
 corpus = LazyCorpusLoader('conll2003', ConllChunkCorpusReader, '.*\.(test|train).*', ('LOC', 'PER', 'ORG', 'MISC'), encoding='utf-8')
 test    = load_conll(corpus.chunked_sents('eng.testa'), lemmatizer, truecaser, gazetteer=gazetteer)

### 2012-10-28 src/structured_ner/train_supervised.py

@@ -31,10 +31,8 @@ def train_ner(lang, labels, data, heldout, test, feature_sets, verbose=False, ad
 
     perceptron.train(data, heldout, verbose=verbose, run_label=run_label)
 
-    test_result,comb,_ = perceptron.test(test)
+    test_result, _ = perceptron.test(test)
     codecs.open('../eval/%s_test.txt' % run_label, 'w', encoding='utf-8').write(test_result)
-    
-    matrix= confMatrix(comb)
 
     if verbose:
         codecs.open('../eval/%s_features.csv' % run_label, 'w', encoding='utf-8').write('\n'.join(perceptron.evaluate_features()))

### 2012-10-29 doc/report/paper.tex
paper outline
++

@@
-\documentclass[a4paper,10pt]{article}
-%\documentclass[a4paper,10pt]{scrartcl}
+\documentclass[11pt]{article}
+\usepackage{acl2011}
+\usepackage[utf8]{inputenc}
+\usepackage{linguex}
+\usepackage{latexsym}  
+\usepackage{graphicx}
+\usepackage{multirow}
+\usepackage{verbatim}
+\usepackage{array}
+\usepackage{subfigure}
+\usepackage{hyperref}
+\usepackage{amsmath}
+% \usepackage{xcolor}
+% \definecolor{dark-red}{rgb}{0.4,0.15,0.15}
+% \definecolor{dark-blue}{rgb}{0.15,0.15,0.4}
+% \definecolor{medium-blue}{rgb}{0,0,0.5}
+% \hypersetup{
+%     colorlinks, linkcolor={dark-red},
+%     citecolor={dark-blue}, urlcolor={medium-blue}}
+\title{Named Entity Recognition using Structured Prediction}
 
-\usepackage[utf8]{inputenc}
-
-\title{Named Entity Recognition using Structured Prediction}
-\author{}
-\date{}
-
-\usepackage{hyperref}
-\usepackage{xcolor}
-\definecolor{dark-red}{rgb}{0.4,0.15,0.15}
-\definecolor{dark-blue}{rgb}{0.15,0.15,0.4}
-\definecolor{medium-blue}{rgb}{0,0,0.5}
-\hypersetup{
-    colorlinks, linkcolor={dark-red},
-    citecolor={dark-blue}, urlcolor={medium-blue}
-}
+% \author{Joachim Daiber \\
+%   %Affiliation / Address line 1 \\
+%  
+%   {\tt email@domain} \\\And
+%   Carmen Klaussner \\
+%   %Affiliation / Address line 1 \\
+% 
+%   {\tt carmen@wordsmith.de} \\
+%   }
+% 
+% \date{today}
 
@@
 
-{\small
-\bibliographystyle{plain}
-\bibliography{paper.bib}
-}
+\section{Introduction}
+Named Entity Recognition (NER) is an information extraction task and has first been introduced as part of the 6th MUC (Sixth Message Understanding Conference)
+that was focusing on the extraction of structured information from unstructured text, such as company names from newspaper articles \cite{ner-sekine2007}.
+The most prominent types are generally known under the name of \emph{enamex} types, which are comprised of \texttt{Person Names}, \texttt{Organisations} and \texttt{Locations}. 
+An additional type \texttt{Miscellaneous} captures person names outside the classic \emph{enamax} type.
+Apart from these there is \emph{timex}, which covers date \& time expressions and \emph{numex} which is for monetary values \& percent. 
+Table \ref{NETypes} gives an overview of the different NE types.
+
+\begin{table}[h!]
+\scriptsize
+\begin{tabular}{l|l}
+\hline
+\bf Named Entity Type & \bf Example \\
+\hline
+Person Names & Greta Garbo, John, Mary \\
+Organisations& Benetton, Coca-Cola Gmbh\\
+Locations&  Paris, Australia, Bristol\\
+Miscellaneous& World Cup\\
+ Date Expression& 01-02-2012, January, 6th, 1998 \\
+Time Expression & 10 p.m.\\
+Monetary Value &  \$15, 100 Euro   \\
+Percent 30\%&   \\
+\end{tabular}
+\caption{NE Examples}
+\label{table:NETypes}
+\end{table}
+
+In considering named entities, it is important to distinguish between mention of entities that are non - specific as 
+time expressions, such as \emph{In June}, referring to any possible year or \emph{the prof} as a person name , which itself is not a specific entity, but a deictic reference
+, which may point back to the mention of a real \emph{NE}, as in the following coreference chain: \\
+
+\ex. $ Prof.\; Bateman \Rightarrow he \Rightarrow the \; prof$
+
+\subsection*{Issues in Named Entity Recognition}
+Among the most common problems for NER is the issue of ambiguity, that is in particular \emph{Polysemy} \cite{ner-sekine2007}, 
+the property of some lexical representation having
+more than one possible meaning, and \emph{Metonymy}, which refers to the concept of the part-whole/whole-part relation between two expressions. 
+\emph{Polysemy} could become an issue for NER when the lexical representation of an item could point to two different \emph{NE} types.
+This is quite frequent with \texttt{Person Names} and \texttt{Locations}, since many people are named after cities, such as \emph{Paris} or \emph{Georgia}. 
+Often the context will not be disambiguating as in \ref{ex:Paris}.
+
+\ex. \emph{Paris is beautiful.} \label{ex:Paris}
+
+\emph{Metonymy} is frequently an issue in literary texts and news data (which is often used in NER), where two items that are in a part-whole
+relationship, are substituted for each other respectively. Example \ref{ex:Lon} shows an instance of \emph{whole-part}, where \emph{London} is supposedly substituted
+for the \textbf{Goverment in London}.
+
+\ex. \emph{\textbf{London} decided to increase the 1200 military personnel involved in Olympic security.} \label{ex:Lon}
+
+\subsection*{Methods employed for NER}
+For \emph{NER}, there exist both rule-based and statisical approaches. 
+Rule-based methods make use of the underlying rules governing languages to extract named entities. 
+However, this approach is quite time-consuming and requires extensive work of computational linguists \cite{ner-sekine2007}
+and although the results are 
+often high in precision, lacks considerably in recall.
+
+In regard to statistical approaches, supervised-learning is the most common method applied in \emph{NER}. Although, there are unsupervised approaches, 
+their performance is not as high as for SL applications.
+
+Prominent algorithms in NER include maximum entropy
+% Overview algorithms /methods used 
+% machine learning/neural networks/rule-based 
+% maximum entropy MM
+% 
+% Base line for NER
+
+\section{Structured Prediction}
+Structured Prediction \cite{strlearn} 
+is a supervised-learning approach and sets itself apart from Non-structured Prediction through the form of its output. 
+Prediction maps an input x to an output y: \\
+$ x \rightarrow y $.
+
+Non-structured output is atomic, thus it is binary prediction for a two-class problem and may corresponds to more than one of more than 2 possible labels for a 
+multiclass problem. 
+The output of Structured Prediction is structured and gives back a sequence/tree. 
+Thus, the output could be the complete sentence with its corresponding labels.
+
+\ex. \emph{The motor company 'Ford' was founded and incorporated  by Henry Ford on June, 16th 1903.} \label{PredEx1a}
+
+would trigger the labels: 
+\ex. x The motor company 'Ford' was founded and incorporated  by Henry Ford on June, 16th 1903. \label{PredEx1b} % include labels 
+     %y 
+Furthermore, Structured Prediction is different from similar approaches such as HMM through the combination of features and the use of label interactions.
+
+\section{Our structured Perceptron}
+Our NER structured perceptron for the languages English, German, Dutch and Spanish is trained and tested on the \emph{CoNLL 2003} and \emph{CoNLL 2002} data sets
+respectively.
+
+\paragraph*{Learning}
+Labels of the whole sentence (0) for zero entity
+\paragraph*{Structure}
+Structured Perceptron with Averaging
+
+\paragraph*{Decoding}
+Viterbi algorithm (Markov assumption, only 1 prev.
+label)
+
+\subsection*{Features}
+The features employed in the system can be divided into three categories: node, label and gazetteer features. 
+We describe each of the three groups in the following.
+
+\paragraph*{Node features}
+These are only present on the word in question: the \emph{Token}, suffix and prefix, capitalisation.
+
+Table \ref{table:node} shows the various node features with an example respectively.
+
+\begin{table}[h!]
+\begin{tabular}{l|l}
+\hline
+\bf Feature & \bf Example \\
+\hline
+Token &  \\
+Suffix& Amster\textbf{dam}\\
+Prefix&  \textbf{San} Sebastian\\
+Captitalized& \textbf{B}enetton\\
+Number Pattern & \\
+UPPERCASE &  \textbf{BENETTON}\\
+POS-tag &  Benetton \textbf{NNP}   \\
+Lemma &   \\
+\end{tabular}
+\caption{Node Features}
+\label{table:node}
+\end{table}
+
+\paragraph*{Label Interaction Features}
+These features register which label has been assigned to the previous token and takes into account the most likely sequence. 
+%current token and last label for prepositions or possessive ’s
+
+\paragraph*{Gazetteer Features}
+In order to create gazetteer lists for the more common named entities, we designed a \emph{SPARQL} query, that would retrieve entries from \emph{DBPedia} for all
+languages. The reliability of the respective list is learnt be the perceptron.
+
+\section{Experiments/ Evaluation}
+In the following section we present our experiments and the evalution of our system.
+\subsection*{Experiments}
+
+\subsection*{Evaluation}
+For the evaluation of the system we used \emph{Precision} and \emph{Recall} as shown in \ref{Precision} and \ref{Recall} respectively.
+The general formula for the \emph{F-Score} is shown in \ref{Fscore}. Since we rate both \emph{Precision} and \emph{Recall} evenly, we 
+use the harmonic mean as shown in \ref{F1}.
+
+\ex. \emph{Precision} = $ \frac{gold\; tag \bigcap predicted}{predicted}$ \label{Precision}\\
+
+\ex. \emph{Recall} = $ \frac{gold \;tag \bigcap predicted}{gold\; tag}$ \label{Recall}\\
+
+\ex. $F_{\beta}$ = $ (1+\beta^2)*\frac{precision *recall}{\beta^2* precision + recall}$ \label{Fscore}\\
+
+\ex. $F_1$ = $ 2*\frac{precision *recall}{precision + recall}$ \label{F1}\\
+
+\section{Discussion of Results}
+
+\subsection*{Some Challenges} % maybe in previous section
+
+\section{Future Work \& Conclusion}
+
+\bibliographystyle{acl}
+\bibliography{ML}
 

### 2012-11-03 doc/report/paper.tex
paper++

@@
 
+\author{Joachim Daiber \\
+  University of Groningen \\
+  %Affiliation / Address line 2 \\
+  %Affiliation / Address line 3 \\
+  {\tt jodaiber@gmx.de} \\\And
+  Carmen Klaussner \\
+  University of Groningen \\
+  %Affiliation / Address line 2 \\
+  %Affiliation / Address line 3 \\
+  {\tt Carmen@wordsmith.de} \\}
+
+\date{}
+
 \begin{document}
@@
 Named Entity Recognition (NER) is an information extraction task and has first been introduced as part of the 6th MUC (Sixth Message Understanding Conference)
-that was focusing on the extraction of structured information from unstructured text, such as company names from newspaper articles \cite{ner-sekine2007}.
+that was focusing on the extraction of structured information from unstructured text, such as company names from newspaper articles \cite{nadeau2007survey}.
 The most prominent types are generally known under the name of \emph{enamex} types, which are comprised of \texttt{Person Names}, \texttt{Organisations} and \texttt{Locations}. 
@@
 Apart from these there is \emph{timex}, which covers date \& time expressions and \emph{numex} which is for monetary values \& percent. 
-Table \ref{NETypes} gives an overview of the different NE types.
+Table \ref{table:NETypes} gives an overview of the different NE types.
 
@@
 Monetary Value &  \$15, 100 Euro   \\
-Percent 30\%&   \\
+Percent &   30\%
 \end{tabular}
@@
 \subsection*{Issues in Named Entity Recognition}
-Among the most common problems for NER is the issue of ambiguity, that is in particular \emph{Polysemy} \cite{ner-sekine2007}, 
+Among the most common problems for NER is the issue of ambiguity, that is in particular \emph{Polysemy} \cite{nadeau2007survey}, 
 the property of some lexical representation having
@@
 Rule-based methods make use of the underlying rules governing languages to extract named entities. 
-However, this approach is quite time-consuming and requires extensive work of computational linguists \cite{ner-sekine2007}
+However, this approach is high maintenance, quite time-consuming and requires extensive work of computational linguists \cite{nadeau2007survey}
 and although the results are 
@@
 In regard to statistical approaches, supervised-learning is the most common method applied in \emph{NER}. Although, there are unsupervised approaches, 
-their performance is not as high as for SL applications.
+their performance is not as high as for SL applications yet.
 
@@
 %
-% Base line for NER
+
+\subsection*{CoNLL data set \& baseline}
+The training and test dataset was taken from the \emph{CoNLL Shared Task 2002} \cite{TjongKimSang:2003:ICS:1119176.1119195} and the \emph{CoNLL Shared Task 2003} \cite{tksintro2002conll}
+for Spanish \& Dutch and English \& German respectively.
+
+% A baseline rate was computed for the English and the
+% German test sets. It was produced by a system which
+% only identified entities which had a unique class in
+% the training data. If a phrase was part of more than
+% one entity, the system would select the longest one.
+% All systems that participated in the shared task have
+% outperformed the baseline system.
+% For all the Fβ=1 rates we have estimated sig-
+% nificance boundaries by using bootstrap resampling
+% (Noreen, 1989). From each output file of a system,
+% 250 random samples of sentences have been chosen
+% and the distribution of the Fβ=1 rates in these sam-
+% ples is assumed to be the distribution of the perfor-
+% mance of the system. We assume that performance
+% A is significantly different from performance B if A
+% is not within the center 90% of the distribution of B.
+
+The baseline for \emph{Named Entity Recognition} in the four different languages is visualised in table \
+\begin{table}[h!]
+\scriptsize
+\begin{tabular}{l|l|l|l}
+\hline
+\bf Language & \bf Precision & \bf Recall & \bf $F_1$-measure \\ \hline
+Spanish &             26.27 \% & 56.48 \% & 35.86  \%        \\
+Dutch  &             64.38 \%  &45.19   \%    & 53.10 \%  \\
+English &              71.91\%& 50.90 \%  & 59.61 $\pm$ 1.2 \%\\
+German &      31.86 \%  & 28.89 \% & 30.30  $\pm$ \%1.3 \\
+\end{tabular}
+\caption{NER Baseline}
+\label{table:Base}
+\end{table}
 
@@
 
+\begin{table}[h!]
+\begin{tabular}{l |l| l| l| l| l| l}
+
+\bf Language & \multicolumn{3}{c|}{ \bf testA}&\multicolumn{3}{c}{ \bf testB}\\\hline 
+             & Precision & Recall & $F_1$-measure & Precision & Recall & $F_1$-measure \\ \hline
+Spanish &       &          &     &          &               & \\
+Dutch  &         &          &     &          &               &   \\
+English &        &          &     &          &               &       \\
+German &      &          &       &          &             & \\
+\end{tabular}
+\caption{NER Structured Prediction Results }
+\label{table:Results}
+\end{table}
+
 \section{Discussion of Results}
@@
 \bibliographystyle{acl}
-\bibliography{ML}
+\bibliography{paper}
 

### 2012-11-03 doc/report/paper.tex
Trees and so on

@@
 \usepackage{amsmath}
+\usepackage{qtree}
+
 % \usepackage{xcolor}
@@
 
+\DeclareMathOperator*{\argmax}{arg\,max}
+
 \begin{document}
@@
 
-In considering named entities, it is important to distinguish between mention of entities that are non - specific as 
-time expressions, such as \emph{In June}, referring to any possible year or \emph{the prof} as a person name , which itself is not a specific entity, but a deictic reference
-, which may point back to the mention of a real \emph{NE}, as in the following coreference chain: \\
+In considering named entities, it is important to distinguish between mention of entities that are non-specific as 
+time expressions, such as \emph{In June}, referring to any possible year or \emph{the prof} as a person name, which itself is not a specific entity, 
+but a deictic reference, which may point back to the mention of a real \emph{NE}, as in the following coreference chain: \\
 
-\ex. $ Prof.\; Bateman \Rightarrow he \Rightarrow the \; prof$
+\ex. $ Prof.\; Bateman \Rightarrow he \Rightarrow the \; prof$ \\
+
+Named Entities can consist of more than one syntactic type. The usual overall type for a named entity is noun phrase, which can be both simple and complex. 
+An example for a simple heterogenous entity is displayed in syntactic tree represenation in ~\ref{tree1}.
+ \Tree
+    [.NP [.N' John ]  ] \label{tree1}
+
+Most languages allow for a recursive definition of, for instance, a noun phrase that in theory can grow infinitely big. 
+An \textbf{NP} can consist of a determiner, such as \emph{the} and a common noun or simply a propernoun. \textbf{PP} stands
+for prepositional phrase and consists of a preposition and a noun phrase. 
+The syntactic rules for PP-attachment are displayed in \ref{PP}.
+\begin{figure}
+$ NP \rightarrow NP PP$ \\
+$ PP \rightarrow P NP$ 
+\caption{Recursive Noun Phrase}
+\label{PP}
+\end{figure}   
+These rules allow for constructions such as in sentence~\ref{PPrec}.
+
+\ex. The princess of Scotland is in the castle in England with the brown bricks on its roof... \label{PPrec}
+
+We distinguish between PP-complements and PP-modifiers, where modifiers simply add to the description of the noun phrase, maybe
+to set it apart and complements are more inherent to the meaning of the entity. 
+The PP-attachment \emph{of Scotland} is semantically defining for the noun \emph{princess}, whereas \emph{with the brown bricks on its roof} 
+just adds to the description of \emph{the castle in England}.
+Regarding named entities, one wants to differentiate between a PP-attachment that defines the entity an one that is irrelevant.
+
+\begin{enumerate}
+\item \parbox[t]{2.4in}{ (a)~\Tree 
+   [.NP [ [.DT the ] [.N\1 [.N\1 Princess ] [.PP [.P of ] [.NP England ] ] ] ] ]}
+\parbox[t]{2in}{ (b)~\Tree 
+   [.NP  [.DT the ] [.N\1 [.N\1 Princess ] ] [.PP [.P from ] [.NP England ] ]  ] ]  }
+\end{enumerate}
 
@@
 However, this approach is high maintenance, quite time-consuming and requires extensive work of computational linguists \cite{nadeau2007survey}
-and although the results are 
-often high in precision, lacks considerably in recall.
+and although the results are often high in precision, lacks considerably in recall.
 
@@
 
-Prominent algorithms in NER include maximum entropy
-% Overview algorithms /methods used 
-% machine learning/neural networks/rule-based 
-% maximum entropy MM
-%
+Prominent algorithms in NER include maximum entropy models,
 
-\subsection*{CoNLL data set \& baseline}
-The training and test dataset was taken from the \emph{CoNLL Shared Task 2002} \cite{TjongKimSang:2003:ICS:1119176.1119195} and the \emph{CoNLL Shared Task 2003} \cite{tksintro2002conll}
-for Spanish \& Dutch and English \& German respectively.
+neural network
 
+\subsection*{CoNLL Data Set \& Baseline Performance}
+The training and test dataset was taken from the \emph{CoNLL Shared Task 2002} \cite{tksintro2002conll} and the \emph{CoNLL Shared Task 2003} \cite{TjongKimSang:2003:ICS:1119176.1119195}.
+for Spanish \& Dutch and English \& German respectively. The ConNLL shared tasks are organised challenges, where the organiser propose a task and provide the
+training and test data. For 2002/2003 it was Named Entity Recognition. Also we base our system on the approach that won the challenge for English and German in 2003. 
+The baseline rate for all languages was produced by a system, that had a unique class in the training data. In case that a phrase was part of more than one entity, 
+the system would choose the longest one \cite{TjongKimSang:2003:ICS:1119176.1119195}. 
+The data is annotated with POS-tags and NE-tags, as well as Bio-tags that indicate the position in the Named Entity. 
 % A baseline rate was computed for the English and the
@@
 \section{Structured Prediction}
-Structured Prediction \cite{strlearn} 
-is a supervised-learning approach and sets itself apart from Non-structured Prediction through the form of its output. 
-Prediction maps an input x to an output y: \\
-$ x \rightarrow y $.
+Structured Prediction \cite{strlearn} is a supervised-learning approach that maps an input \textbf{x} to an output \textbf{y}: \\
 
-Non-structured output is atomic, thus it is binary prediction for a two-class problem and may corresponds to more than one of more than 2 possible labels for a 
+$ x \rightarrow y $. \\
+
+Structured and  Non-structured prediction differ in the form of their output. 
+Non-structured output is atomic; it is binary prediction for a two-class problem and may corresponds to more than one of more than two possible labels for a 
 multiclass problem. 
-The output of Structured Prediction is structured and gives back a sequence/tree. 
-Thus, the output could be the complete sentence with its corresponding labels.
+Structured Prediction gives back the prediction for the whole sequence. 
+For the sentence in \ref{PredEx1a}, it would return the corresponding labels combination in example~\ref{PredEx1b}.
 
-\ex. \emph{The motor company 'Ford' was founded and incorporated  by Henry Ford on June, 16th 1903.} \label{PredEx1a}
+\ex. \emph{The motor company 'Ford' was founded and incorporated  by Henry Ford.} \label{PredEx1a}
+%The motor company 'Ford' was founded and incorporated  by Henry Ford on June, 16th 1903.
 
-would trigger the labels: 
-\ex. x The motor company 'Ford' was founded and incorporated  by Henry Ford on June, 16th 1903. \label{PredEx1b} % include labels 
-     %y 
-Furthermore, Structured Prediction is different from similar approaches such as HMM through the combination of features and the use of label interactions.
+\ex. \textbf{x:} The motor company 'Ford' was founded and incorporated by Henry Ford.\\ \textbf{y:} 0  0  ORG  0  0 0 0 0 PER PER.  \label{PredEx1b}
 
-\section{Our structured Perceptron}
-Our NER structured perceptron for the languages English, German, Dutch and Spanish is trained and tested on the \emph{CoNLL 2003} and \emph{CoNLL 2002} data sets
+Also, Structured Prediction is different from similar approaches through its combination of features and label interactions instead
+of concentrating mainly on label interactions like HMM or a rich set of features like local classifiers.
+
+\section{Our Implementation}
+% problem external data / use Dbpedia - gazetteer
+Our NER system for the languages English, German, Dutch and Spanish is trained and tested on the \emph{CoNLL 2003} and \emph{CoNLL 2002} data sets
 respectively.
@@
 Labels of the whole sentence (0) for zero entity
+
 \paragraph*{Structure}
@@
 \paragraph*{Decoding}
-Viterbi algorithm (Markov assumption, only 1 prev.
-label)
+In this system we are using viterbi decoding, thus we are looking for the prediction $\hat{y}$, which is the sequence with the highest overall score:
+
+\[
+\hat{y} = \argmax_{\mathbf{y} \in \mathcal{Y}^{n}} \mathbf{w} \cdot \boldsymbol{f}(\mathbf{x}, i, y_{i-1}, y_{i})
+\]
+
+This sequence can be computed efficiently in $ O( N^2 |\mathbf{x}| ) $ using the Viterbi algorithm. 
+As a first step, the trellis has to be constructed, then we have to find the $ a \in \mathcal{Y}$ in the 
+last column of the trellis with maximal score $\delta_n(a)$. From this, the sequence can be recovered via back-tracking trough the trellis. 
+The score of the best sequence ending in $a$ is:
+
+\[
+\delta_i(a) = \max_{\mathbf{y} \in \mathcal{Y}^{n}, y_n = a} \sum_{j=1}^{n}{\mathbf{w} \cdot \boldsymbol{f}(\mathbf{x}, j, y_{j-1}, y_{j})}
+\]
+
+\noindent And $\delta_i(a)$ can be calculated recursively:
+
+\begin{align*}
+\delta_1(a) &= \mathbf{w} \cdot \boldsymbol{f}(\mathbf{x}, 1, \emptyset, a) \\
+\delta_i(a) &= \max_{b \in \mathcal{Y}} \delta_{i-1}(b) + \mathbf{w} \cdot \boldsymbol{f}(\mathbf{x}, i, b, a) \\
+\end{align*}
 
@@
 \hline
-Token &  \\
+Token &  \textbf{``Amsterdam''}\\
 Suffix& Amster\textbf{dam}\\
@@
 POS-tag &  Benetton \textbf{NNP}   \\
-Lemma &   \\
+Lemma &  produced $Rightarrow$ produce \\
 \end{tabular}
@@
 These features register which label has been assigned to the previous token and takes into account the most likely sequence. 
+Thus, for the sequence: ``Jack London went to New York'' the NE tag combination of \emph{PER} and \emph{PER} in example~\ref{seq1}is more likely 
+than\emph{PER} and \emph{LOC} as presented in \ref{seq2}.
+.
 %current token and last label for prepositions or possessive ’s
+
+\ex. Jack London went to New York. \\ PER   PER  0  0  LOC   LOC. \label{seq1}
+
+\ex. Jack London went to New York. \\ PER   LOC  0  0  LOC   LOC. \label{seq2}
 
@@
 \begin{table}[h!]
+\scriptsize
 \begin{tabular}{l |l| l| l| l| l| l}
@@
 \bf Language & \multicolumn{3}{c|}{ \bf testA}&\multicolumn{3}{c}{ \bf testB}\\\hline 
-             & Precision & Recall & $F_1$-measure & Precision & Recall & $F_1$-measure \\ \hline
+             & Precision & Recall & $F_1$ & Precision & Recall & $F_1$ \\ \hline
 Spanish &       &          &     &          &               & \\

### 2012-11-06 doc/report/paper.tex
Report++.

@@
 
+\newcommand{\namedentity}{Named Entity} 
+\newcommand{\Oo}{\texttt O}
+
 \begin{document}
@@
 \scriptsize
-\begin{tabular}{l|l}
-\hline
+\begin{tabular}{l l}
 \bf Named Entity Type & \bf Example \\
@@
 Monetary Value &  \$15, 100 Euro   \\
-Percent &   30\%
+Percent &   30\% \\
 \end{tabular}
@@
 
-\subsection*{CoNLL Data Set \& Baseline Performance}
+\subsection*{CoNLL Data and Baseline Performance}
 The training and test dataset was taken from the \emph{CoNLL Shared Task 2002} \cite{tksintro2002conll} and the \emph{CoNLL Shared Task 2003} \cite{TjongKimSang:2003:ICS:1119176.1119195}.
@@
 \scriptsize
-\begin{tabular}{l|l|l|l}
+\begin{tabular}{|l|l|l|l|}
 \hline
 \bf Language & \bf Precision & \bf Recall & \bf $F_1$-measure \\ \hline
-Spanish &             26.27 \% & 56.48 \% & 35.86  \%        \\
-Dutch  &             64.38 \%  &45.19   \%    & 53.10 \%  \\
-English &              71.91\%& 50.90 \%  & 59.61 $\pm$ 1.2 \%\\
-German &      31.86 \%  & 28.89 \% & 30.30  $\pm$ \%1.3 \\
+Spanish &             26.27\% & 56.48\% & 35.86\%        \\
+Dutch  &             64.38\%  &45.19\%    & 53.10\%  \\
+English &              71.91\%& 50.90\%  & 59.61 $\pm$ 1.2\%\\
+German &      31.86\%  & 28.89\% & 30.30  $\pm$ 1.3\% \\
+\hline
 \end{tabular}
@@
 
-\ex. \textbf{x:} The motor company 'Ford' was founded and incorporated by Henry Ford.\\ \textbf{y:} 0  0  ORG  0  0 0 0 0 PER PER.  \label{PredEx1b}
+\ex. \textbf{x:} The motor company 'Ford' was founded and incorporated by Henry Ford.\\ \textbf{y:} O  O  ORG  O  O O O O PER PER.  \label{PredEx1b}
 
@@
 
-\section{Our Implementation}
+\section{Implementation}
+
+\subsection{Learning and Decoding}
+
 % problem external data / use Dbpedia - gazetteer
-Our NER system for the languages English, German, Dutch and Spanish is trained and tested on the \emph{CoNLL 2003} and \emph{CoNLL 2002} data sets
-respectively.
+Our implementation for the languages English, German, Dutch and Spanish is trained and tested on the \emph{CoNLL 2003} and \emph{CoNLL 2002} data sets respectively.
 
-\paragraph*{Learning}
-Labels of the whole sentence (0) for zero entity
+As the predicted structure, we use the labels for every token in the full sentence. If a token is not a \namedentity, it is assigned the label \Oo.
 
-\paragraph*{Structure}
-Structured Perceptron with Averaging
-
-\paragraph*{Decoding}
-In this system we are using viterbi decoding, thus we are looking for the prediction $\hat{y}$, which is the sequence with the highest overall score:
+Learning is implemented with a Structured Perceptron.
+To avoid overfitting, we average the parameters after the last iteration.
+For finding the best sequence of labels, we use a modified version of the Viterbi algorithm.
+The algorithm finds the sequence $\hat{y}$, such that:
 
 \[
-\hat{y} = \argmax_{\mathbf{y} \in \mathcal{Y}^{n}} \mathbf{w} \cdot \boldsymbol{f}(\mathbf{x}, i, y_{i-1}, y_{i})
+\hat{y} = \argmax_{\mathbf{y} \in \mathcal{Y}^{n}} \sum_{i=1}^{n}\mathbf{w} \cdot \boldsymbol{f}(\mathbf{x}, i, y_{i-1}, y_{i})
 \]
 
-This sequence can be computed efficiently in $ O( N^2 |\mathbf{x}| ) $ using the Viterbi algorithm. 
+This sequence can be computed efficiently in $ O( N^2 |\mathbf{x}| ) $ via dynamic programming. 
 As a first step, the trellis has to be constructed, then we have to find the $ a \in \mathcal{Y}$ in the 
@@
 
-\noindent And $\delta_i(a)$ can be calculated recursively:
+\noindent This function $\delta_i(a)$ can be defined recursively as:
 
@@
 
-\subsection*{Features}
+\subsection{Features}
 The features employed in the system can be divided into three categories: node, label and gazetteer features. 

### 2012-11-06 doc/report/paper.tex
Example layouts, some writing.

@@
 \usepackage{qtree}
+%\usepackage{gb4e}
+
+%\usepackage{gb4e}
 
@@
 %     citecolor={dark-blue}, urlcolor={medium-blue}}
-\title{Named Entity Recognition using Structured Prediction}
+\title{Named Entity Recognition with Structured Prediction}
 
@@
 The most prominent types are generally known under the name of \emph{enamex} types, which are comprised of \texttt{Person Names}, \texttt{Organisations} and \texttt{Locations}. 
-An additional type \texttt{Miscellaneous} captures person names outside the classic \emph{enamax} type.
+An additional type \texttt{Miscellaneous} captures names outside the classic \emph{enamax} type.
 Apart from these there is \emph{timex}, which covers date \& time expressions and \emph{numex} which is for monetary values \& percent. 
@@
 \scriptsize
-\begin{tabular}{l l}
+\begin{tabular}{| l | l |}
+\hline
 \bf Named Entity Type & \bf Example \\
@@
 Percent &   30\% \\
+\hline
 \end{tabular}
@@
 
-\ex. \textbf{x:} The motor company 'Ford' was founded and incorporated by Henry Ford.\\ \textbf{y:} O  O  ORG  O  O O O O PER PER.  \label{PredEx1b}
+\exg. \textbf{x:} The motor company ' Ford ' was founded and incorporated by Henry Ford .\\
+      \textbf{y:} O O O O ORG  O  O O O O O PER PER O  \label{PredEx1b} \\
 
@@
 
-Learning is implemented with a Structured Perceptron.
-To avoid overfitting, we average the parameters after the last iteration.
+Learning is implemented using the Structured Perceptron algorithm.
+To avoid overfitting, we average the parameters after the last iteration \cite{collins2002discriminative}.
 For finding the best sequence of labels, we use a modified version of the Viterbi algorithm.
@@
 \paragraph*{Node features}
-These are only present on the word in question: the \emph{Token}, suffix and prefix, capitalisation.
+Node features are features that only depend on the current token (node): the \emph{Token}, suffix and prefix, capitalisation.
 
@@
 \begin{table}[h!]
-\begin{tabular}{l|l}
+\begin{tabular}{| l | l |}
 \hline
@@
 POS-tag &  Benetton \textbf{NNP}   \\
-Lemma &  produced $Rightarrow$ produce \\
+Lemma &  produced $\Rightarrow$ produce \\
+\hline
 \end{tabular}
@@
 
-\ex. Jack London went to New York. \\ PER   PER  0  0  LOC   LOC. \label{seq1}
+\exg. Jack London went to New York .\\
+      PER   PER   O    O  LOC LOC  O \\\label{seq1}
 
-\ex. Jack London went to New York. \\ PER   LOC  0  0  LOC   LOC. \label{seq2}
+\exg. Jack London went to New York . \\ 
+      PER  LOC    O    O  LOC LOC O \\\label{seq2}
 

### 2012-11-07 doc/report/paper.tex
Minor paper changes.

@@
 \usepackage{qtree}
+
 %\usepackage{gb4e}
@@
 
+\begin{figure*}[ht]
+
 \ex. \emph{The motor company 'Ford' was founded and incorporated  by Henry Ford.} \label{PredEx1a}
@@
       \textbf{y:} O O O O ORG  O  O O O O O PER PER O  \label{PredEx1b} \\
+\caption{Input and predicted structure for the Named Entity task.}
+
+\end{figure*}
 
@@
 To avoid overfitting, we average the parameters after the last iteration \cite{collins2002discriminative}.
-For finding the best sequence of labels, we use a modified version of the Viterbi algorithm.
+For finding the best sequence of labels for a sentence, we use a modified version of the Viterbi algorithm.
 The algorithm finds the sequence $\hat{y}$, such that:
@@
 
-Table \ref{table:node} shows the various node features with an example respectively.
+Table~\ref{table:node} shows the various node features with an example respectively.
 
@@
 
-\section{Experiments/ Evaluation}
+\section{Experiments and Evaluation}
 In the following section we present our experiments and the evalution of our system.
@@
 \scriptsize
-\begin{tabular}{l |l| l| l| l| l| l}
+\begin{tabular}{| l | l l l| l l l |}
 
-\bf Language & \multicolumn{3}{c|}{ \bf testA}&\multicolumn{3}{c}{ \bf testB}\\\hline 
+\hline
+\bf Language & \multicolumn{3}{c|}{ \bf Test A}&\multicolumn{3}{c|}{ \bf Test B}\\
              & Precision & Recall & $F_1$ & Precision & Recall & $F_1$ \\ \hline
@@
 German &      &          &       &          &             & \\
+\hline
 \end{tabular}

### 2012-11-08 doc/report/paper.tex
Some changes, abstract placeholder.

@@
 %     citecolor={dark-blue}, urlcolor={medium-blue}}
-\title{Named Entity Recognition with Structured Prediction}
+\title{Named Entity Recognition as Structured Prediction}
 
@@
   University of Groningen \\
-  %Affiliation / Address line 2 \\
+  2397331\\
   %Affiliation / Address line 3 \\
@@
   University of Groningen \\
-  %Affiliation / Address line 2 \\
+  2401541\\
   %Affiliation / Address line 3 \\
@@
 
-\section{Introduction}
+\begin{abstract}
+Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+\end{abstract}
+
+\section{Application: Named Entity Recognition}
+
 Named Entity Recognition (NER) is an information extraction task and has first been introduced as part of the 6th MUC (Sixth Message Understanding Conference)
@@
 
-\subsection*{Issues in Named Entity Recognition}
+\subsection{Issues in Named Entity Recognition}
 Among the most common problems for NER is the issue of ambiguity, that is in particular \emph{Polysemy} \cite{nadeau2007survey}, 
@@
 
-\subsection*{Methods employed for NER}
+\subsection{Methods employed for NER}
 For \emph{NER}, there exist both rule-based and statisical approaches. 
@@
 
-\subsection*{CoNLL Data and Baseline Performance}
+\subsection{CoNLL Data and Baseline Performance}
 The training and test dataset was taken from the \emph{CoNLL Shared Task 2002} \cite{tksintro2002conll} and the \emph{CoNLL Shared Task 2003} \cite{TjongKimSang:2003:ICS:1119176.1119195}.

### 2012-11-08 src/structured_ner/demo.py
Tiding up, parallel processing.

@@ -12,10 +12,6 @@ from case.MosesTrueCaser import MosesTrueCaser
 @route('/')
 def index():
     return """
-
-
-
-
 <!DOCTYPE html PUBLIC "-//W3C//DTD XHTML 1.0 Strict//EN"
     "http://www.w3.org/TR/xhtml1/DTD/xhtml1-strict.dtd">
 
@@ -190,9 +186,7 @@ def ner(sent):
         f = ""
         for ft in features:
             f += ft.replace("<", "-").replace(">", "-") + "<br />"
-
         s += "<span class='ne_label' rel='tooltip' title='%s'>%s</span>" % (f, label)
-
         s += "</span>"
 
     return s

### 2012-11-08 src/structured_ner/evaluate.py

@@ -8,60 +8,79 @@ from nltk.stem.wordnet import WordNetLemmatizer
 from features.features_labels import LabelInteractionFeatures
 from features.features_node import SimpleNodeFeatures
 from features.features_gazetteer import GazetteerFeatures
-from case.MosesTrueCaser import MosesTrueCaser
+from case.moses_truecaser import MosesTrueCaser
 from train_supervised import train_ner
 
-conll2002_labels = ['LOC', 'PER', 'ORG', 'MISC', 'O']
+import sys
+sys.modules['cPickle']=pickle
+
+import multiprocessing
+
+
 conll2003 = LazyCorpusLoader('conll2003', ConllChunkCorpusReader, '.*\.(test|train).*', ('LOC', 'PER', 'ORG', 'MISC'), encoding='utf-8')
 
 ########################################
 # Feature sets and Gazetteers for Dutch
 ########################################
 
-lemmatizer = SnowballStemmer("dutch")
-truecaser  = MosesTrueCaser(open("models/truecase/nl"))
-nl_gazetteer = GazetteerFeatures(['data/loc_nl.txt', 'data/person_nl.txt', 'data/org_nl.txt'], truecaser)
-
-dutch_train     = load_conll(conll2002.chunked_sents('ned.train'), lemmatizer, None, nl_gazetteer )
-dutch_heldout   = load_conll(conll2002.chunked_sents('ned.testa'), lemmatizer, truecaser, nl_gazetteer)
-dutch_test      = load_conll(conll2002.chunked_sents('ned.testb'), lemmatizer, truecaser, nl_gazetteer)
-
-dutch_ner        = train_ner('nl', conll2002_labels, dutch_train, dutch_heldout, dutch_test, [SimpleNodeFeatures(), LabelInteractionFeatures(), nl_gazetteer], verbose=True)
-pickle.dump(dutch_ner, open("models/%s.pickle" % 'nl', 'w'))
-
-dutch_ner2       = train_ner('nl', conll2002_labels, dutch_train, dutch_heldout, dutch_test, [SimpleNodeFeatures(), LabelInteractionFeatures()], verbose=True)
-dutch_ner1       = train_ner('nl', conll2002_labels, dutch_train, dutch_heldout, dutch_test, [SimpleNodeFeatures()], verbose=True)
-
[…]

### 2012-11-08 src/structured_ner/features/features_labels.py

@@ -1,6 +1,7 @@
 from feature_generator import FeatureSet
 
 class LabelInteractionFeatures(FeatureSet):
+    """Set of features for interaction between NE labels."""
 
     def generate_features(self, feature_storage, sequence, i, last_label, label):
 
@@ -10,10 +11,20 @@ class LabelInteractionFeatures(FeatureSet):
         feature_id = feature_storage.add_feature("label_bigram:%s-%s" % (last_label, label))
         features_fired.append(feature_id)
 
+        #" as part of the entity:
+        if sequence.x[i][0] == '"':
+            feature_id = feature_storage.add_feature("''_and_last_POS:%s-%s" % (sequence.x[i][1], label))
+            features_fired.append(feature_id)
+
+
         #Last label and current POS tag:
-        if sequence.x[i][1] in ['IN', ')', '(', 'PRP', 'POS']:
-            feature_id = feature_storage.add_feature("labels_and_pos_tag:%s-%s-%s" % (last_label, sequence.x[i][0], label))
+        def attachment(j):
+            return sequence.x[j][0] in ['&'] or sequence.x[j][1] in ['IN', ')', '(', 'PRP', 'POS', 'Conj', 'Prep', '&', 'Num', 'Fe', 'Fz', 'SP']
+
+        if attachment(i):
+            feature_id = feature_storage.add_feature("attachment_problem:%s-%s-%s-%s" % (last_label, sequence.x[i][0], str(attachment(i-1)), label))
             features_fired.append(feature_id)
 
         return features_fired
+
 

### 2012-11-08 src/structured_ner/gazetteer/dbpedia_gazetteer.py

@@ -4,23 +4,25 @@ import sys
 
 class GazetteerExtractor():
 
+    """
+    A GazetteerExtractor automatically retrieves lists of names from a public LOD endpoint.
+    Currently the lists are for the classes Person, Organization, Location and SportsEvent.
+    The list can easily be extended, since the mapping to labels is done automatically by
+    the Perceptron learning algorithm.
+    """
+
+    dbp_classes = ['Person', 'Organisation', 'Place', 'SportsEvent']
+
     def __init__(self, lang):
 
-        persons          = self.getGazetteerEntries(lang, 'Person')
-        print >>sys.stderr, "Retrieved %d person names." % len(persons)
-        codecs.open("data/person_%s.txt" % lang, encoding='utf-8', mode='w').write('\n'.join(persons))
+        for dbp_class in GazetteerExtractor.dbp_classes:
+            list = self.retrieve_name_list(lang, dbp_class)
+            print >>sys.stderr, "Retrieved %d names: %s" % (len(list), dbp_class)
 
-        organizations    = self.getGazetteerEntries(lang, 'Organisation')
-        print >>sys.stderr, "Retrieved %d organization names." % len(organizations)
-        codecs.open("data/org_%s.txt" % lang, encoding='utf-8', mode='w').write('\n'.join(organizations))
-
-        locations        = self.getGazetteerEntries(lang, 'Place')
-        print >>sys.stderr, "Retrieved %d location names." % len(locations)
-        codecs.open("data/loc_%s.txt" % lang, encoding='utf-8', mode='w').write('\n'.join(locations))
+            codecs.open("../data/%s_%s.txt" % (dbp_class, lang), encoding='utf-8', mode='w').write('\n'.join(list))
 
 
-
-    def getGazetteerEntries(self, lang, dbp_class):
+    def retrieve_name_list(self, lang, dbp_class):
 
         repeat  = True
         entries = []
@@ -56,9 +58,9 @@ class GazetteerExtractor():
 
         return entries
 
-
-GazetteerExtractor('en')
-GazetteerExtractor('nl')
-GazetteerExtractor('es')
-GazetteerExtractor('de')
+if __name__ == '__main__':
+    GazetteerExtractor('en')
+    GazetteerExtractor('nl')
+    GazetteerExtractor('es')
+    GazetteerExtractor('de')
 

### 2012-11-08 src/structured_ner/gazetteer/wiktionary_category_contents.sh

@@ -1,3 +1,5 @@
 #! /bin/sh
 
+# Retrieves all words in a specific category of Wiktionary
+
 curl -H "Accept: text/xml" "http://en.wiktionary.org/w/api.php?action=query&list=categorymembers&cmtitle=Category:$1&cmlimit=500&format=xml" | egrep -o '<cm pageid="[0-9]+" ns="[0-9]+" title="[^"]+"' | sed 's/.* title=\"\([^"]*\)\"$/\1/g' | grep -v ^Category: > $1

### 2012-11-08 src/structured_ner/linear_classifier.py

@@ -39,6 +39,7 @@ class LinearClassifier(object):
         incorrect = 0
 
         errors = []
+        true_and_predicted = []
 
         gold_count_for_label      = dict([ (l, 0) for l in self.labels ])
         predicted_count_for_label = dict([ (l, 0) for l in self.labels ])
@@ -50,6 +51,8 @@ class LinearClassifier(object):
             for i in range(len(z)):
                 if z[i] != sentence.y[i]:
                     errors.append( sentence.x[i][0] + ' ' + sentence.x[i][1] + ' ' + sentence.true_case[i] + ' ' + sentence.y[i] + ' ' + z[i] )
+
+                true_and_predicted.append([sentence.x[i][0], sentence.y[i], z[i]])
 
             total += len(sentence.y)
             incorrect += len( [i for i in range(len(sentence.y)) if sentence.y[i] != z[i]])
@@ -93,7 +96,7 @@ class LinearClassifier(object):
         out += "Errors:\n"
         out += '\n'.join(errors)
 
-        return out, 1.0 - (incorrect/total)
+        return out, true_and_predicted, 1.0 - (incorrect/total)
 
 
     def evaluate_features(self):

### 2012-11-08 src/structured_ner/sentence.py

@@ -1,4 +1,5 @@
 from __future__ import division
+from nltk.stem.api import StemmerI
 
 def all_uppercase(tokens, tags):
     return sum([ (tokens[i].isupper() or (tags[i] in ['CD', '.', ',', ':'])) for i in range(len(tokens))]) / float(len(tokens)) >= 0.95
@@ -34,7 +35,10 @@ def sentence_from_conll(chunked_sent, lemmatizer, truecaser, gazetteer):
     else:
         gazetteer_entries = map(lambda _: 'O', true_case)
 
-    lemmas = map(lambda tc: lemmatizer.lemmatize(tc), true_case)
+    if isinstance(lemmatizer, StemmerI):
+        lemmas = map(lambda tc: lemmatizer.stem(tc), true_case)
+    else:
+        lemmas = map(lambda tc: lemmatizer.lemmatize(tc), true_case)
 
     return Sentence(x, y, lemmas, true_case, gazetteer_entries)
 

### 2012-11-08 src/structured_ner/structured_perceptron.py

@@ -23,7 +23,7 @@ class StructuredPerceptron(LinearClassifier):
         self.w = np.zeros(self.n_features, dtype=float)
 
         training_accuracy = [0.0]
-        heldout_accuracy = [0.0]
+        heldout_accuracy  = [0.0]
 
         for i_epoch in xrange(self.n_epochs):
 
@@ -39,7 +39,7 @@ class StructuredPerceptron(LinearClassifier):
             training_accuracy.append(accuracy)
 
             if verbose:
-                _, acc = self.test(heldout)
+                _, _, acc = self.test(heldout)
                 heldout_accuracy.append(acc)
 
             #Stop if the error on the training data does not decrease

### 2012-11-08 src/structured_ner/util.py

@@ -1,24 +1,0 @@
-def entities_from_list(list):
-    b = False
-    current_start = 0
-    current_tag = None
-
-    entities = []
-
-    for i in range(len(list)):
-      if list[i] == 'O' or i == len(list)-1 or (b and list[i] != current_tag):
-          if b:
-              entities.append( (current_start, i-1, current_tag) )
-              b = False
-              current_start = 0
-              current_tag = None
-
-      if list[i] != 'O' and not b:
-          b = True
-          current_start = i
-          current_tag = list[i]
-
-      if i == len(list)-1 and list[i] != 'O':
-          entities.append( (i, i, list[i]) )
-
-    return entities

### 2012-11-08 doc/report/paper.tex

@@
 \usepackage{qtree}
+\usepackage{textcomp}
 
@@
 that was focusing on the extraction of structured information from unstructured text, such as company names from newspaper articles \cite{nadeau2007survey}.
-The most prominent types are generally known under the name of \emph{enamex} types, which are comprised of \texttt{Person Names}, \texttt{Organisations} and \texttt{Locations}. 
+The most prominent types are generally known under the name of \emph{enamex} types, which are comprised of \texttt{Person Names}, 
+\texttt{Organisations} and \texttt{Locations}. 
 An additional type \texttt{Miscellaneous} captures names outside the classic \emph{enamax} type.
-Apart from these there is \emph{timex}, which covers date \& time expressions and \emph{numex} which is for monetary values \& percent. 
+Apart from the above, there is the \emph{timex} type, which covers date \& time expressions and \emph{numex} which is for monetary values \& percent. 
 Table \ref{table:NETypes} gives an overview of the different NE types.
@@
 Locations&  Paris, Australia, Bristol\\
-Miscellaneous& World Cup\\
+Miscellaneous& World Cup 2014\\
  Date Expression& 01-02-2012, January, 6th, 1998 \\
 Time Expression & 10 p.m.\\
-Monetary Value &  \$15, 100 Euro   \\
+Monetary Value &  \$15, \textsterling 100    \\
 Percent &   30\% \\
@@
 
-In considering named entities, it is important to distinguish between mention of entities that are non-specific as 
-time expressions, such as \emph{In June}, referring to any possible year or \emph{the prof} as a person name, which itself is not a specific entity, 
-but a deictic reference, which may point back to the mention of a real \emph{NE}, as in the following coreference chain: \\
+In considering named entities, it is important to distinguish between named entities and the mention of entities that are in fact non-specific. % reformulate
+The difference lies in the reference of the entity. A named entity refers to a specific, unique person/time/location, whereas in 
+mere mentions of entites, these can refer to multiple events, such as the time expression: \emph{In June}, referring to any possible year.
+Also, there could be a reference to a person, as in: \emph{the prof}, which in itself is not a specific entity.
+However, it would be a deictic reference and point back to the mention of a real \emph{NE}, as in the following coreference chain: \\
 
-\ex. $ Prof.\; Bateman \Rightarrow he \Rightarrow the \; prof$ \\
+\ex. $ Prof.\; Bateman \Rightarrow he \Rightarrow the \; prof$ \label{cor}\\
 

### 2012-11-08 src/structured_ner/case/truecaser.py
removing conf matrix tests.

@@ -1,0 +1,8 @@
+class TrueCaser:
+
+    def case_pairs(self, token_tags):
+        tokens, tags = zip(*token_tags)
+        return self.case(list(tokens), list(tags))
+
+    def case(self, tokens, pos_tags):
+        pass

### 2012-11-08 src/structured_ner/case/TrueCaser.py
.

@@ -1,8 +1,0 @@
-class TrueCaser:
-
-    def case_pairs(self, token_tags):
-        tokens, tags = zip(*token_tags)
-        return self.case(list(tokens), list(tags))
-
-    def case(self, tokens, pos_tags):
-        pass

### 2012-11-08 doc/report/paper.tex
++++lingsec

@@
 
-Named Entities can consist of more than one syntactic type. The usual overall type for a named entity is noun phrase, which can be both simple and complex. 
-An example for a simple heterogenous entity is displayed in syntactic tree represenation in ~\ref{tree1}.
- \Tree
+Named Entities can consist of more than one syntactic type. Syntactic types feature in syntactic rules that define, how sentences can be build.
+Typical rules are: \\
+$S \rightarrow NP VP $ \footnote{ A sentence S breaks down into a noun phrase NP and a verb phrase VP} \\
+$ VP \rightarrow V (NP)$ \footnote{ A verb phrase VP consists of a verb followed by an optional NP} \\
+
+The usual overall type for a named entity is noun phrase, which can be both simple and complex, meaning it can consist of a only a noun or 
+also additional other syntactic types.
+An example for a simple homogenous entity is displayed in syntactic tree representation in ~\ref{tree1}.
+
+\begin{enumerate}
+ \item \Tree
     [.NP [.N' John ]  ] \label{tree1}
+\end{enumerate}
 
+\emph{N'} signals an intermediate state in the syntactic tree where adjuncts can be joined to the noun phrase. 
+The closer the entity is joined to a tree, the more it usually contributes to the meaning. 
 Most languages allow for a recursive definition of, for instance, a noun phrase that in theory can grow infinitely big. 
-An \textbf{NP} can consist of a determiner, such as \emph{the} and a common noun or simply a propernoun. \textbf{PP} stands
-for prepositional phrase and consists of a preposition and a noun phrase. 
-The syntactic rules for PP-attachment are displayed in \ref{PP}.
+A \textbf{NP} can consist of a determiner, such as \emph{the} and a common noun or simply a propernoun. \textbf{PP} stands
+for prepositional phrase (PP) and consists of a preposition and a noun phrase. 
+The syntactic rules for PPs are displayed in \ref{PP}.
 \begin{figure}
-$ NP \rightarrow NP PP$ \\
-$ PP \rightarrow P NP$ 
+$ NP \rightarrow NP \;PP$ \\
+$ PP \rightarrow P \;NP$ \\
 \caption{Recursive Noun Phrase}
@@
 
-\ex. The princess of Scotland is in the castle in England with the brown bricks on its roof... \label{PPrec}
+\ex. The Bank of Scotland is in the brown building with the brown bricks on its roof... \label{PPrec}
 
-We distinguish between PP-complements and PP-modifiers, where modifiers simply add to the description of the noun phrase, maybe
-to set it apart and complements are more inherent to the meaning of the entity. 
-The PP-attachment \emph{of Scotland} is semantically defining for the noun \emph{princess}, whereas \emph{with the brown bricks on its roof} 
-just adds to the description of \emph{the castle in England}.
-Regarding named entities, one wants to differentiate between a PP-attachment that defines the entity an one that is irrelevant.
+We distinguish between PP-complements and PP-modifiers, where modifiers simply add to the description of the noun phrase, and complements are more inherent to the meaning of the entity. 
+The PP \emph{of Scotland} is semantically defining for the noun \emph{bank}, whereas \emph{with the brown bricks on its roof} 
+just adds to the description of \emph{brown building}.
+Regarding named entities, one wants to differentiate between a PP that defines the entity an one that is not inherent to its meaning. 
+The phrases in \ref{bos} and \ref{Nirvana} give examples of how a syntactic tree defines semantic connection.
+In \emph{Bank of Scotland}, the PP \emph{of Scotland} is attached directly, whereas in \ref{Nirvana} both \emph{Kurt Cobain} and \emph{Nirvana} constitute own 
+concepts and are joined at a later level to express this distance.
+Thus, in connection with Named Enitites, we want to be able to capture types of the first example as a NE, but not of the second.
 
 \begin{enumerate}
-\item \parbox[t]{2.4in}{ (a)~\Tree 
-   [.NP [ [.DT the ] [.N\1 [.N\1 Princess ] [.PP [.P of ] [.NP England ] ] ] ] ]}
+\item \parbox[t]{2.4in}{ (a)~\Tree  \label{bos}
+   [.NP [ [.DT the ] [.N\1 [.N Bank ] [.PP [.P of ] [.NP Scotland ] ] ] ] ]}
 \parbox[t]{2in}{ (b)~\Tree 
-   [.NP  [.DT the ] [.N\1 [.N\1 Princess ] ] [.PP [.P from ] [.NP England ] ]  ] ]  }
+   [.NP  [.N\1 [.N\1  \qroof{Kurt Cobain}.N  ]  [.PP [.P of ] [.N\1 [.N\1 [.N Nirvana ] ] ] ] ] ]  } \label{Nirvana}
 \end{enumerate}

### 2012-11-09 doc/report/paper.tex
++

@@
 \begin{enumerate}
-\item \parbox[t]{2.4in}{ (a)~\Tree  \label{bos}
-   [.NP [ [.DT the ] [.N\1 [.N Bank ] [.PP [.P of ] [.NP Scotland ] ] ] ] ]}
+\item \parbox[t]{2.4in}{ (a)~\Tree  
+   [.NP [ [.DT the ] [.N\1 [.N Bank ] [.PP [.P of ] [.NP Scotland ] ] ] ] ] } \label{bos}
 \parbox[t]{2in}{ (b)~\Tree 
@@
 \subsection{CoNLL Data and Baseline Performance}
-The training and test dataset was taken from the \emph{CoNLL Shared Task 2002} \cite{tksintro2002conll} and the \emph{CoNLL Shared Task 2003} \cite{TjongKimSang:2003:ICS:1119176.1119195}.
-for Spanish \& Dutch and English \& German respectively. The ConNLL shared tasks are organised challenges, where the organiser propose a task and provide the
-training and test data. For 2002/2003 it was Named Entity Recognition. Also we base our system on the approach that won the challenge for English and German in 2003. 
-The baseline rate for all languages was produced by a system, that had a unique class in the training data. In case that a phrase was part of more than one entity, 
-the system would choose the longest one \cite{TjongKimSang:2003:ICS:1119176.1119195}. 
-The data is annotated with POS-tags and NE-tags, as well as Bio-tags that indicate the position in the Named Entity. 
+The training and test data for Spanish \& Dutch was taken from the \emph{CoNLL Shared Task 2002} \cite{tksintro} and 
+the \emph{CoNLL Shared Task 2003} \cite{TjongKimSang:2003:ICS:1119176.1119195} provided the one for English \& German.
+The ConNLL shared tasks are organised challenges, where the organisers propose a task and provide the participants with the
+training and test data. 
+In 2002 and 2003, the task was Named Entity Recognition. 
+In both years, the baseline rate for the individual languages was produced by a system, that had a unique class in the training data. 
+In case that a phrase was part of more than one entity, the system would choose the longest one \cite{TjongKimSang:2003:ICS:1119176.1119195}. 
+Table \ref{table:Base} shows the baseline for the four different languages.  
+The data is annotated with part-of-speech tags and named-entity tags, as well as Bio-tags that indicate the position in the Named Entity. 
+Our approach is based on the system that won the challenge for English and German in 2003.
+
 % A baseline rate was computed for the English and the
@@
 
-The baseline for \emph{Named Entity Recognition} in the four different languages is visualised in table \
 \begin{table}[h!]
@@
 \exg. \textbf{x:} The motor company ' Ford ' was founded and incorporated by Henry Ford .\\
-      \textbf{y:} O O O O ORG  O  O O O O O PER PER O  \label{PredEx1b} \\
+      \textbf{y:}  O   O      O     0 ORG  O  O     O     O       O        O PER   PER  O  \label{PredEx1b} \\
 \caption{Input and predicted structure for the Named Entity task.}

### 2012-11-09 doc/report/paper.tex
++++methods

@@
 \emph{Metonymy} is frequently an issue in literary texts and news data (which is often used in NER), where two items that are in a part-whole
-relationship, are substituted for each other respectively. Example \ref{ex:Lon} shows an instance of \emph{whole-part}, where \emph{London} is supposedly substituted
-for the \textbf{Goverment in London}.
+relationship, are substituted for each other respectively. Example \ref{ex:Lon} shows an instance of \emph{whole-part}, 
+where \emph{London} is supposedly substituted for the \textbf{Goverment in London}.
 
@@
 \subsection{Methods employed for NER}
-For \emph{NER}, there exist both rule-based and statisical approaches. 
-Rule-based methods make use of the underlying rules governing languages to extract named entities. 
-However, this approach is high maintenance, quite time-consuming and requires extensive work of computational linguists \cite{nadeau2007survey}
-and although the results are often high in precision, lacks considerably in recall.
+For \emph{NER}, there exist two general approaches: one is rule-based and the other statistical. 
+Rule-based methods make use of the underlying rules governing a language to extract the named entities. 
+This approach, however, is rather high maintenance and requires extensive work by experts such as computational linguists \cite{nadeau2007survey}.
+The results are often quite high in precision, but lack considerably in recall. % rules are partly language-dependent
 
-In regard to statistical approaches, supervised-learning is the most common method applied in \emph{NER}. Although, there are unsupervised approaches, 
-their performance is not as high as for SL applications yet.
-
+For statistical approaches, supervised-learning is the most commonly method applied in \emph{NER}. 
+There are unsupervised methods, but they cannot compete with SL applications yet.
 Prominent algorithms in NER include maximum entropy models,
@@
 Table \ref{table:Base} shows the baseline for the four different languages.  
-The data is annotated with part-of-speech tags and named-entity tags, as well as Bio-tags that indicate the position in the Named Entity. 
-Our approach is based on the system that won the challenge for English and German in 2003.
+The data is annotated with part-of-speech tags and named-entity tags, as well as Bio-tags that indicate the position in the Named Entity. %put in ``tokenized, chunked''
+Our approach is based on the system of \cite{strlearn} that won the challenge for English and German in 2003.
 
@@
 For the evaluation of the system we used \emph{Precision} and \emph{Recall} as shown in \ref{Precision} and \ref{Recall} respectively.
+Precision is the percentage of Named Entities found by the system that are correct. Recall represents the the percentage of 
+named entites present in the corpus that were found by the system. 
+A named entity is only deemed correct if it is an exact match for the corresponding data file.
 The general formula for the \emph{F-Score} is shown in \ref{Fscore}. Since we rate both \emph{Precision} and \emph{Recall} evenly, we 

### 2012-11-09 doc/report/paper.tex
++++methods

@@
 Locations&  Paris, Australia, Bristol\\
-Miscellaneous& World Cup 2014\\
+Miscellaneous& World Cup 2014, Italien\\
  Date Expression& 01-02-2012, January, 6th, 1998 \\
@@
 There are unsupervised methods, but they cannot compete with SL applications yet.
-Prominent algorithms in NER include maximum entropy models,
-
+Among the most prominent algorithms in NER are maximum entropy models. For CoNLL 2003, five of the sixteen systems
+used this statistical learning method; three of them in isolation and they seem to be a very suitable choice for this task, since the three highest results in 
+English and the two highest results in German were obtained by systems that employed this approach \cite{TjongKimSang:2003:ICS:1119176.1119195}.
+The best system in both German and English \cite{Florian:2003:NER:1119176.1119201} used a combination of classifiers: a robust linear classifier, a maximum entropy, a transformation-based learning, and a hidden Markov model which are connected through a linear interpolation of the classifiers’ % combination scheme
+class probability distribution. Particular importance is also assigned to creating a rich feature space, also including a gazetteer feature. 
+This system gains a precision of 88.99\% on the English test set and  83.87\% on the German one. 
 neural network
@@
 The data is annotated with part-of-speech tags and named-entity tags, as well as Bio-tags that indicate the position in the Named Entity. %put in ``tokenized, chunked''
-Our approach is based on the system of \cite{strlearn} that won the challenge for English and German in 2003.
+Our approach is based on the system of \cite{strlearn} that won the challenge for Spanish and Dutch in 2002.
 

### 2012-11-09 doc/report/paper.tex
++++methods

@@
 \begin{abstract}
-Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+This paper describes an approach to \namedentity Recognition using Structured Prediction. The System was applied to four different languages: Spanish, Dutch, English and
+German. Various sets of features were employed among them a Gazetteer feature that obtained a look-up list from the database DBPedia and which proved very useful.
+The results were...
 \end{abstract}
@@
 An additional type \texttt{Miscellaneous} captures names outside the classic \emph{enamax} type.
-Apart from the above, there is the \emph{timex} type, which covers date \& time expressions and \emph{numex} which is for monetary values \& percent. 
-Table \ref{table:NETypes} gives an overview of the different NE types.
+Apart from the above, there is also the \emph{timex} type, which covers date and time expressions and \emph{numex} for monetary values and percentages. 
+Table \ref{table:NETypes} gives an overview of the different \namedentity (NE) types.
 
@@
 
-In considering named entities, it is important to distinguish between named entities and the mention of entities that are in fact non-specific. % reformulate
-The difference lies in the reference of the entity. A named entity refers to a specific, unique person/time/location, whereas in 
-mere mentions of entites, these can refer to multiple events, such as the time expression: \emph{In June}, referring to any possible year.
-Also, there could be a reference to a person, as in: \emph{the prof}, which in itself is not a specific entity.
-However, it would be a deictic reference and point back to the mention of a real \emph{NE}, as in the following coreference chain: \\
+In considering \namedentity, it is important not to confuse them with non-specific entities. % reformulate
+The difference lies in the reference of the entity. A \namedentity refers to a specific, unique person/time/location, whereas
+mere mentions of entites can refer to multiple events, such as the time expression: \emph{In June}, which could refer to any year.
+Also, there could be a reference to a person, as in: \emph{the prof}, which in itself is not a \namedentity.
+However, it would be a deictic reference and point back to the mention of a real \namedentity, as in the following coreference chain: \\
 
@@
 
-Named Entities can consist of more than one syntactic type. Syntactic types feature in syntactic rules that define, how sentences can be build.
-Typical rules are: \\
-$S \rightarrow NP VP $ \footnote{ A sentence S breaks down into a noun phrase NP and a verb phrase VP} \\
-$ VP \rightarrow V (NP)$ \footnote{ A verb phrase VP consists of a verb followed by an optional NP} \\
+A \namedentity  can consist of more than one syntactic type. Syntactic types feature in syntactic rules that define how sentences can be build.
+Typical high-level rules include: \\
+$s \rightarrow np\quad vp $ \footnote{ A sentence s breaks down into a noun phrase np and a verb phrase vp} \\
+$ vp \rightarrow v\quad (np)$ \footnote{ A verb phrase vp consists of a verb followed by an optional np} \\
 
-The usual overall type for a named entity is noun phrase, which can be both simple and complex, meaning it can consist of a only a noun or 
-also additional other syntactic types.
-An example for a simple homogenous entity is displayed in syntactic tree representation in ~\ref{tree1}.
+Thus, sentences are often represented as syntactic trees that model the dependencies among the types.
+The overall type for a \namedentity is usually a noun phrase (np), which can be both simple and complex, meaning it can consist of a only a noun phrase or 
+also additional other syntactic types, such as prepositional phrases (pp), that consist of a preposition and a noun phrase. 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% - can be left out if not enough space/too detailed
+It is therefore a recursive definition and in theory, noun phrases can grow infinitely big, as shown by the following recursive rules:  %
 
-\begin{enumerate}
- \item \Tree
-    [.NP [.N' John ]  ] \label{tree1}
-\end{enumerate}
+\ex. $ np \rightarrow np \quad pp$ \\
+     $ pp \rightarrow p \quad np$ 
+ %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\emph{N'} signals an intermediate state in the syntactic tree where adjuncts can be joined to the noun phrase. 
-The closer the entity is joined to a tree, the more it usually contributes to the meaning. 
-Most languages allow for a recursive definition of, for instance, a noun phrase that in theory can grow infinitely big. 
-A \textbf{NP} can consist of a determiner, such as \emph{the} and a common noun or simply a propernoun. \textbf{PP} stands
-for prepositional phrase (PP) and consists of a preposition and a noun phrase. 
-The syntactic rules for PPs are displayed in \ref{PP}.
+We distinguish between pp-complements and pp-modifiers; where complements are non-optional elements of the noun phrase as they are
+inherent to the meaning of the entity and modifiers simply add to the description and are optional. 
+The syntactic trees in \ref{fig:bos} and \ref{fig:nirvana}  show two examples for a complex noun phrase containing a \namedentity.
+\emph{N'} signals an intermediate state in the syntactic tree where modifiers can be joined.
+In \ref{fig:bos}, (\emph{Bank of Scotland}), the pp \emph{of Scotland} is attached directly; it is central to the meaning, 
+whereas in \ref{fig:nirvana} both \emph{Kurt Cobain} and \emph{Nirvana} constitute own concepts and are joined at a later level to express this distance.
+Thus, in connection with Named Entities, we want to be able to capture types of the first example as a \namedentity, but not of the second more distantly joined 
+relation.
+% \begin{figure}
+% \begin{enumerate}
+% \item \parbox[t]{2.4in}{~\Tree  
+%    [.NP [ [.DT the ] [.N\1 [.N Bank ] [.PP [.P of ] [.NP Scotland ] ] ] ] ] } \label{bos}
+% \item \parbox[t]{2in}{~\Tree 
+%    [.NP  [.N\1 [.N\1  \qroof{Kurt Cobain}.N  ]  [.PP [.P of ] [.N\1 [.N\1 [.N Nirvana ] ] ] ] ] ]  } \label{Nirvana}
+% \end{enumerate} \label{Tree}
+% \end{figure}
+
 \begin{figure}
-$ NP \rightarrow NP \;PP$ \\
-$ PP \rightarrow P \;NP$ \\
-\caption{Recursive Noun Phrase}
-\label{PP}
+\parbox[t]{2.4in}{~\Tree  
+   [.NP [ [.DT the ] [.N\1 [.N Bank ] [.PP [.P of ] [.NP Scotland ] ] ] ] ] }
+\label{fig:bos}
+\caption{\namedentity: Complement Meaning}
 \end{figure}
-These rules allow for constructions such as in sentence~\ref{PPrec}.
 
-\ex. The Bank of Scotland is in the brown building with the brown bricks on its roof... \label{PPrec}
+\begin{figure}
+\parbox[t]{2in}{~\Tree 
+   [.NP  [.N\1 [.N\1  \qroof{Kurt Cobain}.N  ]  [.PP [.P of ] [.N\1 [.N\1 [.N Nirvana ] ] ] ] ] ]  }
+\label{fig:nirvana}
+\caption{\namedentity: Reference Meaning}
+\end{figure}
 
-We distinguish between PP-complements and PP-modifiers, where modifiers simply add to the description of the noun phrase, and complements are more inherent to the meaning of the entity. 
-The PP \emph{of Scotland} is semantically defining for the noun \emph{bank}, whereas \emph{with the brown bricks on its roof} 
-just adds to the description of \emph{brown building}.
-Regarding named entities, one wants to differentiate between a PP that defines the entity an one that is not inherent to its meaning. 
-The phrases in \ref{bos} and \ref{Nirvana} give examples of how a syntactic tree defines semantic connection.
-In \emph{Bank of Scotland}, the PP \emph{of Scotland} is attached directly, whereas in \ref{Nirvana} both \emph{Kurt Cobain} and \emph{Nirvana} constitute own 
-concepts and are joined at a later level to express this distance.
-Thus, in connection with Named Enitites, we want to be able to capture types of the first example as a NE, but not of the second.
-
-\begin{enumerate}
-\item \parbox[t]{2.4in}{ (a)~\Tree  
-   [.NP [ [.DT the ] [.N\1 [.N Bank ] [.PP [.P of ] [.NP Scotland ] ] ] ] ] } \label{bos}
-\parbox[t]{2in}{ (b)~\Tree 
-   [.NP  [.N\1 [.N\1  \qroof{Kurt Cobain}.N  ]  [.PP [.P of ] [.N\1 [.N\1 [.N Nirvana ] ] ] ] ] ]  } \label{Nirvana}
-\end{enumerate}
+% \ex. The Bank of Scotland is in the brown building with the brown bricks on its roof... \label{PPrec}
+% 
+% The PP \emph{of Scotland} is semantically defining for the noun \emph{bank}, whereas \emph{with the brown bricks on its roof} 
+% just adds to the description of \emph{brown building}.
 
 \subsection{Issues in Named Entity Recognition}
-Among the most common problems for NER is the issue of ambiguity, that is in particular \emph{Polysemy} \cite{nadeau2007survey}, 
-the property of some lexical representation having
-more than one possible meaning, and \emph{Metonymy}, which refers to the concept of the part-whole/whole-part relation between two expressions. 
-\emph{Polysemy} could become an issue for NER when the lexical representation of an item could point to two different \emph{NE} types.
+Among the most common challenges for NER is the issue of ambiguity, in particular \emph{Polysemy} and \emph{Metonymy} \cite{nadeau2007survey}.
+Polysemy refers to the property of a word that is a lexical representation having more than one possible meaning.
+\emph{Polysemy} could become an issue for NER when this lexical representation points to two different \namedentity types.
 This is quite frequent with \texttt{Person Names} and \texttt{Locations}, since many people are named after cities, such as \emph{Paris} or \emph{Georgia}. 
-Often the context will not be disambiguating as in \ref{ex:Paris}.
+Often the context will not be disambiguating as in example~\ref{ex:Georgia}.
 
-\ex. \emph{Paris is beautiful.} \label{ex:Paris}
+\ex. \emph{Georgia is beautiful.} \label{ex:Georgia}
 
-\emph{Metonymy} is frequently an issue in literary texts and news data (which is often used in NER), where two items that are in a part-whole
-relationship, are substituted for each other respectively. Example \ref{ex:Lon} shows an instance of \emph{whole-part}, 
-where \emph{London} is supposedly substituted for the \textbf{Goverment in London}.
+\emph{Metonymy} is the relationship of a part-whole relation between two items.
+It is frequently employed in literary texts and news data (which is often used in NER), where one item is substituted for another. 
+Example \ref{ex:Lon} shows an instance of a \emph{whole-part}, where \emph{London} probably stands for the smaller 'item': \textbf{Goverment in London}.
 
@@
 \subsection{Methods employed for NER}
-For \emph{NER}, there exist two general approaches: one is rule-based and the other statistical. 
-Rule-based methods make use of the underlying rules governing a language to extract the named entities. 
-This approach, however, is rather high maintenance and requires extensive work by experts such as computational linguists \cite{nadeau2007survey}.
-The results are often quite high in precision, but lack considerably in recall. % rules are partly language-dependent
+For NER, there exist two general approaches: one is rule-based and the other statistical. 
+Rule-based methods make use of the underlying rules governing a language to extract \namedentity. 
+This approach, however, is rather high maintenance and requires extensive work by experts, such as computational linguists \cite{nadeau2007survey}.
+The results can be quite high in precision, but lack considerably in recall. % rules are partly language-dependent
 
-For statistical approaches, supervised-learning is the most commonly method applied in \emph{NER}. 
+For statistical approaches, supervised-learning is the most common method applied in NER. 
 There are unsupervised methods, but they cannot compete with SL applications yet.
-Among the most prominent algorithms in NER are maximum entropy models. For CoNLL 2003, five of the sixteen systems
-used this statistical learning method; three of them in isolation and they seem to be a very suitable choice for this task, since the three highest results in 
-English and the two highest results in German were obtained by systems that employed this approach \cite{TjongKimSang:2003:ICS:1119176.1119195}.
+Among the most prominent algorithms in NER are maximum entropy models. 
+For CoNLL 2003, five of the sixteen systems used this statistical learning method and three of of those used maximum entropy models in isolation.
+Maximum entropy models seem a very suitable choice for the NER task, since systems using this approach were ranked among the three highest results in 
+English and the two highest results in German \cite{TjongKimSang:2003:ICS:1119176.1119195}.
 The best system in both German and English \cite{Florian:2003:NER:1119176.1119201} used a combination of classifiers: a robust linear classifier, a maximum entropy, a transformation-based learning, and a hidden Markov model which are connected through a linear interpolation of the classifiers’ % combination scheme
@@
 \subsection{CoNLL Data and Baseline Performance}
-The training and test data for Spanish \& Dutch was taken from the \emph{CoNLL Shared Task 2002} \cite{tksintro} and 
-the \emph{CoNLL Shared Task 2003} \cite{TjongKimSang:2003:ICS:1119176.1119195} provided the one for English \& German.
+The training and test data for Spanish \& Dutch was taken from the CoNLL Shared Task 2002 \cite{tksintro} and 
+the CoNLL Shared Task 2003 \cite{TjongKimSang:2003:ICS:1119176.1119195} provided the one for English \& German.
 The ConNLL shared tasks are organised challenges, where the organisers propose a task and provide the participants with the
@@
       \textbf{y:}  O   O      O     0 ORG  O  O     O     O       O        O PER   PER  O  \label{PredEx1b} \\
-\caption{Input and predicted structure for the Named Entity task.}
+\caption{Input and predicted structure for the \namedentity task.}
 
@@
 
+This paper presented a system for \namedentity Recognition using Structured Prediction.
+
 \bibliographystyle{acl}

### 2012-11-09 doc/report/paper.tex
Merge+minor changes.

@@
 \begin{abstract}
-This paper describes an approach to \namedentity Recognition using Structured Prediction. The System was applied to four different languages: Spanish, Dutch, English and
-German. Various sets of features were employed among them a Gazetteer feature that obtained a look-up list from the database DBPedia and which proved very useful.
-The results were...
+Named Entity Recognition is a task
+
 \end{abstract}
@@
 \paragraph*{Gazetteer Features}
-In order to create gazetteer lists for the more common named entities, we designed a \emph{SPARQL} query, that would retrieve entries from \emph{DBPedia} for all
-languages. The reliability of the respective list is learnt be the perceptron.
+A technique commonly used in Named Entity Recognition is the use of external lists of names of a specific type (gazetteer).
+In order to create gazetteer lists for the more common named entities, we use a \emph{SPARQL} query to retrieve entries from \emph{DBpedia} for all required languages.
+
+\begin{figure}
+\begin{verbatim}
+PREFIX dbp: <http://dbpedia.org/ontology/>
+SELECT ?label
+WHERE {
+    ?entity rdf:type dbp:CLASS.
+    ?entity rdfs:label ?label.
+    FILTER (lang(?label) = LANGUAGE)
+}
+\end{verbatim}
+\caption{SPARQL query for retrieving names of the class \texttt{CLASS} from DBpedia.}
+\end{figure}
 
@@
 
-\begin{table}[h!]
-\scriptsize
+\begin{table*}[t]
+\centering
 \begin{tabular}{| l | l l l| l l l |}
@@
 \label{table:Results}
-\end{table}
+\end{table*}
 

### 2012-11-10 doc/report/paper.tex
++++methods

--- 2012-11-09 doc/report/paper.tex
@@
 \begin{figure}
-\parbox[t]{2.4in}{~\Tree  
-   [.NP [ [.DT the ] [.N\1 [.N Bank ] [.PP [.P of ] [.NP Scotland ] ] ] ] ] }
+\Tree  
+   [.NP [ [.DT the ] [.N\1 [.N Bank ] [.PP [.P of ] [.NP Scotland ] ] ] ] ]
+
+\caption{\namedentity: Complement Meaning}
 \label{fig:bos}
-\caption{\namedentity: Complement Meaning}
 \end{figure}
@@
 \begin{figure}
-\parbox[t]{2in}{~\Tree 
-   [.NP  [.N\1 [.N\1  \qroof{Kurt Cobain}.N  ]  [.PP [.P of ] [.N\1 [.N\1 [.N Nirvana ] ] ] ] ] ]  }
+\Tree 
+   [.NP  [.N\1 [.N\1  \qroof{Kurt Cobain}.N  ]  [.PP [.P of ] [.N\1 [.N\1 [.N Nirvana ] ] ] ] ] ]
+
+\caption{\namedentity: Reference Meaning}
 \label{fig:nirvana}
-\caption{\namedentity: Reference Meaning}
 \end{figure}
@@
 There are unsupervised methods, but they cannot compete with SL applications yet.
-Among the most prominent algorithms in NER are maximum entropy models. 
+For the shared task in 2002 \cite{tksintro}, 12 systems participated in the challenge and presented a wide variety of approaches to the \namedentity task.
+The most successful system in the task, used AdaBoost applied to fixed-depth decision trees. Special attention was given to a variety of input features and the features were also
+partly adapted for each language with an F-score of 81.39 on the Spanish test set and 77.05 on the Dutch one.
+The second highest system \cite{Florian:2002:NER:1118853.1118863} employed three stacked
+learners (transformation-based learning for obtaining base-level non-typed Named Entites, Snow for improving the quality of these entities and the forward-backward algorithm for
+finding categories for the Named Entites), where the combination of these three algorithms substantially outperformed any subset of them, with an F-score of 79.05 on the Spanish test
+set and 74.99 on the Dutch one.
+Among the most prominent algorithms in NER are maximum entropy models, whose benefit in regard to this task became most apparent in CoNLL 2003. 
 For CoNLL 2003, five of the sixteen systems used this statistical learning method and three of of those used maximum entropy models in isolation.
@@
 English and the two highest results in German \cite{TjongKimSang:2003:ICS:1119176.1119195}.
-The best system in both German and English \cite{Florian:2003:NER:1119176.1119201} used a combination of classifiers: a robust linear classifier, a maximum entropy, a transformation-based learning, and a hidden Markov model which are connected through a linear interpolation of the classifiers’ % combination scheme
+The best system in both German and English \cite{Florian:2003:NER:1119176.1119201} used a combination of classifiers: a robust linear classifier, 
+a maximum entropy, a transformation-based learning, and a hidden Markov model which are connected through a linear interpolation of the classifiers’ % combination scheme
 class probability distribution. Particular importance is also assigned to creating a rich feature space, also including a gazetteer feature. 
 This system gains a precision of 88.99\% on the English test set and  83.87\% on the German one. 
-neural network
+More recent approaches to NER include Conditional Random Fields and Neural Networks.
 
@@
 The data is annotated with part-of-speech tags and named-entity tags, as well as Bio-tags that indicate the position in the Named Entity. %put in ``tokenized, chunked''
-Our approach is based on the system of \cite{strlearn} that won the challenge for Spanish and Dutch in 2002.
-
-% A baseline rate was computed for the English and the
-% German test sets. It was produced by a system which
-% only identified entities which had a unique class in
-% the training data. If a phrase was part of more than
-% one entity, the system would select the longest one.
-% All systems that participated in the shared task have
-% outperformed the baseline system.
-% For all the Fβ=1 rates we have estimated sig-
-% nificance boundaries by using bootstrap resampling
-% (Noreen, 1989). From each output file of a system,
-% 250 random samples of sentences have been chosen
-% and the distribution of the Fβ=1 rates in these sam-
-% ples is assumed to be the distribution of the perfor-
-% mance of the system. We assume that performance
-% A is significantly different from performance B if A
-% is not within the center 90% of the distribution of B.
+Our approach is based on \cite{strlearn}, the group that won the challenge for Spanish and Dutch in 2002.
 
@@
 % problem external data / use Dbpedia - gazetteer
-Our implementation for the languages English, German, Dutch and Spanish is trained and tested on the \emph{CoNLL 2003} and \emph{CoNLL 2002} data sets respectively.
+Our implementation for the languages English, German, Dutch and Spanish is trained and tested on the CoNLL 2003 and CoNLL 2002 data sets respectively.
 

### 2012-11-10 doc/report/paper.tex
++++methods

@@
 
-In considering \namedentity, it is important not to confuse them with non-specific entities. % reformulate
-The difference lies in the reference of the entity. A \namedentity refers to a specific, unique person/time/location, whereas
-mere mentions of entites can refer to multiple events, such as the time expression: \emph{In June}, which could refer to any year.
+In considering Named Entities, it is important not to confuse them with non-specific entities. % reformulate
+The difference lies in the reference of the entity. A Named Entity refers to a specific, unique person/time/location, whereas
+\emph{normal} entites can refer to multiple events, such as the time expression: \emph{In June}, which could refer to any year.
 Also, there could be a reference to a person, as in: \emph{the prof}, which in itself is not a \namedentity.
-However, it would be a deictic reference and point back to the mention of a real \namedentity, as in the following coreference chain: \\
+%However, it would be a deictic reference and point back to the mention of a real \namedentity, as in the following coreference chain: \\
 
-\ex. $ Prof.\; Bateman \Rightarrow he \Rightarrow the \; prof$ \label{cor}\\
-
-A \namedentity  can consist of more than one syntactic type. Syntactic types feature in syntactic rules that define how sentences can be build.
+%\ex. $ Prof.\; Bateman \Rightarrow he \Rightarrow the \; prof$ \label{cor}\\
+A Named Entity  can consist of more than one syntactic type. Syntactic types feature in syntactic rules that define how sentences can be build.
 Typical high-level rules include: \\
@@
 Thus, sentences are often represented as syntactic trees that model the dependencies among the types.
-The overall type for a \namedentity is usually a noun phrase (np), which can be both simple and complex, meaning it can consist of a only a noun phrase or 
+The overall type for a Named Entity is usually a noun phrase (np), which can be both simple and complex, meaning it can consist of a only a noun phrase or 
 also additional other syntactic types, such as prepositional phrases (pp), that consist of a preposition and a noun phrase. 
@@
 Polysemy refers to the property of a word that is a lexical representation having more than one possible meaning.
-\emph{Polysemy} could become an issue for NER when this lexical representation points to two different \namedentity types.
+\emph{Polysemy} could become an issue for NER when this lexical representation points to two different Named Entity types.
 This is quite frequent with \texttt{Person Names} and \texttt{Locations}, since many people are named after cities, such as \emph{Paris} or \emph{Georgia}. 
@@
 For NER, there exist two general approaches: one is rule-based and the other statistical. 
-Rule-based methods make use of the underlying rules governing a language to extract \namedentity. 
+Rule-based methods make use of the underlying rules governing a language to extract Named Entities. 
 This approach, however, is rather high maintenance and requires extensive work by experts, such as computational linguists \cite{nadeau2007survey}.
@@
 There are unsupervised methods, but they cannot compete with SL applications yet.
-For the shared task in 2002 \cite{tksintro}, 12 systems participated in the challenge and presented a wide variety of approaches to the \namedentity task.
+For the shared task in 2002 \cite{tksintro}, 12 systems participated in the challenge and presented a wide variety of approaches to the Named Entity task.
 The most successful system in the task, used AdaBoost applied to fixed-depth decision trees. Special attention was given to a variety of input features and the features were also
@@
 This system gains a precision of 88.99\% on the English test set and  83.87\% on the German one.
-More recent approaches to NER include Conditional Random Fields and Neural Networks.
 
@@
 \section{Structured Prediction}
-Structured Prediction \cite{strlearn} is a supervised-learning approach that maps an input \textbf{x} to an output \textbf{y}: \\
+Prediction \cite{strlearn} is a supervised-learning approach, where a system uses training data to learn the prediction
+of unseen data. It consequently maps an input \textbf{x} to an output \textbf{y}: \\
 
@@
 
-Structured and  Non-structured prediction differ in the form of their output. 
-Non-structured output is atomic; it is binary prediction for a two-class problem and may corresponds to more than one of more than two possible labels for a 
-multiclass problem. 
-Structured Prediction gives back the prediction for the whole sequence. 
-For the sentence in \ref{PredEx1a}, it would return the corresponding labels combination in example~\ref{PredEx1b}.
+Structured Prediction and Non-structured prediction differ in the form of their output.
+Non-Structured Prediction returns an atomic output: it is binary prediction for a two-class problem and one of many labels for a multiclass problem. 
+Structured Prediction, on the other hand, returns the prediction for a whole sequence. 
+For the sentence in \ref{PredEx1a}, it would return the corresponding label combination in example~\ref{PredEx1b}.
 
@@
 
-Also, Structured Prediction is different from similar approaches through its combination of features and label interactions instead
+In comparison to other similar approaches, Structured Prediction sets itself apart through the combination of features and label interactions instead
 of concentrating mainly on label interactions like HMM or a rich set of features like local classifiers.
@@
 
-Learning is implemented using the Structured Perceptron algorithm.
-To avoid overfitting, we average the parameters after the last iteration \cite{collins2002discriminative}.
-For finding the best sequence of labels for a sentence, we use a modified version of the Viterbi algorithm.
-The algorithm finds the sequence $\hat{y}$, such that:
+Learning is implemented using the Structured Perceptron algorithm. To avoid overfitting, we average the parameters after the last iteration \cite{collins2002discriminative}. 
+For finding the best sequence of labels for a sentence, we use a modified version of the Viterbi algorithm. The algorithm finds the sequence $\hat{y}$, such that:
 

### 2012-11-10 doc/report/paper.tex
++++methods

@@
 Apart from the above, there is also the \emph{timex} type, which covers date and time expressions and \emph{numex} for monetary values and percentages. 
-Table \ref{table:NETypes} gives an overview of the different \namedentity (NE) types.
+Table \ref{table:NETypes} gives an overview of the different Named Entity (NE) types.
 
@@
 The data is annotated with part-of-speech tags and named-entity tags, as well as Bio-tags that indicate the position in the Named Entity. %put in ``tokenized, chunked''
-Our approach is based on \cite{strlearn}, the group that won the challenge for Spanish and Dutch in 2002.
+Our approach is based on \cite{strlearn}, of the group that won the challenge for Spanish and Dutch in 2002.
 
@@
 \section{Structured Prediction}
-Prediction \cite{strlearn} is a supervised-learning approach, where a system uses training data to learn the prediction
+Prediction \cite{Taskar:2005:LSP:1102351.1102464} is a supervised-learning approach, where a system uses training data to learn the prediction
 of unseen data. It consequently maps an input \textbf{x} to an output \textbf{y}: \\
@@
 \section{Implementation}
+In this section, we describe our implementation, the learning and decoding algorithm and the features.
 

### 2012-11-10 doc/report/paper.tex
Gazetteer and Issues.

--- 2012-11-10 doc/report/paper.tex
@@
       \textbf{y:}  O   O      O     0 ORG  O  O     O     O       O        O PER   PER  O  \label{PredEx1b} \\
-\caption{Input and predicted structure for the \namedentity task.}
+\caption{Input and predicted structure for the Named Entity Recognition task.}
 
@@
 These features register which label has been assigned to the previous token and takes into account the most likely sequence. 
-Thus, for the sequence: ``Jack London went to New York'' the NE tag combination of \emph{PER} and \emph{PER} in example~\ref{seq1}is more likely 
-than\emph{PER} and \emph{LOC} as presented in \ref{seq2}.
-.
+Thus, for the sequence: ``Jack London went to New York'' the NE tag combination of \emph{PER} and \emph{PER} in example~\ref{seq1} is more likely than\emph{PER} and \emph{LOC} in \ref{seq2}.
 %current token and last label for prepositions or possessive ’s
@@
 A technique commonly used in Named Entity Recognition is the use of external lists of names of a specific type (gazetteer).
-In order to create gazetteer lists for the more common named entities, we use a \emph{SPARQL} query to retrieve entries from \emph{DBpedia} for all required languages.
+In order to create gazetteer lists for more common named entities, we use a \emph{SPARQL} query  (Figure~\ref{fig:sparql}) to automatically retrieve entries from DBpedia for all required languages from a public SPARQL endpoint.
 
@@
 \begin{verbatim}
-PREFIX dbp: <http://dbpedia.org/ontology/>
+PREFIX d: <http://dbpedia.org/ontology/>
 SELECT ?label
 WHERE {
-    ?entity rdf:type dbp:CLASS.
+    ?entity rdf:type   d:CLASS.
     ?entity rdfs:label ?label.
@@
 \caption{SPARQL query for retrieving names of the class \texttt{CLASS} from DBpedia.}
+\label{fig:sparql}
+
 \end{figure}
 
+Given the lists of entities retrieved, we then mark the longest matches of known entities in the input using a directed acyclic word graph.
+The mention is annotated with a reference to its list of origin, which is then used as a feature in our model.
+With this approach, it is not necessary to manually map the lists to target labels and the model can learn the reliability of each list.
+
+\subsection{Other Issues}
+
+One of the issues that we needed to address is the case of headlines.
+Headlines, mainly in the English data, are often written entirely in uppercase letters.
+This is problematic because from most of the data, the model might learn that a word in all uppercase letters is indicative of an organization (e.g.\ IBM, AIG, NYSE) and might thus mistake the tokens in headlines for organizations.
+We address this issue by converting all-uppercase headlines into their most likely original case (truecasing).
+To estimate the most likely case of a token from a monolingual corpus, we use the truecasing model from the Moses statistical machine translation system \cite{koehn2007moses}.
+
 \section{Experiments and Evaluation}
-In the following section we present our experiments and the evalution of our system.
+In the following section, we present our experiments and the evaluation of our system.
 \subsection*{Experiments}
@@
 
-\subsection*{Some Challenges} % maybe in previous section
-
 \section{Future Work \& Conclusion}

### 2012-11-10 doc/report/paper.tex
Eval+layout.

@@
 Table \ref{table:Base} shows the baseline for the four different languages.  
-The data is annotated with part-of-speech tags and named-entity tags, as well as Bio-tags that indicate the position in the Named Entity. %put in ``tokenized, chunked''
-Our approach is based on \cite{strlearn}, of the group that won the challenge for Spanish and Dutch in 2002.
+The data is annotated with part-of-speech tags and named-entity tags, as well as BIO-tags that indicate the position in the Named Entity. %put in ``tokenized, chunked''
+%Our approach is based on \cite{strlearn}, of the group that won the challenge for Spanish and Dutch in 2002.
 
@@
 \section{Experiments and Evaluation}
-In the following section, we present our experiments and the evaluation of our system.
-\subsection*{Experiments}
 
-\subsection*{Evaluation}
-For the evaluation of the system we used \emph{Precision} and \emph{Recall} as shown in \ref{Precision} and \ref{Recall} respectively.
-Precision is the percentage of Named Entities found by the system that are correct. Recall represents the the percentage of 
-named entites present in the corpus that were found by the system. 
-A named entity is only deemed correct if it is an exact match for the corresponding data file.
-The general formula for the \emph{F-Score} is shown in \ref{Fscore}. Since we rate both \emph{Precision} and \emph{Recall} evenly, we 
-use the harmonic mean as shown in \ref{F1}.
+\subsection{Experiments}
 
-\ex. \emph{Precision} = $ \frac{gold\; tag \bigcap predicted}{predicted}$ \label{Precision}\\
+\subsubsection{Influence of feature sets}
 
-\ex. \emph{Recall} = $ \frac{gold \;tag \bigcap predicted}{gold\; tag}$ \label{Recall}\\
+In the following, we show the influence of each of the three sets of features on the resulting accuracy.
 
-\ex. $F_{\beta}$ = $ (1+\beta^2)*\frac{precision *recall}{\beta^2* precision + recall}$ \label{Fscore}\\
+%INSERT MATLAB PIC HERE
 
-\ex. $F_1$ = $ 2*\frac{precision *recall}{precision + recall}$ \label{F1}\\
+\subsubsection{Extending monolingual gazetteers}
+
+% RESULTS FOR DUTCH
+
+\subsection{Evaluation}
 
@@
 
-\section{Discussion of Results}
+For the evaluation of the system, we used the standard measure of \emph{precision} and \emph{recall} as shown below.
 
-\section{Future Work \& Conclusion}
+\begin{align*}
+     precision =  \frac{  |\,gold\,\cap\,predicted\,| }{ |\,predicted\,| } \\
+     recall = \frac{ |\,gold\,\cap\,predicted\,| }{ |\,gold\,| }
+\end{align*}
+
+Precision is the percentage of Named Entities found by the system that are correct.
+Recall is the percentage of Named Entities present in the corpus that were found by the system.
+A Named Entity is only deemed correct if it is an exact match for the corresponding data file.
+The general formula for the \emph{F-Score} is shown below.
+Since we consider \emph{precision} and \emph{recall} as equally important, we use the $F_{1}$ score, which is the harmonic mean of precision and recall.
+
+\begin{align*}
+  F_{\beta} &=  (1+\beta^2)*\frac{precision *recall}{\beta^2* precision + recall}\\
+  F_1       &= 2*\frac{precision *recall}{precision + recall}
+\end{align*}
+
+\section{Discussion}
+
+\section{Future Work and Conclusion}
 

### 2012-11-10 doc/report/paper.tex
+++

--- 2012-11-10 doc/report/paper.tex
@@
 \usepackage{textcomp}
+\usepackage{placeins}
+%\usepackage{float}
 
@@
 \date{}
+%\restylefloat{figure} 
+%\restylefloat{table}
 
@@
 \begin{abstract}
-Named Entity Recognition is a task
+Named Entity Recognition is a fundamental task for a wide variety of
+natural language processing applications, such as
+question answering, information retrieval.
 
@@
 This approach, however, is rather high maintenance and requires extensive work by experts, such as computational linguists \cite{nadeau2007survey}.
-The results can be quite high in precision, but lack considerably in recall. % rules are partly language-dependent
+The results are quite reasonable, but not readily portable to other domains or languages. % rules are partly language-dependent
 
@@
 
-\begin{table}[h!]
+\begin{table}[!h]
 \begin{tabular}{| l | l |}
@@
 
+%\FloatBarrier 
 \section{Experiments and Evaluation}
+\begin{table*}[t!]
+\scriptsize
+\begin{tabular}{| l | l l l| l l l |}
+\hline
+\bf Language & \multicolumn{3}{c|}{ \bf Test A}&\multicolumn{3}{c|}{ \bf Test B}\\
+             & Precision & Recall & $F_1$ & Precision & Recall & $F_1$ \\ \hline
+Spanish &       &          &     &          &               & \\
+Dutch  &         &          &     &          &               &   \\
+English &        &          &     &          &               &       \\
+German &      &          &       &          &             & \\
+\hline
+\end{tabular}
+\caption{NER Structured Prediction Results }
+\label{table:Results}
+\end{table*}
 In the following section, we present our experiments and the evaluation of our system.
@@
 \subsection*{Evaluation}
+
 For the evaluation of the system we used \emph{Precision} and \emph{Recall} as shown in \ref{Precision} and \ref{Recall} respectively.
@@
 
-\begin{table*}[t]
-\centering
-\begin{tabular}{| l | l l l| l l l |}
-
-\hline
-\bf Language & \multicolumn{3}{c|}{ \bf Test A}&\multicolumn{3}{c|}{ \bf Test B}\\
-             & Precision & Recall & $F_1$ & Precision & Recall & $F_1$ \\ \hline
-Spanish &       &          &     &          &               & \\
-Dutch  &         &          &     &          &               &   \\
-English &        &          &     &          &               &       \\
-German &      &          &       &          &             & \\
-\hline
-\end{tabular}
-\caption{NER Structured Prediction Results }
-\label{table:Results}
-\end{table*}
-
+\FloatBarrier
 \section{Discussion of Results}

### 2012-11-11 doc/report/paper.tex
Started Proof-reading

--- 2012-11-10 doc/report/paper.tex
@@
 \usepackage{textcomp}
+\usepackage{todonotes}
 
@@
 \begin{abstract}
-Named Entity Recognition is a task
+In this paper, we apply the method of structure prediction to the information extraction task Named Entity Recognition (NER).
+We illustrate a number issues relevant to the NER task, describe our implementation and compare it to state-of-the art methods.
+%We use a structured version of the Perceptron algorithm with averaging and Viterbi decoding.
 
@@
 that was focusing on the extraction of structured information from unstructured text, such as company names from newspaper articles \cite{nadeau2007survey}.
-The most prominent types are generally known under the name of \emph{enamex} types, which are comprised of \texttt{Person Names}, 
-\texttt{Organisations} and \texttt{Locations}. 
-An additional type \texttt{Miscellaneous} captures names outside the classic \emph{enamax} type.
-Apart from the above, there is also the \emph{timex} type, which covers date and time expressions and \emph{numex} for monetary values and percentages. 
-Table \ref{table:NETypes} gives an overview of the different Named Entity (NE) types.
+The most prominent types of named entities are generally known under the name ENAMEX and are comprised of \texttt{Person Names}, \texttt{Organisations} and \texttt{Locations}.
+An additional type \texttt{Miscellaneous} captures names outside the classic ENAMEX types.
+Apart from the above, there is also the TIMEX type, which covers date and time expressions and \emph{NUMEX} for monetary values and percentages.
+Table~\ref{table:NETypes} provides an overview of the different Named Entity (NE) types.
 
@@
 \end{tabular}
-\caption{NE Examples}
+\caption{Named Entity Examples}
 \label{table:NETypes}
@@
 The difference lies in the reference of the entity. A Named Entity refers to a specific, unique person/time/location, whereas
-\emph{normal} entites can refer to multiple events, such as the time expression: \emph{In June}, which could refer to any year.
+\emph{normal} entities can refer to multiple events, such as the time expression: \emph{In June}, which could refer to any year.
 Also, there could be a reference to a person, as in: \emph{the prof}, which in itself is not a \namedentity.
@@
 A Named Entity  can consist of more than one syntactic type. Syntactic types feature in syntactic rules that define how sentences can be build.
-Typical high-level rules include: \\
-$s \rightarrow np\quad vp $ \footnote{ A sentence s breaks down into a noun phrase np and a verb phrase vp} \\
-$ vp \rightarrow v\quad (np)$ \footnote{ A verb phrase vp consists of a verb followed by an optional np} \\
+Typical high-level rules include:
+
+\ex. $\text{S} \rightarrow \text{NP}\  \text{VP} $ \footnote{ A sentence S breaks down into a noun phrase NP and a verb phrase VP} \\
+$ \text{VP} \rightarrow \text{V}\  (\text{NP})$ \footnote{ A verb phrase VP consists of a verb followed by an optional NP}
 
 Thus, sentences are often represented as syntactic trees that model the dependencies among the types.
-The overall type for a Named Entity is usually a noun phrase (np), which can be both simple and complex, meaning it can consist of a only a noun phrase or 
-also additional other syntactic types, such as prepositional phrases (pp), that consist of a preposition and a noun phrase. 
+The overall type for a Named Entity is usually a noun phrase (NP), which can be both simple and complex, meaning it can consist of a only a noun phrase or 
+also additional other syntactic types, such as prepositional phrases (PP), that consist of a preposition and a noun phrase. 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% - can be left out if not enough space/too detailed
@@
 
-\ex. $ np \rightarrow np \quad pp$ \\
-     $ pp \rightarrow p \quad np$ 
+\ex. $ \text{NP} \rightarrow \text{NP} \  \text{PP}$ \\
+     $ \text{PP} \rightarrow \text{P} \  \text{NP}$ 
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-We distinguish between pp-complements and pp-modifiers; where complements are non-optional elements of the noun phrase as they are
-inherent to the meaning of the entity and modifiers simply add to the description and are optional. 
+% \ex. The Bank of Scotland is in the brown building with the brown bricks on its roof... \label{PPrec}
+% 
+% The PP \emph{of Scotland} is semantically defining for the noun \emph{bank}, whereas \emph{with the brown bricks on its roof} 
+% just adds to the description of \emph{brown building}.
+
+\subsection{Issues in Named Entity Recognition}
+
+To illustrate linguistically motivated problems in NER, we consider the following problem, which is addressed in our model by a feature.
+
+We distinguish between PP-complements and PP-modifiers. Complements are non-optional elements of the noun phrase as they are
+inherent to the meaning of the entity. Modifiers add to the description and are optional.
 The syntactic trees in \ref{fig:bos} and \ref{fig:nirvana} show two examples for a complex noun phrase containing a \namedentity.
 \emph{N'} signals an intermediate state in the syntactic tree where modifiers can be joined.
-In \ref{fig:bos}, (\emph{Bank of Scotland}), the pp \emph{of Scotland} is attached directly; it is central to the meaning, 
+In \ref{fig:bos}, (\emph{Bank of Scotland}), the PP \emph{of Scotland} is attached directly; it is central to the meaning, 
 whereas in \ref{fig:nirvana} both \emph{Kurt Cobain} and \emph{Nirvana} constitute own concepts and are joined at a later level to express this distance.
-Thus, in connection with Named Entities, we want to be able to capture types of the first example as a \namedentity, but not of the second more distantly joined 
-relation.
+Thus, in Named Entity Recognition, we want to be able to capture types of the first example as a single \namedentity~(\emph{Bank of Scotland}), but treat the second example as two separate Named Entities (\emph{Kurt Cobain} and \emph{Nirvana}).
+
 % \begin{figure}
@@
 
-% \ex. The Bank of Scotland is in the brown building with the brown bricks on its roof... \label{PPrec}
-% 
-% The PP \emph{of Scotland} is semantically defining for the noun \emph{bank}, whereas \emph{with the brown bricks on its roof} 
-% just adds to the description of \emph{brown building}.
-
-\subsection{Issues in Named Entity Recognition}
-Among the most common challenges for NER is the issue of ambiguity, in particular \emph{Polysemy} and \emph{Metonymy} \cite{nadeau2007survey}.
+Generally, one of the most common challenges for NER is the issue of ambiguity, in particular \emph{Polysemy} and \emph{Metonymy} \cite{nadeau2007survey}.
 Polysemy refers to the property of a word that is a lexical representation having more than one possible meaning.
-\emph{Polysemy} could become an issue for NER when this lexical representation points to two different Named Entity types.
-This is quite frequent with \texttt{Person Names} and \texttt{Locations}, since many people are named after cities, such as \emph{Paris} or \emph{Georgia}. 
-Often the context will not be disambiguating as in example~\ref{ex:Georgia}.
+Polysemy can become an issue for NER when this lexical representation points to two different Named Entity types.
+This is quite frequent with \texttt{Person Names} and \texttt{Locations}, since many people are named after cities and vice versa, for example \emph{Paris} or \emph{Georgia}. 
+Often, context is not sufficient to disambiguate such names as in example~\ref{ex:Georgia}.
 
@@
 
-\emph{Metonymy} is the relationship of a part-whole relation between two items.
-It is frequently employed in literary texts and news data (which is often used in NER), where one item is substituted for another. 
-Example \ref{ex:Lon} shows an instance of a \emph{whole-part}, where \emph{London} probably stands for the smaller 'item': \textbf{Goverment in London}.
+% So lange ich nicht sicher weiß, wie damit in den CONLL Annotationen umgegangen wird, find ich das ein bisschen gefährlich...
+%\emph{Metonymy} is the relationship of a part-whole relation between two items.
+%It is frequently employed in literary texts and news data (which is often used in NER), where one item is substituted for another. 
+%Example \ref{ex:Lon} shows an instance of a \emph{whole-part}, where \emph{London} probably stands for the smaller 'item': \textbf{Goverment in London}.
 
-\ex. \emph{\textbf{London} decided to increase the 1200 military personnel involved in Olympic security.} \label{ex:Lon}
+%\ex. \emph{\textbf{London} decided to increase the 1200 military personnel involved in Olympic security.} %\label{ex:Lon}
 
-\subsection{Methods employed for NER}
-For NER, there exist two general approaches: one is rule-based and the other statistical. 
+\subsection{Approaches to NER}
+
+There exist two general approaches: one is rule-based and the other statistical. 
 Rule-based methods make use of the underlying rules governing a language to extract Named Entities. 
-This approach, however, is rather high maintenance and requires extensive work by experts, such as computational linguists \cite{nadeau2007survey}.
-The results can be quite high in precision, but lack considerably in recall. % rules are partly language-dependent
+This approach, however, is high in maintenance and requires extensive work by experts, such as computational linguists \cite{nadeau2007survey}.
+The results can have high precision, but generally lack considerably in recall. % rules are partly language-dependent
 
-For statistical approaches, supervised-learning is the most common method applied in NER. 
-There are unsupervised methods, but they cannot compete with SL applications yet.
-For the shared task in 2002 \cite{tksintro}, 12 systems participated in the challenge and presented a wide variety of approaches to the Named Entity task.
-The most successful system in the task, used AdaBoost applied to fixed-depth decision trees. Special attention was given to a variety of input features and the features were also
-partly adapted for each language with an F-score of 81.39 on the Spanish test set and 77.05 on the Dutch one.
-The second highest system \cite{Florian:2002:NER:1118853.1118863} employed three stacked
-learners (transformation-based learning for obtaining base-level non-typed Named Entites, Snow for improving the quality of these entities and the forward-backward algorithm for
-finding categories for the Named Entites), where the combination of these three algorithms substantially outperformed any subset of them, with an F-score of 79.05 on the Spanish test
-set and 74.99 on the Dutch one.
+For statistical approaches, supervised learning is the most common method applied in NER.
+There are unsupervised methods, however they cannot currently compete with supervised applications.
+For the CoNLL shared task in 2002 \cite{tksintro}, 12 systems participated in the challenge and presented a wide variety of approaches to the Named Entity task.
+The most successful system used AdaBoost applied to fixed-depth decision trees. Special attention was given to a variety of input features and the features were also
+partly adapted for each language with an F-score of 81.39 on the Spanish test set and 77.05 on Dutch.
+The second best system \cite{Florian:2002:NER:1118853.1118863} employed three stacked
+learners (transformation-based learning for obtaining base-level non-typed Named Entites, Snow for improving the quality of these entities and the forward-backward algorithm for finding the categories for the Named Entites), where the combination of these three algorithms substantially outperformed any subset of them, with an F-score of 79.05 on the Spanish test set and 74.99 on Dutch.
 Among the most prominent algorithms in NER are maximum entropy models, whose benefit in regard to this task became most apparent in CoNLL 2003. 
-For CoNLL 2003, five of the sixteen systems used this statistical learning method and three of of those used maximum entropy models in isolation.
-Maximum entropy models seem a very suitable choice for the NER task, since systems using this approach were ranked among the three highest results in 
+For CoNLL 2003, five of the sixteen systems used this statistical learning method and three used maximum entropy models in isolation.
+Maximum entropy models appear to be a suitable choice for the NER task, since systems using this approach were ranked among the three highest results in 
 English and the two highest results in German \cite{TjongKimSang:2003:ICS:1119176.1119195}.
 The best system in both German and English \cite{Florian:2003:NER:1119176.1119201} used a combination of classifiers: a robust linear classifier, 
-a maximum entropy, a transformation-based learning, and a hidden Markov model which are connected through a linear interpolation of the classifiers’ % combination scheme
+a maximum entropy classifier, a transformation-based learner, and a hidden Markov model. The models were connected through a linear interpolation of the classifiers’ % combination scheme
 class probability distribution. Particular importance is also assigned to creating a rich feature space, also including a gazetteer feature. 
-This system gains a precision of 88.99\% on the English test set and  83.87\% on the German one.
+This system achieves a precision of 88.99\% on the English test set and 83.87\% German.
 
 \subsection{CoNLL Data and Baseline Performance}
-The training and test data for Spanish \& Dutch was taken from the CoNLL Shared Task 2002 \cite{tksintro} and 
-the CoNLL Shared Task 2003 \cite{TjongKimSang:2003:ICS:1119176.1119195} provided the one for English \& German.
-The ConNLL shared tasks are organised challenges, where the organisers propose a task and provide the participants with the
-training and test data. 
-In 2002 and 2003, the task was Named Entity Recognition. 
-In both years, the baseline rate for the individual languages was produced by a system, that had a unique class in the training data. 
-In case that a phrase was part of more than one entity, the system would choose the longest one \cite{TjongKimSang:2003:ICS:1119176.1119195}. 
-Table \ref{table:Base} shows the baseline for the four different languages.  
-The data is annotated with part-of-speech tags and named-entity tags, as well as BIO-tags that indicate the position in the Named Entity. %put in ``tokenized, chunked''
+The training and test data for Spanish and Dutch was taken from the CoNLL Shared Task 2002 \cite{tksintro} and 
+the CoNLL Shared Task 2003 \cite{TjongKimSang:2003:ICS:1119176.1119195} provided the data for English and German.
+%The ConNLL shared tasks are organized challenges, where the organisers propose a task and provide the participants with the
+%training and test data. 
+%In 2002 and 2003, the task was Named Entity Recognition. 
+In 2002 and 2003 CoNLL shared tasks on Named Entity Recognition, the baseline rate for the individual languages was produced by a system, that had a unique class in the training data. 
+In case that a phrase was part of more than one entity, the system would choose the longest one \cite{TjongKimSang:2003:ICS:1119176.1119195}.\todo{? hab ich nicht kapiert :D}
+Table~\ref{table:Base} shows the baselines for the four different languages.  
+The data is annotated with part-of-speech tags and named-entity tags, as well as BIO-tags that indicate the position of a token in the Named Entity. %put in ``tokenized, chunked''
 %Our approach is based on \cite{strlearn}, of the group that won the challenge for Spanish and Dutch in 2002.
@@
 \section{Implementation}
-In this section, we describe our implementation, the learning and decoding algorithm and the features.
+In this section, we describe our implementation, the learning and decoding algorithm and the features we used.
 
@@
 For the evaluation of the system, we used the standard measure of \emph{precision} and \emph{recall} as shown below.
-
 \begin{align*}
@@
 Since we consider \emph{precision} and \emph{recall} as equally important, we use the $F_{1}$ score, which is the harmonic mean of precision and recall.
-
 \begin{align*}
   F_{\beta} &=  (1+\beta^2)*\frac{precision *recall}{\beta^2* precision + recall}\\
-  F_1       &= 2*\frac{precision *recall}{precision + recall}
+  F_1       &= 2*\frac{precision *recall}{precision + recall}\\
 \end{align*}

### 2012-11-11 doc/report/paper.tex
todos.

@@
 class probability distribution. Particular importance is also assigned to creating a rich feature space, also including a gazetteer feature. 
-This system achieves a precision of 88.99\% on the English test set and 83.87\% German.
+\todo{Ich würde hier eher den F-Score nennen.}This system achieves a precision of 88.99\% on the English test set and 83.87\% German.
 
@@
 \begin{table}[h!]
-\scriptsize
+\centering
+\small
 \begin{tabular}{|l|l|l|l|}
@@
 \end{tabular}
-\caption{NER Baseline}
+\caption{Named Entity Recognition baselines}
 \label{table:Base}
@@
 As a first step, the trellis has to be constructed, then we have to find the $ a \in \mathcal{Y}$ in the 
-last column of the trellis with maximal score $\delta_n(a)$. From this, the sequence can be recovered via back-tracking trough the trellis. 
+last column of the trellis with maximal score $\delta_n(a)$. From this, the sequence can be recovered via back-tracking through the trellis. 
 The score of the best sequence ending in $a$ is:

### 2012-11-11 doc/report/paper.tex

--- 2012-11-11 doc/report/paper.tex
@@
 \begin{abstract}
-In this paper, we apply the method of structure prediction to the information extraction task Named Entity Recognition (NER).
+In this paper, we apply the method of structure prediction to the information extraction task: Named Entity Recognition (NER). 
 We illustrate a number issues relevant to the NER task, describe our implementation and compare it to state-of-the art methods.
 %We use a structured version of the Perceptron algorithm with averaging and Viterbi decoding.
-
 \end{abstract}
@@
 
-Named Entity Recognition (NER) is an information extraction task and has first been introduced as part of the 6th MUC (Sixth Message Understanding Conference)
+Named Entity Recognition (NER) is a fundamental task for wide variety of natural language processing applications, including information retrieval and question answering
+and thus provides a general basis for semantic analysis of texts \cite{Chen:2010:UDB:1870457.1870473}.
+
+NER is an information extraction task and has first been introduced as part of the 6th MUC (Sixth Message Understanding Conference)
 that was focusing on the extraction of structured information from unstructured text, such as company names from newspaper articles \cite{nadeau2007survey}.
-The most prominent types of named entities are generally known under the name ENAMEX and are comprised of \texttt{Person Names}, \texttt{Organisations} and \texttt{Locations}.
+The most prominent types of Named Entities are generally known under the name ENAMEX and are comprised of \texttt{Person Names}, \texttt{Organisations} and \texttt{Locations}.
 An additional type \texttt{Miscellaneous} captures names outside the classic ENAMEX types.
-Apart from the above, there is also the TIMEX type, which covers date and time expressions and \emph{NUMEX} for monetary values and percentages.
+Apart from the above, there is also the TIMEX type, which covers date and time expressions and NUMEX for monetary values and percentages.
 Table~\ref{table:NETypes} provides an overview of the different Named Entity (NE) types.
@@
 \begin{table}[h!]
-\scriptsize
-\begin{tabular}{| l | l |}
-\hline
+\small
+\begin{tabular}{ l | l }
+
 \bf Named Entity Type & \bf Example \\
@@
 Percent &   30\% \\
-\hline
 \end{tabular}
@@
 
-There exist two general approaches: one is rule-based and the other statistical. 
-Rule-based methods make use of the underlying rules governing a language to extract Named Entities. 
-This approach, however, is high in maintenance and requires extensive work by experts, such as computational linguists \cite{nadeau2007survey}.
-The results can have high precision, but generally lack considerably in recall. % rules are partly language-dependent
+Initially, rule-based methods to extract Named Entities were more common. These make use of the underlying rules governing a language.
+This approach, however, is very high in maintenance and requires extensive work by experts, such as computational linguists \cite{nadeau2007survey}.
+The results achieved were reasonably good, but presents some flaws since constructed rules depend on the language and  the domain and are not readily portable. 
+This is the reason, why nowadays machine learning methods are more attractive for this task as they offer more and faster adaptability.
 
-For statistical approaches, supervised learning is the most common method applied in NER.
+For ML, supervised learning is the most common method applied in NER.
 There are unsupervised methods, however they cannot currently compete with supervised applications.
@@
 This system achieves a precision of 88.99\% on the English test set and 83.87\% German.
+Another, more recent approach involved the use of Deep Belief Nets for NER in Chinese \cite{Chen:2010:UDB:1870457.1870473}.
 
@@
 \begin{table}[h!]
-\begin{tabular}{| l | l |}
-\hline
+\begin{tabular}{ l | l }
 \bf Feature & \bf Example \\
@@
 Lemma &  produced $\Rightarrow$ produce \\
-\hline
+
 \end{tabular}

### 2012-11-11 doc/report/paper.tex
Tables, stuff.

@@
 \begin{abstract}
-In this paper, we apply the method of structure prediction to the information extraction task: Named Entity Recognition (NER). 
-We illustrate a number issues relevant to the NER task, describe our implementation and compare it to state-of-the art methods.
+In this paper, we apply the method of structure prediction to the Named Entity Recognition (NER) information extraction task. 
+We illustrate a number of issues relevant to the NER task, describe our implementation and compare it to state-of-the art methods.
 %We use a structured version of the Perceptron algorithm with averaging and Viterbi decoding.
@@
 \begin{table}[h!]
-\small
-\begin{tabular}{ l | l }
+\scriptsize
+\begin{tabular}{ l l }
 
@@
 \begin{table}[h!]
-\scriptsize
+\small
 \begin{tabular}{|l|l|l|l|}
@@
 \end{tabular}
-\caption{NER Baseline}
+\caption{Named Entity Recognition baselines.}
 \label{table:Base}
@@
 \begin{table}[h!]
-\begin{tabular}{ l | l }
+\small
+\begin{tabular}{ l l }
 \bf Feature & \bf Example \\

### 2012-11-11 doc/report/paper.tex
+++

--- 2012-11-11 doc/report/paper.tex
@@
 This approach, however, is very high in maintenance and requires extensive work by experts, such as computational linguists \cite{nadeau2007survey}.
-The results achieved were reasonably good, but presents some flaws since constructed rules depend on the language and  the domain and are not readily portable. 
+The results achieved were reasonably good, but presented some flaws since constructed rules depend on the language and  the domain and are not readily portable. 
 This is the reason, why nowadays machine learning methods are more attractive for this task as they offer more and faster adaptability.
@@
 class probability distribution. Particular importance is also assigned to creating a rich feature space, also including a gazetteer feature. 
-This system achieves a precision of 88.99\% on the English test set and 83.87\% German.
-Another, more recent approach involved the use of Deep Belief Nets for NER in Chinese \cite{Chen:2010:UDB:1870457.1870473}.
+This system achieves a F-score of 88.76\% on the English test set and 72.41\% on the German one.
+Another, more recent approach involved the use of Deep Belief Nets (DBN) for NER in Chinese \cite{Chen:2010:UDB:1870457.1870473}.
+Due to the special characteristics of Chinese of lacking boundary information, character-based methods are generally given preference over word-based methods.
+In their application for NER, DBN  is used to automatically discover the complicated composite effects of the characters to the NE categories from the input data. 
+The paper reports DBN's strong representative power in regard to detection of more complicated and efficient features. With a precision of 91.45\%,
+it achieves better performance than a SVM and a BP neural networks.
 

### 2012-11-11 doc/report/paper.tex
Minor changes, experiment.

@@
       \textbf{y:}  O   O      O     0 ORG  O  O     O     O       O        O PER   PER  O  \label{PredEx1b} \\
+
 \caption{Input and predicted structure for the Named Entity Recognition task.}
+%\label{fig:predxy}
 
@@
 
-As the predicted structure, we use the labels for every token in the full sentence. If a token is not a \namedentity, it is assigned the label \Oo.
+As the predicted structure, we use the labels for every token in the full sentence (see Figure~3). If a token is not a \namedentity, it is assigned the label \Oo.
 
@@
 For finding the best sequence of labels for a sentence, we use a modified version of the Viterbi algorithm. The algorithm finds the sequence $\hat{y}$, such that:
-
 \[
@@
 The score of the best sequence ending in $a$ is:
-
 \[
@@
 Headlines, mainly in the English data, are often written entirely in uppercase letters.
-This is problematic because from most of the data, the model might learn that a word in all uppercase letters is indicative of an organization (e.g.\ IBM, AIG, NYSE) and might thus mistake the tokens in headlines for organizations.
+This is problematic because from the data, the model might learn that a word in all uppercase letters is indicative of an organization (e.g.\ IBM, AIG, NYSE) and might thus mistake the tokens in headlines for organizations.
 We address this issue by converting all-uppercase headlines into their most likely original case (truecasing).
@@
 \subsubsection{Extending monolingual gazetteers}
+
+The experiment above has shown that the gazetteer features are beneficial for Dutch.
+In this experiment, we want to explore whether the gazetteer features can be improved by extending the monolingual gazetteer with entries from other languages.
+For this, we compare the Dutch-only gazetteer with a gazetteer extended with English data.
 
@@
 
-This paper presented a system for \namedentity Recognition using Structured Prediction.
+In this paper, we presented a \namedentity~Recognition system based on the concept of Structured Prediction.
+While the system does not achieve state-of-the-art performance on the CoNLL shared task, we believe that the results are res
 

### 2012-11-11 doc/report/paper.tex
+conc.

@@
 In this paper, we presented a \namedentity~Recognition system based on the concept of Structured Prediction.
-While the system does not achieve state-of-the-art performance on the CoNLL shared task, we believe that the results are res
+While the system does not achieve state-of-the-art performance on the CoNLL shared task, we believe that the results are reasonable for the following reasons.
+We are using a simple model, which has the advantage that the final application will be fast and it is easily extendable.
+In contrast to the best systems in the CoNLL task, we also do not use ensemble learning methods and we use the same set of features for all languages, which enables simple extension of the system to additional languages.
 

### 2012-11-11 doc/report/paper.tex
+++

--- 2012-11-11 doc/report/paper.tex
@@
 \begin{abstract}
-In this paper, we apply the method of structure prediction to the Named Entity Recognition (NER) information extraction task. 
+In this paper, we apply the method of Structured Prediction to the Named Entity Recognition task. %(NER) information extraction task. 
 We illustrate a number of issues relevant to the NER task, describe our implementation and compare it to state-of-the art methods.
@@
 
-Named Entity Recognition (NER) is a fundamental task for wide variety of natural language processing applications, including information retrieval and question answering
+Named Entity Recognition (NER) is a fundamental task for a wide variety of natural language processing applications, including information retrieval and question answering
 and thus provides a general basis for semantic analysis of texts \cite{Chen:2010:UDB:1870457.1870473}.
@@
 that was focusing on the extraction of structured information from unstructured text, such as company names from newspaper articles \cite{nadeau2007survey}.
-The most prominent types of Named Entities are generally known under the name ENAMEX and are comprised of \texttt{Person Names}, \texttt{Organisations} and \texttt{Locations}.
-An additional type \texttt{Miscellaneous} captures names outside the classic ENAMEX types.
-Apart from the above, there is also the TIMEX type, which covers date and time expressions and NUMEX for monetary values and percentages.
+The most prominent types of Named Entities are generally known under the name of ENAMEX and are comprised of \texttt{Person Names}, \texttt{Organisations} and \texttt{Locations}.
+An additional type: \texttt{Miscellaneous} captures names outside the classic ENAMEX type.
+Apart from the above, there is also the TIMEX type, which covers date and time expressions and the NUMEX type for monetary values and percentages.
 Table~\ref{table:NETypes} provides an overview of the different Named Entity (NE) types.
@@
 \begin{table}[h!]
-\scriptsize
+\small
 \begin{tabular}{ l l }
-
 \bf Named Entity Type & \bf Example \\
@@
 In considering Named Entities, it is important not to confuse them with non-specific entities. % reformulate
-The difference lies in the reference of the entity. A Named Entity refers to a specific, unique person/time/location, whereas
-\emph{normal} entities can refer to multiple events, such as the time expression: \emph{In June}, which could refer to any year.
-Also, there could be a reference to a person, as in: \emph{the prof}, which in itself is not a \namedentity.
+The difference lies in the reference of the entity. A Named Entity refers to a specific, unique person/time/location..., whereas
+\emph{normal} entities can refer to multiple events, such as the time expression: \emph{In June}, which is ambiguous in regard to the year.
+%Also, there could be a reference to a person, as in: \emph{the prof}, which in itself is not a \namedentity.
 %However, it would be a deictic reference and point back to the mention of a real \namedentity, as in the following coreference chain: \\
@@
 
-\ex. $\text{S} \rightarrow \text{NP}\  \text{VP} $ \footnote{ A sentence S breaks down into a noun phrase NP and a verb phrase VP} \\
-$ \text{VP} \rightarrow \text{V}\  (\text{NP})$ \footnote{ A verb phrase VP consists of a verb followed by an optional NP}
+\ex. $\text{S} \rightarrow \text{NP}\  \text{VP} $ \footnote{ A sentence (S) breaks down into a noun phrase (NP) and a verb phrase (VP)} \\
+$ \text{VP} \rightarrow \text{V}\  (\text{NP})$ \footnote{ A verb phrase (VP) consists of a verb followed by an optional NP}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% - can be left out if not enough space/too detailed
-It is therefore a recursive definition and in theory, noun phrases can grow infinitely big, as shown by the following recursive rules:  %
+It is therefore a recursive definition and in theory, noun phrases can grow infinitely big, as shown by the recursive rules in example~\ref{recRules}.  %
 
-\ex. $ \text{NP} \rightarrow \text{NP} \  \text{PP}$ \\
+\ex. $ \text{NP} \rightarrow \text{NP} \  \text{PP}$ \\ \label{recRules}
      $ \text{PP} \rightarrow \text{P} \  \text{NP}$ 
@@
 
-To illustrate linguistically motivated problems in NER, we consider the following problem, which is addressed in our model by a feature.
+In order to illustrate linguistically motivated problems in NER, we consider the following problem, which is addressed in our model by a feature.
 
@@
 whereas in \ref{fig:nirvana} both \emph{Kurt Cobain} and \emph{Nirvana} constitute own concepts and are joined at a later level to express this distance.
-Thus, in Named Entity Recognition, we want to be able to capture types of the first example as a single \namedentity~(\emph{Bank of Scotland}), but treat the second example as two separate Named Entities (\emph{Kurt Cobain} and \emph{Nirvana}).
+Thus, in Named Entity Recognition, we want to be able to capture types of the first example as a single \namedentity~(\emph{Bank of Scotland}), 
+but treat the second example as two separate Named Entities (\emph{Kurt Cobain} and \emph{Nirvana}).
 
@@
 The second best system \cite{Florian:2002:NER:1118853.1118863} employed three stacked
-learners (transformation-based learning for obtaining base-level non-typed Named Entites, Snow for improving the quality of these entities and the forward-backward algorithm for finding the categories for the Named Entites), where the combination of these three algorithms substantially outperformed any subset of them, with an F-score of 79.05 on the Spanish test set and 74.99 on Dutch.
+learners (transformation-based learning for obtaining base-level non-typed Named Entites, sparse network of winnows (Snow) for improving the quality of these entities and the forward-backward algorithm for finding the categories for the Named Entites), where the combination of these three algorithms substantially outperformed any subset of them, with an F-score of 79.05 on the Spanish test set and 74.99 on Dutch.
 Among the most prominent algorithms in NER are maximum entropy models, whose benefit in regard to this task became most apparent in CoNLL 2003. 
@@
 \exg. \textbf{x:} The motor company ' Ford ' was founded and incorporated by Henry Ford .\\
-      \textbf{y:}  O   O      O     0 ORG  O  O     O     O       O        O PER   PER  O  \label{PredEx1b} \\
+      \textbf{y:}  0   O      O     0 ORG  O  O     O     O       O        O PER   PER  O  \label{PredEx1b} \\
 \caption{Input and predicted structure for the Named Entity Recognition task.}
@@
 These features register which label has been assigned to the previous token and takes into account the most likely sequence. 
-Thus, for the sequence: ``Jack London went to New York'' the NE tag combination of \emph{PER} and \emph{PER} in example~\ref{seq1} is more likely than\emph{PER} and \emph{LOC} in \ref{seq2}.
+Thus, for the sequence: ``Jack London went to New York'' the NE tag combination of \emph{PER} and \emph{PER} in example~\ref{seq1} 
+is more likely than\emph{PER} and \emph{LOC} in \ref{seq2}.
 %current token and last label for prepositions or possessive ’s
@@
 
+In addition, there is a prepositional phrase (PP) attachment feature, that registers the existence of a this attachment, that would
+result in a break of the sequence as in
+
+\exg. Bank of Scotland . \\ 
+      ORG   0    ORG   0  \\\label{seq3}
+
 \paragraph*{Gazetteer Features}
 A technique commonly used in Named Entity Recognition is the use of external lists of names of a specific type (gazetteer). 
-In order to create gazetteer lists for more common named entities, we use a \emph{SPARQL} query  (Figure~\ref{fig:sparql}) to automatically retrieve entries from DBpedia for all required languages from a public SPARQL endpoint.
+In order to create gazetteer lists for more common named entities, we use a \emph{SPARQL} query  (Figure~\ref{fig:sparql}) 
+to automatically retrieve entries from DBpedia for all required languages from a public SPARQL endpoint.
 
@@
 Given the lists of entities retrieved, we then mark the longest matches of known entities in the input using a directed acyclic word graph. 
-The mention is annotated with a reference to its list of origin, which is then used as a feature in our model.
-With this approach, it is not necessary to manually map the lists to target labels and the model can learn the reliability of each list.
+The mention is annotated with a reference to its list of origin, which is then used as a feature in our model. With this approach, it is not 
+necessary to manually map the lists to target labels and the model can learn the reliability of each list.
 
@@
 
-One of the issues that we needed to address is the case of headlines.
-Headlines, mainly in the English data, are often written entirely in uppercase letters.
-This is problematic because from most of the data, the model might learn that a word in all uppercase letters is indicative of an organization (e.g.\ IBM, AIG, NYSE) and might thus mistake the tokens in headlines for organizations.
+One of the issues that we needed to address is the case of headlines. Headlines, mainly in the English data, are often written entirely 
+in uppercase letters. This is problematic because from most of the data, the model might learn that a word in all uppercase letters is 
+indicative of an organization (e.g.\ IBM, AIG, NYSE) and might thus mistake the tokens in headlines for organizations. 
 We address this issue by converting all-uppercase headlines into their most likely original case (truecasing). 
-To estimate the most likely case of a token from a monolingual corpus, we use the truecasing model from the Moses statistical machine translation system \cite{koehn2007moses}.
+To estimate the most likely case of a token from a monolingual corpus, we use the truecasing model from the 
+Moses statistical machine translation system \cite{koehn2007moses}.
 
@@
 
-Precision is the percentage of Named Entities found by the system that are correct.
-Recall is the percentage of Named Entities present in the corpus that were found by the system.
-A Named Entity is only deemed correct if it is an exact match for the corresponding data file.
-The general formula for the \emph{F-Score} is shown below.
-Since we consider \emph{precision} and \emph{recall} as equally important, we use the $F_{1}$ score, which is the harmonic mean of precision and recall.
+Precision is the percentage of Named Entities found by the system that are correct. Recall is the percentage of Named Entities present in the corpus that were 
+found by the system. A Named Entity is only deemed correct if it is an exact match for the corresponding data file. The general formula for the \emph{F-Score} 
+is shown below. Since we consider \emph{precision} and \emph{recall} as equally important, we use the $F_{1}$ score, which is the harmonic mean of precision and recall.
 \begin{align*}

### 2012-11-11 doc/report/paper.tex

@@
 \begin{abstract}
-In this paper, we apply the method of Structured Prediction to the Named Entity Recognition task. %(NER) information extraction task. 
+In this paper, we apply the method of structure prediction to the Named Entity Recognition (NER) information extraction task. 
 We illustrate a number of issues relevant to the NER task, describe our implementation and compare it to state-of-the art methods.
@@
 
-Named Entity Recognition (NER) is a fundamental task for a wide variety of natural language processing applications, including information retrieval and question answering
+Named Entity Recognition (NER) is a fundamental task for wide variety of natural language processing applications, including information retrieval and question answering
 and thus provides a general basis for semantic analysis of texts \cite{Chen:2010:UDB:1870457.1870473}.
@@
 that was focusing on the extraction of structured information from unstructured text, such as company names from newspaper articles \cite{nadeau2007survey}.
-The most prominent types of Named Entities are generally known under the name of ENAMEX and are comprised of \texttt{Person Names}, \texttt{Organisations} and \texttt{Locations}.
-An additional type: \texttt{Miscellaneous} captures names outside the classic ENAMEX type.
-Apart from the above, there is also the TIMEX type, which covers date and time expressions and the NUMEX type for monetary values and percentages.
+The most prominent types of Named Entities are generally known under the name ENAMEX and are comprised of \texttt{Person Names}, \texttt{Organisations} and \texttt{Locations}.
+An additional type \texttt{Miscellaneous} captures names outside the classic ENAMEX types.
+Apart from the above, there is also the TIMEX type, which covers date and time expressions and NUMEX for monetary values and percentages.
 Table~\ref{table:NETypes} provides an overview of the different Named Entity (NE) types.
@@
 \begin{table}[h!]
-\small
+\scriptsize
 \begin{tabular}{ l l }
+
 \bf Named Entity Type & \bf Example \\
@@
 In considering Named Entities, it is important not to confuse them with non-specific entities. % reformulate
-The difference lies in the reference of the entity. A Named Entity refers to a specific, unique person/time/location..., whereas
-\emph{normal} entities can refer to multiple events, such as the time expression: \emph{In June}, which is ambiguous in regard to the year.
-%Also, there could be a reference to a person, as in: \emph{the prof}, which in itself is not a \namedentity.
+The difference lies in the reference of the entity. A Named Entity refers to a specific, unique person/time/location, whereas
+\emph{normal} entities can refer to multiple events, such as the time expression: \emph{In June}, which could refer to any year.
+Also, there could be a reference to a person, as in: \emph{the prof}, which in itself is not a \namedentity.
 %However, it would be a deictic reference and point back to the mention of a real \namedentity, as in the following coreference chain: \\
@@
 \exg. \textbf{x:} The motor company ' Ford ' was founded and incorporated by Henry Ford .\\
-      \textbf{y:}  0   O      O     0 ORG  O  O     O     O       O        O PER   PER  O  \label{PredEx1b} \\
+      \textbf{y:}  O   O      O     0 ORG  O  O     O     O       O        O PER   PER  O  \label{PredEx1b} \\
 \caption{Input and predicted structure for the Named Entity Recognition task.}
@@
 These features register which label has been assigned to the previous token and takes into account the most likely sequence. 
-Thus, for the sequence: ``Jack London went to New York'' the NE tag combination of \emph{PER} and \emph{PER} in example~\ref{seq1} 
-is more likely than\emph{PER} and \emph{LOC} in \ref{seq2}.
+Thus, for the sequence: ``Jack London went to New York'' the NE tag combination of \emph{PER} and \emph{PER} in example~\ref{seq1} is more likely than\emph{PER} and \emph{LOC} in \ref{seq2}.
 %current token and last label for prepositions or possessive ’s

### 2012-11-11 doc/report/paper.tex
+++

@@
 \usepackage{todonotes}
+\usepackage{placeins}
 
@@
 In the following, we show the influence of each of the three sets of features on the resulting accuracy.
-
-%INSERT MATLAB PIC HERE
+\FloatBarrier
+\begin{figure*}[t!]
+ \includegraphics[scale=0.5]{Plot.png}
+\end{figure*}
 
@@
 
-This paper presented a system for \namedentity Recognition using Structured Prediction.
+In this paper, we presented a \namedentity~Recognition system based on the concept of Structured Prediction. 
+While the system does not achieve state-of-the-art performance on the CoNLL shared task, we believe that the results 
+are reasonable for the following reasons. We are using a simple model, which has the advantage that the final application 
+will be fast and it is easily extendable. In contrast to the best systems in the CoNLL task, we also do not use ensemble 
+learning methods and we use the same set of features for all languages, which enables simple extension of the system to additional languages.
 

### 2012-11-11 doc/report/paper.tex
Minor changes.

@@
 \usepackage{todonotes}
-\usepackage{placeins}
 
@@
 
-\ex. $\text{S} \rightarrow \text{NP}\  \text{VP} $ \footnote{ A sentence (S) breaks down into a noun phrase (NP) and a verb phrase (VP)} \\
-$ \text{VP} \rightarrow \text{V}\  (\text{NP})$ \footnote{ A verb phrase (VP) consists of a verb followed by an optional NP}
+\ex. $\text{S} \rightarrow \text{NP}\  \text{VP} $ \footnote{ A sentence S breaks down into a noun phrase NP and a verb phrase VP} \\
+$ \text{VP} \rightarrow \text{V}\  (\text{NP})$ \footnote{ A verb phrase VP consists of a verb followed by an optional NP}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% - can be left out if not enough space/too detailed
-It is therefore a recursive definition and in theory, noun phrases can grow infinitely big, as shown by the recursive rules in example~\ref{recRules}.  %
+It is therefore a recursive definition and in theory, noun phrases can grow infinitely big, as shown by the following recursive rules:  %
 
-\ex. $ \text{NP} \rightarrow \text{NP} \  \text{PP}$ \\ \label{recRules}
+\ex. $ \text{NP} \rightarrow \text{NP} \  \text{PP}$ \\
      $ \text{PP} \rightarrow \text{P} \  \text{NP}$ 
@@
 
-In order to illustrate linguistically motivated problems in NER, we consider the following problem, which is addressed in our model by a feature.
+To illustrate linguistically motivated problems in NER, we consider the following problem, which is addressed in our model by a feature.
 
@@
 whereas in \ref{fig:nirvana} both \emph{Kurt Cobain} and \emph{Nirvana} constitute own concepts and are joined at a later level to express this distance.
-Thus, in Named Entity Recognition, we want to be able to capture types of the first example as a single \namedentity~(\emph{Bank of Scotland}), 
-but treat the second example as two separate Named Entities (\emph{Kurt Cobain} and \emph{Nirvana}).
+Thus, in Named Entity Recognition, we want to be able to capture types of the first example as a single \namedentity~(\emph{Bank of Scotland}), but treat the second example as two separate Named Entities (\emph{Kurt Cobain} and \emph{Nirvana}).
 
@@
 The second best system \cite{Florian:2002:NER:1118853.1118863} employed three stacked
-learners (transformation-based learning for obtaining base-level non-typed Named Entites, sparse network of winnows (Snow) for improving the quality of these entities and the forward-backward algorithm for finding the categories for the Named Entites), where the combination of these three algorithms substantially outperformed any subset of them, with an F-score of 79.05 on the Spanish test set and 74.99 on Dutch.
+learners (transformation-based learning for obtaining base-level non-typed Named Entites, Snow for improving the quality of these entities and the forward-backward algorithm for finding the categories for the Named Entites), where the combination of these three algorithms substantially outperformed any subset of them, with an F-score of 79.05 on the Spanish test set and 74.99 on Dutch.
 Among the most prominent algorithms in NER are maximum entropy models, whose benefit in regard to this task became most apparent in CoNLL 2003. 
@@
 \hline
-Token &  \textbf{``Amsterdam''}\\
+Token &  \textbf{Amsterdam}\\
 Suffix& Amster\textbf{dam}\\
-Prefix&  \textbf{San} Sebastian\\
+Prefix&  \textbf{Mc}Cartney\\
 Captitalized& \textbf{B}enetton\\
-Number Pattern & \\
 UPPERCASE &  \textbf{BENETTON}\\
-POS-tag &  Benetton \textbf{NNP}   \\
+Number patterns & 2005\\
+POS tag &  Benetton \textbf{NNP}   \\
 Lemma & produced $\Rightarrow$ produce \\
@@
 \end{tabular}
-\caption{Node Features}
+
+\caption{\normalsize Examples of node features.}
 \label{table:node}
@@
 \paragraph*{Label Interaction Features}
-These features register which label has been assigned to the previous token and takes into account the most likely sequence. 
-Thus, for the sequence: ``Jack London went to New York'' the NE tag combination of \emph{PER} and \emph{PER} in example~\ref{seq1} is more likely than\emph{PER} and \emph{LOC} in \ref{seq2}.
+These features model the interaction between the labels of adjacent tokens and take into account the most likely sequence. 
+For the sequence ``Jack London went to New York'', the label combination of \emph{PER} and \emph{PER} as in \ref{seq1} is more likely than \emph{PER} and \emph{LOC} in \ref{seq2}.
 %current token and last label for prepositions or possessive ’s
@@
 
-In addition, there is a prepositional phrase (PP) attachment feature, that registers the existence of a this attachment, that would
-result in a break of the sequence as in
-
-\exg. Bank of Scotland . \\ 
-      ORG   0    ORG   0  \\\label{seq3}
-
 \paragraph*{Gazetteer Features}
 A technique commonly used in Named Entity Recognition is the use of external lists of names of a specific type (gazetteer).
-In order to create gazetteer lists for more common named entities, we use a \emph{SPARQL} query  (Figure~\ref{fig:sparql}) 
-to automatically retrieve entries from DBpedia for all required languages from a public SPARQL endpoint.
+In order to create gazetteer lists for more common named entities, we use a \emph{SPARQL} query  (Figure~\ref{fig:sparql}) to automatically retrieve entries from DBpedia for all required languages from a public SPARQL endpoint.
 
@@
 Given the lists of entities retrieved, we then mark the longest matches of known entities in the input using a directed acyclic word graph.
-The mention is annotated with a reference to its list of origin, which is then used as a feature in our model. With this approach, it is not 
-necessary to manually map the lists to target labels and the model can learn the reliability of each list.
+The mention is annotated with a reference to its list of origin, which is then used as a feature in our model.
+With this approach, it is not necessary to manually map the lists to target labels and the model can learn the reliability of each list.
 
@@
 
-One of the issues that we needed to address is the case of headlines. Headlines, mainly in the English data, are often written entirely 
-in uppercase letters. This is problematic because from most of the data, the model might learn that a word in all uppercase letters is 
-indicative of an organization (e.g.\ IBM, AIG, NYSE) and might thus mistake the tokens in headlines for organizations. 
+One of the issues that we needed to address is the case of headlines.
+Headlines, mainly in the English data, are often written entirely in uppercase letters.
+This is problematic because from the data, the model might learn that a word in all uppercase letters is indicative of an organization (e.g.\ IBM, AIG, NYSE) and might thus mistake the tokens in headlines for organizations.
 We address this issue by converting all-uppercase headlines into their most likely original case (truecasing).
-To estimate the most likely case of a token from a monolingual corpus, we use the truecasing model from the 
-Moses statistical machine translation system \cite{koehn2007moses}.
+To estimate the most likely case of a token from a monolingual corpus, we use the truecasing model from the Moses statistical machine translation system \cite{koehn2007moses}.
 
@@
 In the following, we show the influence of each of the three sets of features on the resulting accuracy.
-\FloatBarrier
-\begin{figure*}[t!]
- \includegraphics[scale=0.5]{Plot.png}
-\end{figure*}
+
+%INSERT MATLAB PIC HERE
 
 \subsubsection{Extending monolingual gazetteers}
+
+The experiment above has shown that the gazetteer features are beneficial for Dutch.
+In this experiment, we want to explore whether the gazetteer features can be improved by extending the monolingual gazetteer with entries from other languages.
+For this, we compare the Dutch-only gazetteer with a gazetteer extended with English data.
 
@@
 
-Precision is the percentage of Named Entities found by the system that are correct. Recall is the percentage of Named Entities present in the corpus that were 
-found by the system. A Named Entity is only deemed correct if it is an exact match for the corresponding data file. The general formula for the \emph{F-Score} 
-is shown below. Since we consider \emph{precision} and \emph{recall} as equally important, we use the $F_{1}$ score, which is the harmonic mean of precision and recall.
+Precision is the percentage of Named Entities found by the system that are correct.
+Recall is the percentage of Named Entities present in the corpus that were found by the system.
+A Named Entity is only deemed correct if it is an exact match for the corresponding data file.
+The general formula for the \emph{F-Score} is shown below.
+Since we consider \emph{precision} and \emph{recall} as equally important, we use the $F_{1}$ score, which is the harmonic mean of precision and recall.
 \begin{align*}
@@
 In this paper, we presented a \namedentity~Recognition system based on the concept of Structured Prediction.
-While the system does not achieve state-of-the-art performance on the CoNLL shared task, we believe that the results 
-are reasonable for the following reasons. We are using a simple model, which has the advantage that the final application 
-will be fast and it is easily extendable. In contrast to the best systems in the CoNLL task, we also do not use ensemble 
-learning methods and we use the same set of features for all languages, which enables simple extension of the system to additional languages.
+While the system does not achieve state-of-the-art performance on the CoNLL shared task, we believe that the results are reasonable for the following reasons: We are using a simple model, which has the advantage that the final application will be fast and it is easily extendable.
+In contrast to the best systems in the CoNLL task, we also do not use ensemble learning methods and we use the same set of features for all languages, which enables simple extension of the system to additional languages.
 

### 2012-11-11 doc/report/paper.tex
+++

--- 2012-11-11 doc/report/paper.tex
@@
 %In 2002 and 2003, the task was Named Entity Recognition. 
-In 2002 and 2003 CoNLL shared tasks on Named Entity Recognition, the baseline rate for the individual languages was produced by a system, that had a unique class in the training data. 
+In 2002 and 2003 CoNLL shared tasks were on Named Entity Recognition. The baseline system first decides on the most frequent Named Entity class C in the training data
+and then labels all occurrences of a Named Entity found in the test data with class C regardless of context. 
+Any named entities that did not occur in the training data are also not assigned any class in the test data.
 In case that a phrase was part of more than one entity, the system would choose the longest one \cite{TjongKimSang:2003:ICS:1119176.1119195}.\todo{? hab ich nicht kapiert :D}
@@
 \begin{figure*}[t!]
+\centering
  \includegraphics[scale=0.5]{Plot.png}

### 2012-11-11 doc/report/paper.tex
Changes.

--- 2012-11-11 doc/report/paper.tex
@@
 In the following, we show the influence of each of the three sets of features on the resulting accuracy.
-\FloatBarrier
-\begin{figure*}[t!]
- \includegraphics[scale=0.5]{Plot.png}
+\begin{figure*}[t]
+ \centering
+ \includegraphics[scale=0.4]{Plot.png}
 \end{figure*}
@@
 While the system does not achieve state-of-the-art performance on the CoNLL shared task, we believe that the results 
-are reasonable for the following reasons. We are using a simple model, which has the advantage that the final application 
+are reasonable for the following reasons: We are using a simple model, which has the advantage that the final application
 will be fast and it is easily extendable. In contrast to the best systems in the CoNLL task, we also do not use ensemble 

### 2012-11-11 doc/report/paper.tex

@@
 In the following, we show the influence of each of the three sets of features on the resulting accuracy.
-\begin{figure*}[t]
+\FloatBarrier
+\begin{figure*}[t!]
 \centering
- \includegraphics[scale=0.4]{Plot.png}
+ \includegraphics[scale=0.5]{Plot.png}
 \end{figure*}

### 2012-11-11 doc/report/paper.tex
.

@@
 
-\section{Structured Prediction}
-Prediction \cite{Taskar:2005:LSP:1102351.1102464} is a supervised-learning approach, where a system uses training data to learn the prediction
-of unseen data. It consequently maps an input \textbf{x} to an output \textbf{y}: \\
+\section{Implementation}
+In this section, we describe our implementation, the learning and decoding algorithm and the features we used.
 
-$ x \rightarrow y $. \\
+\subsection{Structured Prediction}
+Structured Prediction \cite{Taskar:2005:LSP:1102351.1102464} is an approach to supervised learning that is often used in natural language processing.
+Assuming our goal is to learn a function ${ \boldsymbol{f}\colon \mathcal{X}\to\mathcal{Y} }$, in structured prediction, we assume that $y \in \mathcal{Y}$ is not atomic (e.g.\ binary prediction), but that the prediction is a structure (e.g.\ a syntactic tree for a sentence or a sequence of part-of-speech tags).
 
-Structured Prediction and Non-structured prediction differ in the form of their output.
-Non-Structured Prediction returns an atomic output: it is binary prediction for a two-class problem and one of many labels for a multiclass problem. 
-Structured Prediction, on the other hand, returns the prediction for a whole sequence. 
-For the sentence in \ref{PredEx1a}, it would return the corresponding label combination in example~\ref{PredEx1b}.
+The structured prediction approach we follow has advantages over traditional approaches like hidden Markov models.
+While a hidden Markov model can model interactions between labels as transition probabilities, it is restricted to tokens and Maximum Likelihood learning.
+Another alternative would be to use a classifier for each token and then concatenate the results, however this approach would not take into account label interactions.
+With structured prediction, we can use features and model label interactions at the same time.
+
+\subsection{Learning and Decoding}
+
+% problem external data / use Dbpedia - gazetteer
+Our implementation for the languages English, German, Dutch and Spanish is trained and tested on the CoNLL 2003 and CoNLL 2002 data sets respectively.
 
@@
 \end{figure*}
-
-In comparison to other similar approaches, Structured Prediction sets itself apart through the combination of features and label interactions instead
-of concentrating mainly on label interactions like HMM or a rich set of features like local classifiers.
-
-\section{Implementation}
-In this section, we describe our implementation, the learning and decoding algorithm and the features we used.
-
-\subsection{Learning and Decoding}
-
-% problem external data / use Dbpedia - gazetteer
-Our implementation for the languages English, German, Dutch and Spanish is trained and tested on the CoNLL 2003 and CoNLL 2002 data sets respectively.
 

### 2012-11-11 doc/report/paper.tex
+++

@@
 We distinguish between PP-complements and PP-modifiers. Complements are non-optional elements of the noun phrase as they are
-inherent to the meaning of the entity. Modifiers add to the description and are optional.
-The syntactic trees in \ref{fig:bos} and \ref{fig:nirvana} show two examples for a complex noun phrase containing a \namedentity.
+inherent to the meaning of the entity. Modifiers only add to the description and are optional.
+The syntactic trees in \ref{fig:bos} and \ref{fig:nirvana} show two examples for a complex noun phrase containing Named Entites.
 \emph{N'} signals an intermediate state in the syntactic tree where modifiers can be joined.
@@
 
-Generally, one of the most common challenges for NER is the issue of ambiguity, in particular \emph{Polysemy} and \emph{Metonymy} \cite{nadeau2007survey}.
-Polysemy refers to the property of a word that is a lexical representation having more than one possible meaning.
-Polysemy can become an issue for NER when this lexical representation points to two different Named Entity types.
+Generally, one of the most common challenges for NER is the issue of ambiguity, in particular \emph{Polysemy} \cite{nadeau2007survey}. %and \emph{Metonymy} insert with other section
+Polysemy refers to the property of a lexical representation (\~ word) having more than one possible meaning. Polysemy can become an issue for NER when this lexical representation 
+points to two different Named Entity types.
 This is quite frequent with \texttt{Person Names} and \texttt{Locations}, since many people are named after cities and vice versa, for example \emph{Paris} or \emph{Georgia}. 
@@
 
-Initially, rule-based methods to extract Named Entities were more common. These make use of the underlying rules governing a language.
+Initially, rule-based methods were prevailing in the extraction of Named Entities. These methods make use of the underlying rules governing a language.
 This approach, however, is very high in maintenance and requires extensive work by experts, such as computational linguists \cite{nadeau2007survey}.
-The results achieved were reasonably good, but presented some flaws since constructed rules depend on the language and  the domain and are not readily portable. 
-This is the reason, why nowadays machine learning methods are more attractive for this task as they offer more and faster adaptability.
+The results achieved were reasonably good, but presented some flaws since constructed rules are dependent on the language and the domain and are not readily portable. 
+This is the reason why nowadays machine learning methods are more attractive for this task as they offer more adaptability.
 
 For ML, supervised learning is the most common method applied in NER.
-There are unsupervised methods, however they cannot currently compete with supervised applications.
-For the CoNLL shared task in 2002 \cite{tksintro}, 12 systems participated in the challenge and presented a wide variety of approaches to the Named Entity task.
-The most successful system used AdaBoost applied to fixed-depth decision trees. Special attention was given to a variety of input features and the features were also
-partly adapted for each language with an F-score of 81.39 on the Spanish test set and 77.05 on Dutch.
+Although there are unsupervised methods, they cannot currently compete with supervised applications.
+For the CoNLL shared task in 2002 \cite{tksintro}, 12 systems participated in the challenge and introduced a wide variety of approaches to the Named Entity task.
+The most successful system used AdaBoost applied to fixed-depth decision trees. Special attention was given to a variety of input features, some which were also
+particularly adapted for each language. I achieved an F-score of 81.39 on the Spanish test set and 77.05 on Dutch.
 The second best system \cite{Florian:2002:NER:1118853.1118863} employed three stacked
-learners (transformation-based learning for obtaining base-level non-typed Named Entites, sparse network of winnows (Snow) for improving the quality of these entities and the forward-backward algorithm for finding the categories for the Named Entites), where the combination of these three algorithms substantially outperformed any subset of them, with an F-score of 79.05 on the Spanish test set and 74.99 on Dutch.
+learners (transformation-based learning for obtaining base-level non-typed Named Entites, sparse network of winnows (Snow) for improving the quality of these entities 
+and the forward-backward algorithm for finding the categories for the Named Entites), where the combination of these three algorithms substantially outperformed 
+any subset of them and yielded an F-score of 79.05 on the Spanish test set and 74.99 on Dutch.
 Among the most prominent algorithms in NER are maximum entropy models, whose benefit in regard to this task became most apparent in CoNLL 2003. 
-For CoNLL 2003, five of the sixteen systems used this statistical learning method and three used maximum entropy models in isolation.
-Maximum entropy models appear to be a suitable choice for the NER task, since systems using this approach were ranked among the three highest results in 
-English and the two highest results in German \cite{TjongKimSang:2003:ICS:1119176.1119195}.
+Five of the sixteen systems that participated used maximum entropy models, which appear to be a suitable choice for the NER task, 
+since systems using this approach were ranked among the highest results in English and in German \cite{TjongKimSang:2003:ICS:1119176.1119195}.
 The best system in both German and English \cite{Florian:2003:NER:1119176.1119201} used a combination of classifiers: a robust linear classifier, 
-a maximum entropy classifier, a transformation-based learner, and a hidden Markov model. The models were connected through a linear interpolation of the classifiers’ % combination scheme
-class probability distribution. Particular importance is also assigned to creating a rich feature space, also including a gazetteer feature. 
-This system achieves a F-score of 88.76\% on the English test set and 72.41\% on the German one.
+a maximum entropy classifier, a transformation-based learner, and a hidden Markov model. The models were connected through a linear interpolation 
+of the classifiers’ % combination scheme
+class probability distribution. Particular importance was also assigned to creating a rich feature space, also including a gazetteer feature. 
+This system achieved a F-score of 88.76\% on the English test set and 72.41\% on the German one.
 Another, more recent approach involved the use of Deep Belief Nets (DBN) for NER in Chinese \cite{Chen:2010:UDB:1870457.1870473}.
-Due to the special characteristics of Chinese of lacking boundary information, character-based methods are generally given preference over word-based methods.
-In their application for NER, DBN  is used to automatically discover the complicated composite effects of the characters to the NE categories from the input data. 
+Due to the language's special characteristics of lacking boundary information, character-based methods are generally given preference over word-based methods.
+In their application for NER, DBN  was used to automatically discover the complicated composite effects of the characters to the NE categories from the input data. 
 The paper reports DBN's strong representative power in regard to detection of more complicated and efficient features. With a precision of 91.45\%,
-it achieves better performance than a SVM and a BP neural networks.
+it achieved better performance than both a SVM and a BP neural networks.
 
@@
 %In 2002 and 2003, the task was Named Entity Recognition. 
-In 2002 and 2003 CoNLL shared tasks were on Named Entity Recognition. The baseline system first decides on the most frequent Named Entity class C in the training data
-and then labels all occurrences of a Named Entity found in the test data with class C regardless of context. 
-Any named entities that did not occur in the training data are also not assigned any class in the test data.
-In case that a phrase was part of more than one entity, the system would choose the longest one \cite{TjongKimSang:2003:ICS:1119176.1119195}.\todo{? hab ich nicht kapiert :D}
+In 2002 and 2003 CoNLL shared tasks were on Named Entity Recognition. The baseline system was constructed by first determining
+on the most frequent Named Entity class C in the training data and then labeling all occurrences of a Named Entity found in the test data with class C regardless of context. 
+Any Named Entities that did not occur in the training data were also not assigned any class in the test data.
+Given that a phrase was part of more than one entity, the system would choose the longest one \cite{TjongKimSang:2003:ICS:1119176.1119195}.
 Table~\ref{table:Base} shows the baselines for the four different languages.  
-The data is annotated with part-of-speech tags and named-entity tags, as well as BIO-tags that indicate the position of a token in the Named Entity. %put in ``tokenized, chunked''
+The data was annotated with part-of-speech tags and Named-Entity tags, as well as BIO-tags that indicate the position of a token in the current Named Entity. %put in ``tokenized, chunked''
 %Our approach is based on \cite{strlearn}, of the group that won the challenge for Spanish and Dutch in 2002.
@@
 \section{Implementation}
-In this section, we describe our implementation, the learning and decoding algorithm and the features we used.
+In this section, we describe our implementation, the learning and decoding algorithm and the features that were used.
 
@@
 Structured Prediction \cite{Taskar:2005:LSP:1102351.1102464} is an approach to supervised learning that is often used in natural language processing. 
-Assuming our goal is to learn a function ${ \boldsymbol{f}\colon \mathcal{X}\to\mathcal{Y} }$, in structured prediction, we assume that $y \in \mathcal{Y}$ is not atomic (e.g.\ binary prediction), but that the prediction is a structure (e.g.\ a syntactic tree for a sentence or a sequence of part-of-speech tags).
+Assuming our goal is to learn a function ${ \boldsymbol{f}\colon \mathcal{X}\to\mathcal{Y} }$, in Structured Prediction, we assume that $y \in \mathcal{Y}$ 
+is not atomic (e.g.\ binary prediction), but that the prediction is a structure (e.g.\ a syntactic tree for a sentence or a sequence of part-of-speech tags).
 
-The structured prediction approach we follow has advantages over traditional approaches like hidden Markov models.
-While a hidden Markov model can model interactions between labels as transition probabilities, it is restricted to tokens and Maximum Likelihood learning.
-Another alternative would be to use a classifier for each token and then concatenate the results, however this approach would not take into account label interactions.
-With structured prediction, we can use features and model label interactions at the same time.
+The structured prediction approach we follow has advantages over traditional approaches like hidden Markov models. While a hidden Markov model can model interactions 
+between labels as transition probabilities, it is restricted to tokens and Maximum Likelihood learning. Another alternative would be to use a classifier for each 
+token and then concatenate the results. However, this approach would not take into account label interactions. With Structured Prediction, we can use features and 
+model label interactions at the same time.
 
@@
 \paragraph*{Node features}
-Node features are features that only depend on the current token (node): the \emph{Token}, suffix and prefix, capitalisation.
+Node features are features that only depend on the current token (node): the \emph{Token}, suffix and prefix and capitalisation.
 
@@
 
-In addition, there is a prepositional phrase (PP) attachment feature, that registers the existence of a this attachment, that would
+In addition, there is a prepositional phrase (PP) attachment feature, that registers the existence of an attachment, that would
 result in a break of the sequence as in
@@
 A technique commonly used in Named Entity Recognition is the use of external lists of names of a specific type (gazetteer). 
-In order to create gazetteer lists for more common named entities, we use a \emph{SPARQL} query  (Figure~\ref{fig:sparql}) 
+In order to create gazetteer lists for more common Named Entities, we use a \emph{SPARQL} query  (Figure~\ref{fig:sparql}) 
 to automatically retrieve entries from DBpedia for all required languages from a public SPARQL endpoint.
@@
 
-One of the issues that we needed to address is the case of headlines. Headlines, mainly in the English data, are often written entirely 
+One of the issues that we needed to address was the case of headlines. Headlines, mainly in the English data, are often written entirely 
 in uppercase letters. This is problematic because from most of the data, the model might learn that a word in all uppercase letters is 
-indicative of an organization (e.g.\ IBM, AIG, NYSE) and might thus mistake the tokens in headlines for organizations. 
+indicative of an organisation (e.g.\ IBM, AIG, NYSE) and might thus mistake the tokens in headlines for organisations. 
 We address this issue by converting all-uppercase headlines into their most likely original case (truecasing). 
@@
 
-The experiment above has shown that the gazetteer features are beneficial for Dutch.
-In this experiment, we want to explore whether the gazetteer features can be improved by extending the monolingual gazetteer with entries from other languages.
-For this, we compare the Dutch-only gazetteer with a gazetteer extended with English data.
+The experiment above has shown that the gazetteer feature is beneficial for Dutch. In this experiment, we wanted to explore whether the 
+gazetteer feature can be improved by extending the monolingual gazetteer with entries from other languages. For this, we compare the 
+Dutch-only gazetteer with a gazetteer extended with English data.
 

### 2012-11-11 doc/report/paper.tex
corrections

@@
 Precision is the percentage of Named Entities found by the system that are correct. Recall is the percentage of Named Entities present in the corpus that were 
-found by the system. A Named Entity is only deemed correct if it is an exact match for the corresponding data file. The general formula for the \emph{F-Score} 
-is shown below. Since we consider \emph{precision} and \emph{recall} as equally important, we use the $F_{1}$ score, which is the harmonic mean of precision and recall.
+found by the system. A Named Entity is only deemed correct if it is an exact match for the corresponding data file. The general formula for the \emph{F-score} 
+is shown below. Since we consider precision and recall as equally important, we use the $F_{1}$-score, which is the harmonic mean of precision and recall.
 \begin{align*}

### 2012-11-12 doc/report/paper.tex
Corrections.

@@
 In this paper, we apply the method of structure prediction to the Named Entity Recognition (NER) information extraction task.
-We illustrate a number of issues relevant to the NER task, describe our implementation and compare it to state-of-the art methods.
+We illustrate a number of issues arising in the NER task, describe our implementation and compare it to state-of-the art methods.
 %We use a structured version of the Perceptron algorithm with averaging and Viterbi decoding.
@@
 
-NER is an information extraction task and has first been introduced as part of the 6th MUC (Sixth Message Understanding Conference)
-that was focusing on the extraction of structured information from unstructured text, such as company names from newspaper articles \cite{nadeau2007survey}.
+NER has first been introduced as part of the 6th MUC (Sixth Message Understanding Conference),
+which focused on the extraction of structured information from unstructured text, such as company names from newspaper articles \cite{nadeau2007survey}.
 The most prominent types of Named Entities are generally known under the name ENAMEX and are comprised of \texttt{Person Names}, \texttt{Organisations} and \texttt{Locations}.
@@
 Apart from the above, there is also the TIMEX type, which covers date and time expressions and NUMEX for monetary values and percentages.
-Table~\ref{table:NETypes} provides an overview of the different Named Entity (NE) types.
+Table~\ref{table:NETypes} provides an overview of the different Named Entity (NE) types and examples.
 
@@
 The difference lies in the reference of the entity. A Named Entity refers to a specific, unique person/time/location, whereas
-\emph{normal} entities can refer to multiple events, such as the time expression: \emph{In June}, which could refer to any year.
-Also, there could be a reference to a person, as in: \emph{the prof}, which in itself is not a \namedentity.
+non-specific entities can refer to multiple events, such as the time expression: \emph{In June}, which could refer to the month June of any year.
+Another example, is a reference to a person in \emph{the prof}, which is not a \namedentity.
 %However, it would be a deictic reference and point back to the mention of a real \namedentity, as in the following coreference chain: \\
@@
 %\ex. $ Prof.\; Bateman \Rightarrow he \Rightarrow the \; prof$ \label{cor}\\
-A Named Entity can consist of more than one syntactic type. Syntactic types feature in syntactic rules that define how sentences can be build.
+A Named Entity can consist of more than one syntactic type. Syntactic types feature in syntactic rules that define how sentences can be built.
 Typical high-level rules include:
@@
 
-Thus, sentences are often represented as syntactic trees that model the dependencies among the types.
-The overall type for a Named Entity is usually a noun phrase (NP), which can be both simple and complex, meaning it can consist of a only a noun phrase or 
-also additional other syntactic types, such as prepositional phrases (PP), that consist of a preposition and a noun phrase. 
+Sentences are often represented as syntactic trees that model the dependencies among the types.
+The overall type for a Named Entity is usually a noun phrase (NP), which can be both simple and complex, meaning it can consist of only a noun phrase or 
+additionally other syntactic types, such as prepositional phrases (PP) that consist of a preposition and a noun phrase. 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% - can be left out if not enough space/too detailed
@@
 Generally, one of the most common challenges for NER is the issue of ambiguity, in particular \emph{Polysemy} \cite{nadeau2007survey}. %and \emph{Metonymy} insert with other section
-Polysemy refers to the property of a lexical representation (\~ word) having more than one possible meaning. Polysemy can become an issue for NER when this lexical representation 
+Polysemy refers to the property of a lexical representation~(word) having more than one possible meanings. Polysemy can become an issue for NER when this lexical representation 
 points to two different Named Entity types.
 This is quite frequent with \texttt{Person Names} and \texttt{Locations}, since many people are named after cities and vice versa, for example \emph{Paris} or \emph{Georgia}. 
-Often, context is not sufficient to disambiguate such names as in example~\ref{ex:Georgia}.
+Often, the textual context is not sufficient to disambiguate such names as in example~\ref{ex:Georgia}.
 
@@
 The results achieved were reasonably good, but presented some flaws since constructed rules are dependent on the language and the domain and are not readily portable. 
-This is the reason why nowadays machine learning methods are more attractive for this task as they offer more adaptability.
+This is the reason why nowadays machine learning methods are more attractive for this task as they offer better adaptability.
 
@@
 Although there are unsupervised methods, they cannot currently compete with supervised applications.
-For the CoNLL shared task in 2002 \cite{tksintro}, 12 systems participated in the challenge and introduced a wide variety of approaches to the Named Entity task.
+For the CoNLL shared task in 2002 \cite{tksintro}, 12 systems participated in a challenge and introduced a wide variety of approaches to the Named Entity task.
 The most successful system used AdaBoost applied to fixed-depth decision trees. Special attention was given to a variety of input features, some which were also
-particularly adapted for each language. I achieved an F-score of 81.39 on the Spanish test set and 77.05 on Dutch.
+particularly adapted for each language. This system achieved an F-score of 81.39 on the Spanish test set and 77.05 on Dutch.
 The second best system \cite{Florian:2002:NER:1118853.1118863} employed three stacked
@@
 The best system in both German and English \cite{Florian:2003:NER:1119176.1119201} used a combination of classifiers: a robust linear classifier, 
-a maximum entropy classifier, a transformation-based learner, and a hidden Markov model. The models were connected through a linear interpolation 
+a maximum entropy classifier, a transformation-based learner, and a hidden Markov model. The models were connected through linear interpolation 
 of the classifiers’ % combination scheme
-class probability distribution. Particular importance was also assigned to creating a rich feature space, also including a gazetteer feature. 
-This system achieved a F-score of 88.76\% on the English test set and 72.41\% on the German one.
+class probability distributions. Particular importance was also assigned to creating a rich feature space, also including a gazetteer feature. 
+This system achieved an F-score of 88.76\% on the English test set and 72.41\% on the German one.
 Another, more recent approach involved the use of Deep Belief Nets (DBN) for NER in Chinese \cite{Chen:2010:UDB:1870457.1870473}.
-Due to the language's special characteristics of lacking boundary information, character-based methods are generally given preference over word-based methods.
-In their application for NER, DBN  was used to automatically discover the complicated composite effects of the characters to the NE categories from the input data. 
+Due to the Chinese language's special characteristics of lacking boundary information, character-based methods are generally given preference over word-based methods.
+In their application for NER, the DBN  was used to automatically discover the complicated composite effects of the characters to the NE categories from the input data. 
 The paper reports DBN's strong representative power in regard to detection of more complicated and efficient features. With a precision of 91.45\%,
-it achieved better performance than both a SVM and a BP neural networks.
+it achieved better performance than SVMs and back-propagation neural networks.
 
@@
 %In 2002 and 2003, the task was Named Entity Recognition. 
-In 2002 and 2003 CoNLL shared tasks were on Named Entity Recognition. The baseline system was constructed by first determining
-on the most frequent Named Entity class C in the training data and then labeling all occurrences of a Named Entity found in the test data with class C regardless of context. 
+%In 2002 and 2003 CoNLL shared tasks were on Named Entity Recognition. 
+The baseline system was constructed by first determining the most frequent Named Entity class C in the training data and then labeling all occurrences of a Named Entity found in the test data with class C regardless of context.
 Any Named Entities that did not occur in the training data were also not assigned any class in the test data.
@@
 Table~\ref{table:Base} shows the baselines for the four different languages.  
-The data was annotated with part-of-speech tags and Named-Entity tags, as well as BIO-tags that indicate the position of a token in the current Named Entity. %put in ``tokenized, chunked''
+The data was annotated with part-of-speech tags and Named-Entity tags in BIO format. %put in ``tokenized, chunked''
 %Our approach is based on \cite{strlearn}, of the group that won the challenge for Spanish and Dutch in 2002.
@@
 \section{Implementation}
-In this section, we describe our implementation, the learning and decoding algorithm and the features that were used.
 
@@
 Assuming our goal is to learn a function ${ \boldsymbol{f}\colon \mathcal{X}\to\mathcal{Y} }$, in Structured Prediction, we assume that $y \in \mathcal{Y}$ 
-is not atomic (e.g.\ binary prediction), but that the prediction is a structure (e.g.\ a syntactic tree for a sentence or a sequence of part-of-speech tags).
+is not atomic (e.g.\ binary prediction), but that the prediction is a structure (e.g.\ the syntactic tree for a sentence or a sequence of part-of-speech tags).
 
@@
 
-Learning is implemented using the Structured Perceptron algorithm. To avoid overfitting, we average the parameters after the last iteration \cite{collins2002discriminative}. 
-For finding the best sequence of labels for a sentence, we use a modified version of the Viterbi algorithm. The algorithm finds the sequence $\hat{y}$, such that:
+Learning is implemented using the structured Perceptron algorithm. To avoid overfitting, we average the parameters after the last iteration \cite{collins2002discriminative}. 
+For finding the best sequence of labels for a sentence, we use a modified version of the Viterbi algorithm.
+
+Let $\mathbf{w}$ be the weight vector and $\boldsymbol{f}(\mathbf{x}, i, y_{i-1}, y_{i})$  be a binary feature function of the input vector $\mathbf{x}$ and the predicted labels at position $i$ and ${i-1}$ ($y_{i}$, $y_{i-1}$).
+The Viterbi algorithm finds the sequence $\hat{y}$, such that:
 \[
@@
 
-This sequence can be computed efficiently in $ O( N^2 |\mathbf{x}| ) $ via dynamic programming. 
-As a first step, the trellis has to be constructed, then we have to find the $ a \in \mathcal{Y}$ in the 
-last column of the trellis with maximal score $\delta_n(a)$. From this, the sequence can be recovered via back-tracking trough the trellis. 
+Given $N$ labels to be predicted, this sequence can be computed efficiently in $ O( N^2 |\mathbf{x}| ) $ via dynamic programming. 
+As a first step, a \todo{erklären?}trellis is constructed and we determine the $ a \in \mathcal{Y}$ in the 
+last column of the trellis that provides the maximal score $\delta_n(a)$. From this, the sequence can be recovered via back-tracking trough the trellis. 
 The score of the best sequence ending in $a$ is:
@@
 \noindent This function $\delta_i(a)$ can be defined recursively as:
-
 \begin{align*}
 \delta_1(a) &= \mathbf{w} \cdot \boldsymbol{f}(\mathbf{x}, 1, \emptyset, a) \\
-\delta_i(a) &= \max_{b \in \mathcal{Y}} \delta_{i-1}(b) + \mathbf{w} \cdot \boldsymbol{f}(\mathbf{x}, i, b, a) \\
+\delta_i(a) &= \max_{b \in \mathcal{Y}} \delta_{i-1}(b) + \mathbf{w} \cdot \boldsymbol{f}(\mathbf{x}, i, b, a) \
 \end{align*}
@@
 \paragraph*{Node features}
-Node features are features that only depend on the current token (node): the \emph{Token}, suffix and prefix and capitalisation.
+Node features are features that only depend on the current token (node), e.g. the token itself, its suffixes and prefixes and its capitalisation.
 
-Table~\ref{table:node} shows the various node features with an example respectively.
+Table~\ref{table:node} illustrates node features used in our system with examples.
 
@@
 \hline
-Token &  \textbf{Amsterdam}\\
-Suffix& Amster\textbf{dam}\\
-Prefix&  \textbf{Mc}Cartney\\
-Captitalized& \textbf{B}enetton\\
-UPPERCASE &  \textbf{BENETTON}\\
-Number patterns & 2005\\
-POS tag &  Benetton \textbf{NNP}   \\
-Lemma & produced $\Rightarrow$ produce \\
+Token &  Amsterdam $\Rightarrow$ Amsterdam \\
+Suffix (2-4)& \emph{Amsterdam}  $\Rightarrow$ am, dam, rdam\\
+Prefix (2-4)&  \emph{McCartney}  $\Rightarrow$ Mc, McC, McCa\\
+Capitalised?& \emph{Benetton}  $\Rightarrow$ True\\
+UPPERCASE? &  \emph{BENETTON} $\Rightarrow$ True\\
+Number patterns & \emph{2005}  $\Rightarrow$ year pattern\\
+POS tag &  \emph{Benetton}  $\Rightarrow$  NNP   \\
+Lemma & \emph{produced} $\Rightarrow$ produce \\
 
@@
 \exg. Bank of Scotland . \\ 
-      ORG   0    ORG   0  \\\label{seq3}
+      ORG  O  ORG      O  \\\label{seq3}
 
 \paragraph*{Gazetteer Features}
-A technique commonly used in Named Entity Recognition is the use of external lists of names of a specific type (gazetteer). 
-In order to create gazetteer lists for more common Named Entities, we use a \emph{SPARQL} query  (Figure~\ref{fig:sparql}) 
-to automatically retrieve entries from DBpedia for all required languages from a public SPARQL endpoint.
+A technique commonly used in Named Entity Recognition is the utilisation of external lists of names for a specific type (gazetteer). 
+In order to create gazetteer lists for the more common types of Named Entities, we use a \emph{SPARQL} query 
+to automatically retrieve entries from DBpedia for all required languages from a public SPARQL endpoint (Figure~\ref{fig:sparql}).\todo{citation fur SPARQL?}
 
@@
 
-Given the lists of entities retrieved, we then mark the longest matches of known entities in the input using a directed acyclic word graph. 
-The mention is annotated with a reference to its list of origin, which is then used as a feature in our model. With this approach, it is not 
-necessary to manually map the lists to target labels and the model can learn the reliability of each list.
+Given the retrieved lists of entities, we then mark the longest matches of known entities in the input using a directed acyclic word graph. \todo{cit.}
+The mention is annotated with a reference to its list of origin, which is then used as a feature in our model.
+With this approach, it is not necessary to manually map the lists to target labels and the model can learn the reliability of each list.
+The gazetteer features also distinguish whether a gazetteer entry consist only of a single token or of two or more tokens.
+This distinction is useful since single tokens are likely less reliable than longer matches.
 
@@
 
-One of the issues that we needed to address was the case of headlines. Headlines, mainly in the English data, are often written entirely 
-in uppercase letters. This is problematic because from most of the data, the model might learn that a word in all uppercase letters is 
+One of the issues that we needed to address was the capitalisation of headlines. Headlines, mainly in the English data, are often written entirely 
+in uppercase letters. This is problematic since from most of the data, the model might learn that a word in all uppercase letters is 
 indicative of an organisation (e.g.\ IBM, AIG, NYSE) and might thus mistake the tokens in headlines for organisations. 
 We address this issue by converting all-uppercase headlines into their most likely original case (truecasing). 
-To estimate the most likely case of a token from a monolingual corpus, we use the truecasing model from the 
+To do this, we estimate the most likely case of a token from a monolingual corpus by using the truecasing model from the 
 Moses statistical machine translation system \cite{koehn2007moses}.
@@
 \hline
-\bf Language & \multicolumn{3}{c|}{ \bf Test A}&\multicolumn{3}{c|}{ \bf Test B}\\
+\bf Language & \multicolumn{3}{c|}{ \bf Dev Set}&\multicolumn{3}{c|}{ \bf Test Set}\\
              & Precision & Recall & $F_1$ & Precision & Recall & $F_1$ \\ \hline
@@
 
-For the evaluation of the system, we used the standard measure of \emph{precision} and \emph{recall} as shown below.
+For the evaluation of the system, we used the standard measures of \emph{precision} and \emph{recall} as shown below.
 \begin{align*}
@@
 
-\section{Discussion}
+\section{Discussion and Future Work}
 
-\section{Future Work and Conclusion}
-
-In this paper, we presented a \namedentity~Recognition system based on the concept of Structured Prediction. 
+In this paper, we have presented a \namedentity~Recognition system based on the concept of Structured Prediction.
 While the system does not achieve state-of-the-art performance on the CoNLL shared task, we believe that the results 
 are reasonable for the following reasons: We are using a simple model, which has the advantage that the final application
-will be fast and it is easily extendable. In contrast to the best systems in the CoNLL task, we also do not use ensemble 
+will be fast and it is easily extendable. In contrast to the best systems in the CoNLL tasks, we do not use ensemble 
 learning methods and we use the same set of features for all languages, which enables simple extension of the system to additional languages.
+
+We have demonstrated that the structured Perceptron with averaging is a suitable learning algorithm for the NER task.
+In future work, we plan to use ensemble methods on top of the basic model and experiment with using longer state histories in model.
 
